### 2022-01-23 1_methods.tex
Initial Overleaf Import

\subsection{The continuous constant pH molecular dynamics framework}

In continuous CpHMD, each titratable residue is assigned a titration coordinate, $\lambda_i$ \cite{Lee_Brooks_2004_Proteins,Khandogin_Brooks_2005_Biophys.J.}. 
We attribute to this coordinate an implicit function of an underlying coordinate $\theta_i$ such that $\lambda_i$ is bound between 0 and 1:
\begin{equation}
    \lambda_i = \sin^2\theta_i,
\end{equation}
where 0 and 1 represent the protonated and deprotonated states, respectively. 
Since $\lambda$ continuously evolves between 0 and 1,
in simulation analysis we apply cutoffs to define
protonated ($\lambda^P < 0.2$) and deprotonated states ($\lambda^U > 0.8$). 
Having the titration coordinate in place, we use an extended Hamiltonian to allow the simultaneous propagation of spatial (real) and titration (virtual) coordinates:

\begin{equation}
\begin{split}
\label{eq:Hamiltonian}
\mathcal{H}(\{\bm{r}_a\},\{\theta_i\})=
&\frac{1}{2}\sum_am_a\dot{\bm{r}}_a^2+\frac{1}{2}\sum_im_i\dot{\theta}_i^2+U^{\rm int}(\{\bm{r}_a\})\\
&+U^{\rm hybr}(\{\bm{r}_a\},\{\theta_i\})+\sum_iU^{\ast}(\theta_i),
\end{split}
\end{equation}
where $\bm{r}_a$ refers to the (x,y,z) coordinates
of atom $a$, and $\theta_i$ refers to the titration coordinate of titratable residue $i$. 
The two first terms of Eq.~\ref{eq:Hamiltonian} give the kinetic energies of real atoms and virtual $\lambda$ particles. 
$U^{\rm int}$ represent the titration-independent bonded and non-bond energies. 
The fourth term, $U^{\rm hybr}$, describes the 
non-bond energies, which are dependent on both spatial and titration coordinates. 
For the particle-mesh Ewald (PME) all-atom CpHMD
\cite{Huang_Shen_2016_J.Chem.TheoryComput.}
$U^{\rm hybr}$ includes only Coulomb and van der Waals (vdW) energies, while for
the implicit- and hybrid-solvent CpHMD methods $U^{\rm hybr}$ also includes the generalized Born (GB) energy:

 \begin{equation}
\label{eq:Uhybr}
\begin{split}
U^{\rm hybr}(\{\bm{r}_a\},\{\theta_i\})=U^{\rm elec}(\{\bm{r}_a\},\{\theta_i\}) 
+ U^{\rm vdW}(\{\bm{r}_a\},\{\theta_i\})\\
+ U^{\rm GB}(\{\bm{r}_a\},\{\theta_i\})
\end{split}
\end{equation}

The dependence of the electrostatic energy on the titration coordinates arises from the requirement that
the partial atomic charges on the titrating residue
are linearly interpolated  
between the values in the deprotonated and protonated states with respect to $\lambda$:

% Thus, we can write the charge on atom $j$, $q_j$, on the titrating residue $i$ as
%
\begin{equation}
q_j(\lambda_i) = (1-\lambda_i) q_j^{\rm prot} +  \lambda_i q_j^{\rm unprot},
\label{eq:qlamb}
\end{equation}
where $q_j$ refers to the partial charge of an atom $j$ in the titrating residue $i$.
As to vdW energies, the 
interactions involving a titratable proton are linearly attenuated with respect to $\lambda$:
\begin{equation}
U^{\rm vdW}_{ij}(\bm r_i, \bm r_j, \lambda_i) = (1-\lambda_i)U^{\rm vdW\ast}_{ij}(\bm r_i, \bm r_j),
\end{equation}
where $i$ is the index for the titratable proton, $j$ is the index for all other atoms, and $U^{\rm vdW \ast}_{ij}$ refers to the protonation independent vdW interaction energy. If both $i$ and $j$ are titratable protons,
the interaction is linearly attenuated by the $\lambda$ values of both:
\begin{equation}
U^{\rm vdW}_{ij}(\bm r_i, \bm r_j, \lambda_i, \lambda_j) = (1-\lambda_i)(1-\lambda_j)U^{\rm vdW\ast}_{ij}(\bm r_i, \bm r_j).
\end{equation}
Note, the vdW energy correction is small and typically neglected in constant pH methods except for the CpHMD methods implemented in CHARMM and Amber \cite{Khandogin_Brooks_2005_Biophys.J.,Wallace_Shen_2011_J.Chem.TheoryComput.,Huang_Shen_2016_J.Chem.TheoryComput.,Huang_Shen_2018_J.Chem.Inf.Model.,Harris_Shen_2019_J.Chem.Inf.Model.}.

Finally, the last term in Eq. \ref{eq:Hamiltonian} 
only affects titratable groups and is consisted of 
three biasing potentials:
\begin{equation}
\label{eq:biasingpotential}
U^{\ast}(\theta_i)=-U^{\rm mod}(\lambda_i)+U^{\rm barr}(\lambda_i)+U^{\rm pH}(\lambda_i).
\end{equation}

The first biasing potential $U^{\rm mod}$ describes a potential of mean force (PMF) of model titration along the $\lambda$-coordinate, where the model represents a fully solvent-exposed amino acid, i.e., a blocked single amino acid or a small peptide in solution.
According to the linear response theory, 
model PMF is (in GB solvent) or can be approximated (in explicit solvent) as a quadratic function:
\begin{equation}
\label{eq:Umod_1d}
U^{\rm mod}(\lambda_i)=A(\lambda_i-B)^2.
\end{equation}
Here $A$ and $B$ are parameters that can be determined through free energy calculation methods, such as thermodynamic integration (TI, see section Parameterization). 
By analogy, in the case of coupled double-site titration \cite{Khandogin_Brooks_2005_Biophys.J.}, e.g., histidine or carboxylic acid, the model PMF is second order in both $\lambda$ and $x$, where 
$x$ is a coordinate that represents the tautomer interconversion.
The two-dimensional PMF can be written as a bivariate polynomial,
%
\begin{equation}
\begin{split}
U^{\rm mod}(\lambda_i,x_i) =
&a_0 \lambda_i^2 x_i^2 + a_1 \lambda_i^2 x_i + a_2 \lambda_i x_i^2 + a_3 \lambda_i^2 \\
&+ a_4 x_i^2 + a_5 \lambda_i x_i + a_6 \lambda_i + a_7 x_i + a_8.
\end{split}
\label{eq:double}
\end{equation}
Here $x_{i}=\sin^{2}\theta^{x}_{i}$ is bound between 0 and 1, where $\theta^{x}_{i}$ is the underlying unbound
variable. 
%

$U^{\rm barr}$, the second biasing potential in Eq. \ref{eq:biasingpotential}, describes a barrier or penalty potential in the center of the titration coordinate, 
%
\begin{equation}
U^{\rm barr}(\lambda_i) = -4\beta\left(\lambda_i - \frac{1}{2} \right)^2,
\label{eq:Ubarr}
\end{equation}
where $\beta$ is a parameter that determines the height of the barrier. 
Adding the barrier potential serves to
increase the sampling time at the end-point states (e.g., $\lambda<0.2$ or $\lambda>0.8$) and therefore
minimizing the unphysical vdW and electrostatic interactions associated with mixed state 
(e.g., $0.2 \le \lambda \le 0.8$).

$U^{\rm pH}$, the last term of Eq.~\ref{eq:biasingpotential} 
describes the pH dependence of the deprotonation free energy 
and is given by
%
\begin{equation}
U^{\rm pH}(\lambda_i)=ln(10)k_BT({\rm pH}-\textrm{p}{K_a}^{\rm mod})\lambda_i, 
\label{eq:UpH}
\end{equation}
% 
where $k_B$ is the Boltzmann constant, $T$ is the system temperature, and $\textrm{p}{K_a}^{\rm mod}$ is the model {\pka} for the titratable residue $i$.

### 2022-01-23 1_scope.tex
Initial Overleaf Import

\subsection{Scope} % - Ruibin 
This tutorial mainly covers two CpHMD methods that
have been extensively validated and applied to various real-life problems:
1) the hybrid-solvent CpHMD \cite{Wallace_Shen_2011_J.Chem.TheoryComput.}
in CHARMM \cite{Brooks_Karplus_2009_J.Comput.Chem.}
and
2) the GBNeck2 implicit-solvent CpHMD
\cite{Huang_Shen_2018_J.Chem.Inf.Model.,Harris_Shen_2019_J.Chem.Inf.Model.}
in Amber \cite{Case_Kollman_2018_}. Currently, fully patched Amber 18 and 20 released versions are supported. GBNeck2-CpHMD patches to them can be downloaded from our \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-patches}{cphmd-patches} repository. Details about how to apply them are described on that repository page.
The CHARMM hybrid-solvent CpHMD currently supports CPUs only, 
but the Amber GBNeck2-CpHMD supports both CPU or GPU computing.
We strongly recommend running the CpHMD simulations
via the pH replica-exchange protocol to significantly accelerate convergence of \pka's and conformational sampling; however, if desired, a single pH 
simulation or a set of independent pH simulations can also 
be conducted.

In addition to the above two methods, we will briefly
discuss the PME all-atom CpHMD method implemented in
CHARMM for CPU computing \cite{Huang_Shen_2016_J.Chem.TheoryComput.}
and in Amber for GPU computing (Harris, Liu, and Shen, unpublished data). 
We note, this tutorial does not cover the GBSW implicit-solvent CpHMD method in CHARMM 
\cite{Lee_Brooks_2004_Proteins,Khandogin_Brooks_2005_Biophys.J.,Khandogin_Brooks_2006_Biochemistry,Arthur_Brooks_2016_J.Comput.Chem.a}, which is available for both CPU and GPU computing.
The GBSW-CpHMD method with the temperature replica-exchange protocol has been extensively validated for protein \pka\ 
predictions and pH-dependent conformational dynamics.

### 2022-01-23 2_prereq.tex
Initial Overleaf Import

\section{Prerequisites} % - Ruibin
CpHMD is a specialized MD technique, so the user should have some familiarity with MD simulations and especially with the Amber or CHARMM packages. 
Users with experience with other MD engines, such as Gromacs
\cite{Abraham_Lindahl_2015_SoftwareX}, NAMD\cite{Phillips_Tajkhorshid_2020_J.Chem.Phys.}, 
or OpenMM
\cite{Eastman_Pande_2017_PLoSComput.Biol.} 
are encouraged to go through the introductory tutorials for Amber or CHARMM to get started. 
Before running a CpHMD simulation, it is important to know your research objective. 
If the objective is to predict the {\pka} values of soluble proteins or their protonation states at a certain pH condition, GBNeck2-CpHMD 
\cite{Huang_Shen_2018_J.Chem.Inf.Model.,Harris_Shen_2019_J.Chem.Inf.Model.}
is the best choice and Amber \cite{Case_Kollman_2018_}
experience is required. 
If the objective is to investigate the detailed proton-coupled conformational dynamics, ion transfer, protein-ligand binding/unbinding,
or transmembrane proteins, then 
the hybrid-solvent CpHMD  \cite{Wallace_Shen_2011_J.Chem.TheoryComput.,Huang_Shen_2016_Nat.Commun.} 
or the all-atom PME CpHMD \cite{Huang_Shen_2016_J.Chem.TheoryComput.}
in CHARMM
\cite{Brooks_Karplus_2009_J.Comput.Chem.}
should be used.
We note, a GPU-accelerated all-atom PME CpHMD 
implementation in Amber is forthcoming (Harris, Liu, and Shen, to be published). 
Like all computational chemistry calculations, it is important to know your system as much as possible before any serious simulations. 
With such knowledge, we can decide what salt concentration, temperature, pH range, and titratable residue types to use in CpHMD. Also, the research objective dictates the simulation length. For example, to predict protonation states, simulation can be run until the unprotonated fractions at all pH are converged, typically 10-50 ns per replica. If however, a large conformational transition is of interest, the simulation should be much longer.

\subsection{Background knowledge}
Knowledge of classical MD is required to understand what we can and cannot achieve through a MD simulation. 
The user should also understand the parameters used in energy minimization, heating, equilibration, and production runs. 
Of course, chemistry knowledge of pH, {\pka}, and protonation states and how they may impact the physical properties of the specific biological or material system is essential for obtaining meaningful results.

CpHMD simulations are run on a workstation or high-performance computing (HPC) cluster under a Linux operating system. 
The user is expected to have a good knowledge of Linux commands for accessing and editing files, submitting jobs to servers, and retrieving simulation result files. 
Shell script and preferably also Python skills are required for pre- and post-processing files. 
For simulations on a HPC cluster, the user is required to be familiar with the specific batch scheduler and queuing system, e.g., SLURM (Simple Linux Universal Resource Manager), SGE (Sun Grid Engine), or PBS (Portable Batch System). 
The user is expected to know commands for submitting and checking the status of jobs.
The user is also expected to set environmental variables required for the specific MD engine.
Nowadays, a CpHMD job can also be performed on a workstation with a single or multiple GPUs. 
In this case, the user is required to use commands such as `nvidia-smi' to check the GPU status and `nvcc --version' to check the CUDA version as well. 
The user can use any data visualisation tool(s), but matplotlib in Python and Jupyter notebooks are recommended in our analysis tool set.

\subsection{Software/system requirements} 
For GBNeck2-CpHMD, Amber16 is the minimum version. 
Requirements for installing Amber can be found in the \href{https://ambermd.org/Installation.php}{Amber website}. 
For the GPU version (pmemd.cuda), a Nvidia GPU card with at least compute compatibility sm3.0 should be installed with the latest driver. 
CUDA 10.1 is the primary CUDA version for running it but early versions from CUDA 7.5 also works. 
After successfully installing CUDA, it is important to set three environmental variables in Linux. 
\textbf{CUDA\_HOME} and \textbf{LD\_LIBRARY\_PATH} must point to the directories where the desired CUDA version is installed. 
We also need to specify \textbf{CUDA\_VISIBLE\_DEVICES} to GPU device indices. For example, 
\begin{lstlisting}
$ export CUDA_HOME=/usr/local/cuda-10.1
$ export LD_LIBRARY_PATH=
            /usr/local/cuda-10.1/lib64
$ export CUDA_VISIBLE_DEVICES=0,1
\end{lstlisting}
These variables will direct the Linux system to use the CUDA 10.1 version installed in the /usr/local directory.
In the example above two GPU cards are available,
and the indices 0 and 1 for CUDA\_VISIBLE\_DEVICES indicate the first and second GPU to use, respectively. 
Following installation of Amber, the user should set the environmental variable \textbf{AMBERHOME} to the Amber installation folder. 
\begin{lstlisting}
$ export AMBERHOME=/path_to_Amber_version_xx/
\end{lstlisting}

Before preparing CpHMD inputs of a protein system, 
we need to clean up and fix problems in the PDB file,
e.g., removing non-protein atoms, adding missing residues and atoms, and fixing non-canonical amino acid residues. 
To do that, we can use the Perl script convpdb.pl in \href{http://blue11.bch.msu.edu/mmtsb/Main_Page}{mmtsb tool set} \cite{Feig_Brooks_2004_J.Mol.Graph.Model.} or \href{https://github.com/openmm/pdbfixer}{pdbfixer} in OpenMM \cite{Eastman_Pande_2017_PLoSComput.Biol.}. 
For example, the following commands can be used 
to fix the file 1vii.pdb downloaded from RCSB (\href{www.rcsb.org}{www.rcsb.org}) \cite{Berman_Bourne_2000_NucleicAcidsRes.}.  
\begin{lstlisting}
$ convpdb.pl -nohetero -fixcoo -renumber 1 -out
amber -chain A -segname 1vii.pdb 
\end{lstlisting}
or 
\begin{lstlisting}
$ pdbfixer --add-atoms=heavy 
--keep-heterogens=none --add-residues
--replace-nonstandard 1vii.pdb
\end{lstlisting}
 Alternatively, we can use a homology modeling software such as \href{https://swissmodel.expasy.org/}{SWISS-MODEL} \cite{Waterhouse_Schwede_2018_NucleicAcidsRes.} to fix those problems especially when the number of continuously missing residues is large \cite{Waterhouse_Schwede_2018_NucleicAcidsRes.}.

 The next step is to add capping ligands ACE and NH2 to the N- and C-terminus and to modify ASP, GLU and HIS to AS2, GL2, and HIP residues if these residue types are set as titratable. For AS2 and GL2, a library file \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/blob/master/Files/phmd.lib}{phmd.lib} and a force field modification file \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/blob/master/Files/frcmod.phmd}{frcmod.phmd} are included in Ambertools after version 19 \cite{Case_Kollman_2018_}, or we can load them to tLeap/Leap from the directory containing the two files. The last step is to use tLeap/Leap in Amber/ Ambertools to generate topology and parameter files. Our \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep}{cphmd-prep} tool set uses the pdbfixer tool for the first step and can do all the aforementioned steps to prepare GBNeck2-CpHMD compatible input files.

We use a Python-based \href{https://gitlab.com/shenlab-amber-cphmd/async_ph_replica_exchange}{asynchronous replica exchange} implementation for GPU pH replica-exchange simulations. 
Scripts in \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-analysis}{cphmd-analysis} are used for analyzing CpHMD related files, including convergence checking, {\pka} calculation, and replica-exchange profiling. 
Some simple trajectory based analysis such as RMSD can also be done, but the users can use any tools, e.g., CPPTRAJ \cite{Roe_Thomas_2013_J.Chem.TheoryComput.}, PTRAJ \cite{Roe_Thomas_2013_J.Chem.TheoryComput.}, VMD \cite{Humphrey_Schulten_1996_J.Mol.Graph.}, and MDAnalysis \cite{Michaud-Agrawal_Beckstein_2011_J.Comput.Chem.}. 
% Note that these free tools usually depends on other free softwares, including OpenMM, Ambertools, scipy, pandas, matplotlib, etc. 
% The instructions on how to install those software are covered in their individual pages and not discussed here.

A suggestion to use Python dependent packages is to use Anaconda or Miniconda. With them, the user can create a virtual environment specific for CpHMD related calculations. 
For example, after installing Anaconda or Miniconda, we can type 
\begin{lstlisting}[language=bash]
$ conda create --name cphmd python=3.7
\end{lstlisting}
Note that we choose Python version 3.7 because the OpenMM package currently only supports up to Python 3.7. 
Activate the environment named 'cphmd' by
\begin{lstlisting}[language=bash]
$ conda activate cphmd
\end{lstlisting}
Many packages can be installed through the conda and pip tool.
\begin{lstlisting}[language=bash]
$ conda install -c omnia pdbfixer parmed
$ conda install -c conda-forge ambertools
                 mdanalysis matplotlib 
$ conda install -c anaconda numpy scipy pandas 
$ pip install f90nml
\end{lstlisting}
Whenever the user wants to do some calculations or analysis related to CpHMD, they can activate this 'cphmd' virtual environment and use the installed tools without causing conflict with other software. 
% Python Packages and Version required for Async-Replica Exchange 
% Python Packages and Version required for the CpHMD-Analysis Lambda Parser

For running the hybrid-solvent CpHMD \cite{Wallace_Shen_2011_J.Chem.TheoryComput.}
in CHARMM \cite{Brooks_Karplus_2009_J.Comput.Chem.}
you will need to follow the installation instructions in the \href{https://www.charmm.org/archive/charmm/documentation/installation/}{CHARMM documentation}.
We suggest including the MPI and the Hamiltonian replica exchange options.
\begin{lstlisting}[language=bash]
 $ ./install.com gnu xxlarge M mpif90 +CMPI
 +REPDSTR +ASYNC_PME +GENCOMM
 \end{lstlisting}
If you want to install CHARMM on a local workstation for building and preparing systems or for running a single pH simulation, you can use the following installation command.
\begin{lstlisting}[language=bash]
 $ ./install.com gnu xxlarge
 \end{lstlisting}

% \section{Content and links}

% A tutorial will normally draw on additional files and materials; clearly indicate where and how these are available, with links, and how they are being archived for the long-term and maintained so they stay current.
% You will likely want to reference your GitHub repository as a central point to access all of this information, and then the GitHub repository may link out to other content as needed.

% \section{Checklists}
% Tutorials do not necessarily require the use of a checklist as in Best Practices documents; however, they can include these if desired.
% Several useful checklist formats are available, with examples presented in \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates}.
% One example is shown here.

### 2022-01-23 5_gphmd.tex
Initial Overleaf Import

\section{Workflow for implicit-solvent CpHMD in Amber} 
% - Ruibin

Here we discuss how to run GBNeck2-CpHMD \cite{Huang_Shen_2018_J.Chem.Inf.Model.,Harris_Shen_2019_J.Chem.Inf.Model.} simulations in Amber.
The related files can be found in the GitLab directory 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/gphmd_Amber}{cphmd-tutorial/gphmd\_Amber}.

\subsection{Preparation of structure and input files}
\begin{checklist}{Files generated by cphmd\_prep.sh}
\begin{itemize}
\item Cleaned up PDB file (.pdb)
\item Coordinates and topology file (.rst7)
\item CpHMD job control file (.phmdin)
\item CpHMD parameter file (.parm)
\item Minimization input file (\_mini.mdin)
\item Equilibration input files (\_equil*.mdin)
\item Template production file (template*prod.mdin)
\item Instructions and commands to run CpHMD (README)
\item Python script to run asynchronous pH replica exchange CpHMD if 'async' is specified (apHrex.py)
\item Amber group file to run pH replica-exchange CpHMD if 'reex' is specified (cphmd.groupfile)
\end{itemize}
\end{checklist}

To simplify system preparation for GBNeck2-CpHMD simulations,
we recommend the user to download and install the tool set \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/}{CpHMD-prep}.
Once installed, the shell script \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/tree/master/cphmd_tools/cphmd_prep.sh}{cphmd\_prep.sh} can be called to prepare the structure and input files.
% In the near future, a web server will be provided for this task.
%for preparing the Amber GBNeck2-CpHMD inputs and instructions to run related CpHMD simulations by one line bash command via the 'CpHMD\_Prepare.sh' script. 
\begin{lstlisting}
cphmd_prep.sh 
[-pdb|--pdbid] | [-fil|--filename]
[-cha|--chainid] 
[-mod|--modelid] 
[-con|--conc] 
[-tim|--time] 
[-tem|--temp]
[-dph] 
[-phl|--phlow]
[-phu|--phup] 
[-res|--restype]
[-run|--runmode] 
[-h|--help]
\end{lstlisting}
%[-tes|--test]
%[-ot|--outfiletype] 
The options of 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/tree/master/cphmd_tools/cphmd_prep.sh}{cphmd\_prep.sh}
are explained here.
 \textbf{-pdb|--pdbid}: the PDB ID of the protein used in CpHMD calculations where the corresponding PDB file is downloaded from RCSB; 
\textbf{-fil|--filename}: the PDB file or path used in CpHMD calculations; 
\textbf{-cha|--chainid}: chain IDs to include in simulations with default 'A'; 
\textbf{-mod|--modelid}: the rotamer model ID in the PDB file;
\textbf{-con|--conc}: ionic strength in M with the default 0.15; 
\textbf{-tim|--time}: simulation length in ns for each pH condition (default 10); 
\textbf{-tem|--temperature}: simulation temperature in K (default 300); 
\textbf{-phl|--phlow}: lowest pH (default 6.5);
\textbf{-phu|--phup}: highest pH (default 9.0);
\textbf{-phi|--pHintvl}: pH interval for calculating the individual
simulation pH conditions between phlow and phup; 
\textbf{-res|--restype}: amino acid residue types allowed to be titratable (default 'Asp Glu His Cys Lys'); 
\textbf{-rm|--runmode}: type of simulation, i.e., 'async' for asynchronous pH replica exchange, 
'rex' for traditional synchronous pH replica exchange,
and 'ind' for independent pH simulations; 
\textbf{-t|--test}: if 'T', temporary files are kept for checking;
\textbf{-h|--help}, to display the usage and options.

\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/tree/master/cphmd_tools/cphmd_prep.sh}{cphmd\_prep.sh}
accepts either a PDB ID
or a user-prepared PDB file and calls PDBFixer in
OpenMM\cite{Eastman_Pande_2017_PLoSComput.Biol.}
to extract the desired chain, add missing heavy atoms 
or residues and the terminal capping groups 
(CH$_3$CO and NH$_2$).
Note, ions, ligands, waters, 
and unspecified chains from the PDB file are removed.
% The option -mod|--modelid (default 'A') can be used to specify a particular sidechain rotamer model in the PDB file.
Once the PDB file is cleaned up, tleap in Amber (or Ambertools) is used to build the positions of hydrogen as well as dummy hydrogen atoms according to the specified titratable residue types.
As required by GBNeck2-CpHMD, \cite{Huang_Shen_2018_J.Chem.Inf.Model.,Harris_Shen_2019_J.Chem.Inf.Model.} 
the GB input radii of HD1/HE2 of His, SG of Cys, OD1/OD2 of Asp, and OE1/OE2 of Glu are changed to 1.17, 2.00, 1.40, and 1.40 {\AA} respectively, and the interactions between the dummy carboxylate hydrogens in Asp/Glu are excluded. 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/tree/master/cphmd_tools}{cphmd\_prep.sh}
automatically generates input files as well.
Below we give two examples.

\paragraph{Example 1. Predict the {\pka's} of the EGFR kinase given a PDB ID.}
Here we wish to predict 
the {\pka's} and 
the protonation states
of His, Cys, and Lys at the physiological pH 7.4.
%
\begin{lstlisting}
$ cphmd_prep.sh -pdb 5U8L -mod A -phl 5.5 
-phu 9.5 -res 'His Cys Lys'
\end{lstlisting}
%
Here 5U8L is the PDB ID.
We omit chain ID, because 5U8L contains only one chain (default is chain A), but we specify the model A (there are two rotamer models in the PDB file). 
For PDB file containing NMR models, a specific model can be selected using the -m flag. 
The pH range is 5.5 to 9.5, which extends 2 pH
units above and below the physiological pH 7.4.
Since the (default) pH interval is 0.5 units, the total number of simulation pH conditions is 9. 
In this example, simulation length (-l|--simlength), temperature (-tp|--temperature), and ionic strength (-i|--ionic) are set to the respective default values of 10 ns, 300 K, and 0.15 M.

Upon execution, cphmd\_prep.sh generates a file
`cphmd\_prep.log' and a folder
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/gphmd_Amber/gphmd_kinase_aphrex}{5U8L}, which
contains the files listed in the aforementioned checklist. 
`cphmd\_prep.log' contains the information of the job control parameters and also the names of the files generated for the current job. 
Inside the folder \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/gphmd_Amber/gphmd_kinase_aphrex}{5U8L},
there are multiple files with names like 5U8L\_chainA\_A\_TAG.EXT with different tags TAG and extensions EXT. 
'5U8L' is the PDB ID provided, 'chainA' refers to chain A, and 'A' is the rotamer model ID  provided. Tags include mini for minimization, 
equil* for different equilibration steps, 
and prod for production. 
Extensions include pdb for the PDB file, 
rst7 for the Amber rst7 coordinate file, 
parm for the Amber parm7 parameter file, 
mdin for the Amber input file, 
and phmdin for CpHMD control parameter file. 
We can also specify multiple chains.   
For example, if we want to include both chain A and chain B, 
the command option -cha 'A B' can be used, 
and the file names will start with 5U8L\_chainAB. 
All files in the folder`5U8L' are also zipped as 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/gphmd_Amber/gphmd_kinase_aphrex}{cphmd\_inputs.zip}.

% As implied in the cphmd\_prep.log file, the default mode for running a GBNeck2-CpHMD simulation is asynchronous pH replica exchange, which can be changed to independent pH or classical pH replica exchange modes through specifying the -rm|--runmode flag from 'async' (default) to 'ind' or 'reex'. 
% The README file mentioned in the job\_log file which has instructions of how to run a CpHMD simulation with the provided files. 
% The whole files are in the \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/gphmd\_Amber/gphmd\_snase}{gphmd\_snase} folder of the \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial}{cphmd-tutorial} GitLab repository.

\paragraph{Example 2. Predict the {\pka's} of the protein Snase given a PDB file.}
%
\begin{lstlisting}
$ cphmd_prep.sh -fil 3BDC.pdb -mod A -phl 1.0
-phu 12.0 -tim 4
\end{lstlisting}
%
Here the script takes a PDB filename and the simulations are conducted in a wider pH range from 1.0 to 12.0. 
With a default interval of 0.5 units, this
amounts to 23 simulation pH conditions.
Since the model {\pka's} of Asp/Glu are 3.7/4.3 and 
the model {\pka} of Lys is 10.4, this pH range 
typically allows the titration of
all default residue types 'Asp Glu His Cys Lys'.
For this reason, the residue type (-type) is omitted. 
Another difference from Example 1 is that a much shorter simulation length of 4 ns (-tim 4) is specified.
This is because only a rough estimate of the {\pka} values is desired. 
Similar files are generated as in Example 1; however, 
there are 14 extra Amber input files due to the additional pH conditions.
All related files can be found here \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/gphmd_Amber/gphmd_snase_aphrex/3BDC}{3BDC}.

Several additional files are generated by cphmd\_prep.sh.
%used in production runs depending on running mode specified by the user. 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/gphmd_Amber/gphmd_kinase_aphrex/5U8L/apHREX.py}{apHREX.py}
is a Python script for running asynchronous pH replica exchange simulations if -run is not specified or specified as async. 
The related file
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/gphmd_Amber/gphmd_snase_phrex/3BDC/cphmd.groupfile}{cphmd.groupfile} 
is an Amber style for specifying individual replica files in pH replica-exchange simulations. 
Finally, the file \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/gphmd_Amber/gphmd_snase_phrex/3BDC/README}{README} contains all the commands for successfully running GBNeck2-CpHMD simulations for the current job.

\subsection{Running GBNeck2-CpHMD simulations}
Here we discuss the protocol of running GBNeck2-CpHMD simulations using the two examples discussed above.
All commands for minimization, equilibration, and production can also be found in the file
\href{
https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/gphmd_Amber/gphmd_snase_phrex/3BDC/README}{README}.

\paragraph{Energy minimization}
The energy minimization is performed
with the input file ending with `\_mini.mdin',
e.g., 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/blob/main/gphmd_Amber/gphmd_snase_phrex/3BDC/3BDC_chainA_A_mini.mdin}{3BDC\_chainA\_A\_mini.mdin} for SNase.
During minimization, a harmonic restraint with the
force constant of 50 kcal mol$^{-1}$ \AA$^{-2}$ is placed
on all protein heavy atoms, except for the 
terminal and capping residues where all atoms are allowed to move. 
The minimization lasts 1000 steps, using the steepest descent and 
conjugate gradient algorithms for the first and second 500 steps, respectively. 
To run minimization, we call \textbf{pmemd.MPI} or \textbf{pmemd.cuda} if Amber 20 \cite{Case_Kollman_2021_Amber20} or a later version is used.
\begin{lstlisting}
$ export name='3BDC_chainA_A'
$ mpirun -n 4 
  $AMBERHOME/bin/pmemd.MPI -O 
  -i ${name}_mini.mdin -c ${name}.rst7 
  -p ${name}.parm7 -ref ${name}.rst7 
  -r ${name}_mini.rst7 -o ${name}_mini.out
\end{lstlisting}
% The latter requires at least an Nvidia GPU installed and configured properly as discusses above.
Here, '-n 4' specifies 4 CPU cores. Note, a large number of processors may be specified if hardware allows. 
The `-ref' option 
specifies the reference coordinate file
(3BDC\_ModelA\_A\_fixed.rst7).
The '-l' option specifies the initial coordinate file. It is omitted here, since it the same as the reference file.
If minimization runs error free, a timing info is given out
and two files are generated: 
the file 3BDC\_ModelA\_A\_fixed\_mini.out 
contains the printed energies, and 
*\_mini.rst7 is the restart coordinate file for the next step equilibration.

\paragraph{Equilibration}
% \begin{checklist}{Files needed for equilibration and production}
% %\textbf{}
% \begin{itemize}
% \item coordinates and topology files (\*.psf, \*.crd)
% %\item system CRD file 
% \item periodic boundary file (\*.pbc)
% \item  restart file from equilibration (\*.rst)
% \item production input prod\_hphrex.inp
% \item replica exchange file rep.cmd
% \end{itemize}
% \end{checklist}
Following minimization, we perform a four-stage equilibration
to relax the protein structure at physiological pH 7.5 (or the crystallization pH).
Note, for implicit-solvent simulations heating is not required. 
During the four-stage equilibration (2000 steps each),
the harmonic force constant on the heavy atoms is gradually reduced from 5.0, 2.0, 1.0, to 0.0 kcal/mol/\AA$^2$. 
CpHMD is turned on by setting the iphmd keyword to 1 
(GBNeck2-CpHMD)
in the Amber input file.

Using SNase as an example, the first equilibration stage is run with the following command:
\begin{lstlisting}
$ export name='3BDC_chainA_A'
$ $AMBERHOME/bin/pmemd.cuda -O
  -i ${name}_equil1.mdin -c ${name}_mini.rst7
  -p ${name}.parm7 -ref ${name}_mini.rst7
  -r ${name}_equil1.rst7 -o ${name}_equil1.out
  -x ${name}_equil1.nc -phmdin ${name}.phmdin
  -phmdparm gbneck2_input.parm
  -phmdout ${name}_equil1.lambda 
  -phmdrestrt ${name}_equil1.phmdrst
  $ sed -i 's/QPHMDStart = .true.,/QPHMDStart
        = .false.,/g' ${name}.phmdin
  $ sed -i 's/PHMDRST/PHMDSTRT/g'
        ${name}.phmdrst
\end{lstlisting}
Here, 3BDC\_chainA\_A\_mini.rst7 generated by the minimization step is used as the initial (-c) and reference (-ref) coordinates. 
The Amber job input file 3BDC\_chainA\_A\_equil1.mdin file specifies simulation length, 
restraints, and other MD control parameters.
3BDC\_chainA\_A.phmdin contains job control parameters
for GBNeck2-CpHMD.  
gbneck2\_input.parm, which is a common file
for GBNeck2-CpHMD simulations,
contains the model {\pka} values
and titration parameters.
3BDC\_chainA\_A\_equil1.rst7 is the restart file that contains coordinates and velocities.
3BDC\_chainA\_A\_equil1.phmdrst is the
corresponding CpHMD restart file that contains
the $\lambda$ values and velocities.  
3BDC\_chainA\_A\_equil1.out contains the energy information. 
3BDC\_chainA\_A\_equil1.lambda contains the $\lambda$ coordinates, although it is of little use since they are from the equilibration run. 
The two 'sed' commands modify the names 
for the next step.

For the rest of the equilibration stages, 
the same commands are used
but file names should be changed. For example, we should replace 'equil1' and 'mini1' with 'equil2' and 'mini2' for the second equilibration step, and use 3BDC\_chainA\_A\_equil1.rst7 as the initial and reference coordinate file. 
In addition, we need to use a new -phmdstrt flag and use 3BDC\_chainA\_A\_equil1.phmdrst as a CpHMD restart file. 
After all four equilibration stages are completed,
files with the names *equil4* are generated.

The above example uses pmemd.cuda, but we can also use pmemd.MPI if no GPU is available. The only modification is to replace
`\$AMBERHOME/bin/pmemd.cuda' with `mpirun -n 16 \$AMBERHOME/bin/pmemd.MPI' in the command, assuming 16 cores can be used for the simulation.

\paragraph{Production}
For production runs, we can use either independent pH, 
synchronous pH replica exchange, or asynchronous pH replica exchange.
%(\href{https://gitlab.com/shenlab-amber-cphmd/async_ph_replica_exchange}{Async-pHREX}). 
For independent pH runs, the simulation at different pH 
conditions are run independently either on the same or different machines. 
The running order is arbitrary, and we collect all the output files together for data analysis. 
The input files can be generated using 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/tree/master/cphmd_tools/cphmd_prep.sh}{cphmd\_prep.sh}
with \textbf{-run ind}.
For SNase, the production run 
is invoked with the following command:
\begin{lstlisting}
$ export name='3BDC_chainA_A'
$ $AMBERHOME/bin/pmemd.cuda -O
  -i ${name}_pH7.5_prod.mdin
  -c ${name}_equil4.rst7 -p ${name}.parm7
  -r ${name}_pH7.5_prod1.rst7
  -o ${name}_pH7.5_prod1.out
  -x ${name}_pH7.5_prod1.nc
  -phmdin ${name}.phmdin
  -phmdparm gbneck2_input.parm
  -phmdstrt ${name}_equil4.phmdrst
  -phmdout ${name}_pH7.5_prod1.lambda
  -phmdrestrt ${name}_pH7.5_prod1.phmdrst
  -inf pH7.5.mdinfo
\end{lstlisting}
The restart files generated by equilibration step 4 are used as starting files for the production run. No reference file is needed. 
Along with *.out, *.rst7, *.nc (trajectory), 
and *.phmdrst files, 
$\lambda$ coordinates are saved in 3BDC\_chainA\_A\_fixed\_pH*\_prod1.lambda. 
For the command above, we again use pmemd.cuda
but we can also run on CPUs with pmemd.MPI.
All input files for this example can be found in \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/gphmd_Amber/gphmd_snase_phind}{gphmd\_snase\_phind}.

pH replica-exchange simulations usually converge
much faster and thus require less simulation time than single pH simulations \cite{Wallace_Shen_2011_J.Chem.TheoryComput.,Harris_Shen_2019_J.Chem.Inf.Model.,Harris_Shen_2020_J.Chem.TheoryComput.}. 
The traditional replica-exchange algorithm
allows each replica to run in parallel and 
data exchange is synchronized.
The pH replica-exchange protocol is executed
similarly as the temperature replica-exchange protocol in Amber, 
i.e., we need a group file that contains the commands and inputs for every pH condition and then mpirun is called. 
The commands in the group file are quite similar to those for 
the independent pH runs. 
The input files can be generated using 
\textbf{cphmd\_prep.sh} with the option \textbf{-run rex}.
% For the Snase example, if we add '-run rex' to the command and change -pmax 12.0 to -pmax 12.5, 
% we will have a file named cphmd.groupfile in the 3BDC folder. 
For example,
\begin{lstlisting}
mpirun -np 64 
$AMBERHOME/bin/pmemd.MPI -ng 16 
-groupfile cphmd.groupfile
\end{lstlisting}
Note that we use 4 processes for a single pH replica and a total of 64 processes are needed (-np 64) given the number of pH conditions (-ng 16) in the Snase example command. 
For using mpirun for CpHMD, we should make sure that each pH replica runs on a single node so that there is only minimal data exchange between nodes if the whole job spreads across multiple node. Supposing we have a CPU cluster consisting of 64-core nodes, we should make sure the number of pH replicas is a factor of 64. 
This example uses pmemd.MPI to run on CPU clusters. 
If we have a GPU cluster, we can replace the '-np 64 
\$AMBERHOME/bin/pmemd.MPI -ng 16' with '-np 1 
\$AMBERHOME/bin/pmemd.cuda.MPI -ng 16'. However, if the number of GPUs is less than 16, it might not work properly.  
The input files can be found in the \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/gphmd_Amber/gphmd_snase_phrex}{gphmd\_snase\_phrex} folder.

The traditional replica-exchange algorithm was designed for CPU based HPC clusters, where a large number of processors are available such that all replicas can be run in parallel. 
However, the latter condition cannot be always met for today's GPU workstation or even HPC cluster, i.e., the number of GPUs is smaller than the number of replicas. Thus, we implemented an asynchronous pH replica-exchange algorithm, \cite{Henderson_Shen_2020_J.Chem.Phys.}
in which each replica is run consecutively and data exchange is performed after all replicas are run.
The users can download the related Python program \href{https://gitlab.com/shenlab-amber-cphmd/async_ph_replica_exchange}{async\_ph\_replica\_exchange} that allows an arbitrary number of pH replicas to be run on an arbitrary number of GPUs.

A Python script apHREX.py generated by 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/tree/master/cphmd_tools/cphmd_prep.sh}{cphmd\_prep.sh}
with `-run async' 
is called to run the asynchronous pH replica-exchange simulations for 3BDC. 
\begin{lstlisting}
$ export name='3BDC_chainA_A'
$ python apHREX.py 2000
  template_${name}_prod.mdin ${name}.parm7
  gbneck2_input.parm ${name}.phmdin
  ${name}_cphmd ${name}_equil4.rst7
  1.0 1.5 2.0 2.5 3.0 3.5 4.0 4.5
  5.0 5.5 6.0 6.5 7.0 7.5 8.0 8.5
  9.0 9.5 10.0 10.5 11.0 11.5 12.0
\end{lstlisting}
Output files, including the standard Amber outputs, lambda files, and trajectory files, are concatenated on the fly according to the pH conditions instead of replicas in the pHREX method. 
These files are stored in the 3BDC\_chainA\_A\_cphmd subdirectory.
The input files for this example can be found in \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/gphmd_Amber/gphmd_snase_aphrex}{gphmd\_snase\_aphrex}.
% Here we can describe a few different styles of production run with CpHMD. 
% this section can be broken down into three sections
%   1.) Single-pH runs with CpHMD Simulations in AMBER
%   2.) Async-pHREX with CpHMD simulations in AMBER

\paragraph{Other relevant settings}
For optimum performance in {\pka} prediction,
the default settings in GBNeck2-CpHMD should be followed.
Currently, the Amber ff14SB force field
is recommended for proteins.
A Langevin thermostat with the collision frequency of 1 ps$^{-1}$ is used to maintain the temperature for $\lambda$ dynamics. 
A non-bond cutoff of 999 {\AA} (i.e., no cutoff) is used. 
Files should be saved infrequently to avoid IO that substantially slows down simulations on the GPUs.
The default frequency of saving $\lambda$ files
every 1000 MD steps
is designed for simulations of 10 ns per replica. For longer simulations, less frequent saving should be used.
If analysis of conformational dynamics is desired, a saving frequency of every 10,000 MD steps (20 ps with a 2-fs time step) is recommended
for a simulation of 10 ns per replica. 
However, if only {\pka's} are desired, trajectory saving should be set to the frequency of restart file writing, e.g., 500,000 MD steps (1 ns with a 2-fs time step) or even larger depending on the wall clock time of the simulation (aiming at writing restart file once or twice a day). Writing to the Amber log file should be avoided or as infrequent as possible.

### 2022-01-23 6_analysis.tex
Initial Overleaf Import

\section{Post-simulation analysis}
CpHMD simulations allow the user to determine
the protonation states of all titratable sites
at the simulation pH conditions, 
the {\pka} values, 
or pH- or proton-coupled conformational dynamics. 
Below we will discuss these aspects.

\subsection{Determination of protonation states and {\pka} values}
The lambda file contains the time evolution of $\lambda$ (titration state) and $x$ (tautomeric state) coordinates of titratable groups.
These data allow us to determine the protonation
states at specific pH, calculate the
{\pka} values and tautomer states, and 
monitor convergence of protonation-state sampling.
As an example, Fig.~\ref{Fig:lambda_time} shows the time evolution of $\lambda$ for Cys481 in the BTK kinase
at different pH conditions.

 %--------------------------------------------- 
\begin{figure}[htb!]
    \centering
    \includegraphics[width=3.6in]{figs/btk_c481_lambdavstime.pdf}
    \caption{\textbf{Time series of the $\lambda$ value for Cys481 in the BTK kinase.} The data were saved every 20 ps (taken from Ref \cite{Liu_Shen_2021_J.Med.Chem.}). The pH conditions and fractions of mixed states ($0.2 \le \lambda \le 0.8$) are given. The plots show that the protonation-state sampling at all pH conditions converge after $\sim$10 ns.
    }
\label{Fig:lambda_time}
\end{figure}
 %---------------------------------------------

To facilitate analysis of the lambda files, we developed a Python tool \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-analysis}{cphmd\_anal.py}. 
This tool returns plots of the time series of unprotonated fractions and fitting of the unprotonated fractions to obtain {\pka's} (see discussion below).
For a trajectory at a specific pH, 
we can count the number of frames where a titratable residue $i$ is 
in the protonated or unprotonated state, which is defined 
as $\lambda_i < 0.2$ or $\lambda_i > 0.8$, respectively.
The unprotonated fraction of the residue ($S_i$)
is calculated as
\begin{equation}
S_i = \frac{N_i^{\rm Unprot}}{N_i^{\rm Prot} + N_i^{\rm Unprot}},
\end{equation}
where $N_i^{\rm Prot}$ and $N_i^{\rm Unprot}$ are
the number of protonated and unprotonated frames,
respectively.
Plotting the time series of the $S_i$ value informs
convergence of the protonation state sampling (Fig.~\ref{fig:intro}, step 5).
Once convergence is reached, the {\pka} of residue $i$
can be obtained by fitting the $S_i$ 
values at all simulation pH to 
the generalized Henderson-Hasselbalch equation,
\begin{equation}
S_i = \frac{1}{1+10^{n(\rm pK_{a,i} - pH)}},
\end{equation}
where $n$ is the Hill coefficient.
The resulting best fit is referred to as the titration curve (Fig.~\ref{fig:intro}, step 5).
A significant deviation of $n$ from 1 indicates cooperativity
($n>1$)
or anti-cooperativity ($n<1$) with a neighboring residue.
We note, in the past, $\lambda_i < 0.1$ and $\lambda_i > 0.9$ were used for
defining the protonated and unprotonated states \cite{Khandogin_Brooks_2005_Biophys.J.,Khandogin_Brooks_2006_Biochemistry}; however, our extensive studies (based on more than 100 proteins) showed that the calculated {\pka} value is not sensitive to the cutoff.

\subsection{Analysis of proton-coupled conformational dynamics and rationalization of the calculated {\pka} values}
A major application of 
CpHMD simulations is 
elucidation of pH-dependent or proton-coupled conformational dynamics, which 
in turn rationalizes the calculated {\pka} values. 
This practice can be best explained using an example. 
The pH replica-exchange GBNeck2-CpHMD simulations 
of the BTK kinase
\cite{Liu_Shen_2021_J.Med.Chem.}
gave a {\pka} of about 7.5 for Cys481,
which is one unit lower than the 
model {\pka} of cysteine -- the {\pka} value of an isolated cysteine fully exposed to solvent.
Analysis of the pH replicas (i.e., trajectories at different pH conditions) revealed that Cys481 accepts hydrogen bonds (h-bond) from 
Asn481 and Thr410 when it is in the deprotonated 
thiolate state (Fig.~\ref{Fig:analysis}a).
Note, these h-bonds are absent in the crystal structure (PDB 3pj3).
Following this observation, we plotted the deprotonated fractions of Cys481 at different pH conditions (Fig.~\ref{Fig:analysis}b)
and the occupancies of the h-bond formation 
between Cys481 and Asn484 or Thr410
(Fig.~\ref{Fig:analysis}c).
A comparison between the pH-dependent deprotonation
and h-bond formation demonstrates that the two are correlated, i.e., deprotonation of Cys481 is coupled to the h-bond formation.
It also suggests that the {\pka} downshift of Cys481 relative to the model value
can be attributed to the stabilization of the deprotonated thiolate state by the h-bond formation with Asn484 and Thr410.
We note, while implicit-solvent based CpHMD simulations have been successfully applied to {\pka} predictions and rationalization, hybrid-solvent and all-atom CpHMD simulations offer more accurate description of conformational dynamics.
We refer the user to the studies listed in Table~\ref{Table:applications} as examples.

 %--------------------------------------------- 
\begin{figure}[htb!]
    \centering
    \includegraphics[width=3.3in]{figs/analysis_example.pdf}
    \caption{\textbf{An example analysis of proton titration of a cysteine and coupling with conformational dynamics.}
    (a) A zoomed-in view of the structural environment of Cys481 in the BTK kinase taken from the trajectory at pH 9. 
    The h-bonds with Asn484$\cdots$Cys481 and Thr410$\cdots$Cys481 are shown. 
    The pH 9 replica is used. 
    (b) The unprotonated fractions of Cys481 at different pH conditions. The titration curve represents the best fit to the Henderson-Hasselbalch equation. 
    (c) Occupancy of the h-bond formation between Cys481 thiolate and Asn484 or Thr410
    at different pH conditions. 
  Reprinted with permission from Liu, Zhan et al.\cite{Liu_Shen_2021_J.Med.Chem.} 
    Copyright {2021} American Chemical Society.
    }
\label{Fig:analysis}
\end{figure}
 %---------------------------------------------

### 2022-01-23 7_parameterization.tex
Initial Overleaf Import

\section{Parameterization of model titration} 
% - Jack/Ruibin
CpHMD is a relative free energy simulation approach, whereby the potential of mean force (PMF) of deprotonation of
a titratable amino acid in the protein environment is calculated relative to that of a reference, i.e., a model compound or peptide in solution
\cite{Lee_Brooks_2004_Proteins,Khandogin_Brooks_2005_Biophys.J.}. 
Model compounds, which were used in the early development of CpHMD methods
\cite{Lee_Brooks_2004_Proteins,Khandogin_Brooks_2005_Biophys.J.,Huang_Shen_2016_J.Chem.TheoryComput.},
are blocked single amino acid residues \ce{CH3CO}-X-\ce{NH2} or \ce{CH3CO}-X-\ce{CONH2},
where X represents a titratable amino acid.
In the development of the hybrid-solvent CpHMD \cite{Wallace_Shen_2011_J.Chem.TheoryComput.} and GBNeck2-CpHMD \cite{Huang_Shen_2018_J.Chem.Inf.Model.,Harris_Shen_2019_J.Chem.Inf.Model.}, 
model peptides 
\ce{CH3CO}-AAXAA-\ce{NH2} were used to take advantage of recent experimental data \cite{Thurlkill_Pace_2006_ProteinSci.,Platzer_McIntosh_2014_J.Biomol.NMR}.
Since in the CpHMD simulation, a model PMF is subtracted,
simulation of a model compound or peptide should return a zero PMF at the pH value equal to the model {\pka}.
In other words, protonated and deprotonated states are sampled with equal probabilities ($S\approx0.5$).
To ensure this is the case, the model PMF needs to be accurately determined.

To obtain the model PMF $U^{\rm mod}$, we use a free energy 
simulation method called thermodynamics integration,
in which the mean forces at different values of $\theta$ (for single site titration)
or $\theta$ and $x$ for double site titration
% $\langle\partial U/\partial\theta|_{\theta_i}\rangle$,
are calculated and then analytically
integrated to obtain $U^{\rm mod}$.
The single site titration model is applied
to Cys and Lys. The latter has equivalent protons, so one is selected for titration.
There are two types of double site titration model. His sidechains have two titratable nitrogens with different microscopic {\pka's}, while
carboxylate sidechains (Asp and Glu) have two titratable oxygens but identical microscopic {\pka's}.
These three titration models require
different forms of $U^{\rm mod}$ and methods of fitting.

\subsection{Protocol of parameterization for single site model titration}
For running TI simulations to obtain parameters, we set the options \textbf{prlam} to FALSE, \textbf{prderiv} to TRUE, and \textbf{phtest} to TRUE in the phmdin file, vph{\_}theta to 0 in the phmdstrt file.
For a single site titration model (e.g., Cys and Lys), 
we run CpHMD simulations at
$\theta_i=\left(0.4,0.6,0.785,1.0,1.2,1.4\right)$.
Each simulation is run for 10 ns, from which the mean forces
$\langle\partial U/\partial\theta|_{\theta_i}\rangle$
are calculated. 
Fitting of the mean forces  
to the following analytic form of $\partial U/\partial\theta$ (Eq.~\ref{Eq:dU_dtheta}) returns the two parameters $A$ and $B$ (Fig.~\ref{Fig:parm_cys}).

\begin{equation}
    \frac{\partial U}{\partial\theta}=2A\sin\left(2\theta\right)\left(\sin^{2}\theta-B\right).
    \label{Eq:dU_dtheta}
\end{equation}
As an example, Fig.~\ref{Fig:parm_cys} shows the fitting result for the GBNeck2-CpHMD titration of  Cys model peptide.
We note, fitting should be performed in the $\theta$ space (and not $\lambda$ space) to reduce errors.
 %--------------------------------------------- 
\begin{figure}[htb!]
    \centering
    \includegraphics[width=3.3in]{figs/cys_parm.pdf}
    \caption{\textbf{Example of fitting A and B parameters for a single site titration model.}
    }
\label{Fig:parm_cys}
\end{figure}
 %---------------------------------------------

\paragraph{Protocol of parameterization for histidine model titration}
For a double site titration model 
with two different microscopic \pka's
(e.g. His), the two-dimensional PMF (Eq.~\ref{eq:double}) can be reduced to the
following form
\cite{Khandogin_Brooks_2005_Biophys.J.},

\begin{align}
U^{\rm mod}= & A_{10}\lambda^{2}x^{2}+2\left(A_{1}B_{1}-A_{0}B_{0}\right)\lambda x \nonumber \\ & + 2\left(A_{0}B_{0}-A_{1}B_{1}-A_{10}B_{10}\right)\lambda^{2}x \nonumber \\ & + A_{1}\lambda^{2}-2A_{1}B_{1}\lambda,
\label{eq:PMF_his}
\end{align}
where $\lambda=sin^2\theta$ and $x=sin^2\theta^x$. 
The six parameters $A_{0}$, $B_{0}$, $A_{1}$,  $B_{1}$, $A_{10}$, and $B_{10}$ are the parameters
in the quadratic functions
that describe the one-dimensional processes
\cite{Khandogin_Brooks_2005_Biophys.J.}.
$A_0$ and $B_0$ correspond to \ce{HIP <=> HID},
i.e., titration at N$\epsilon$;
$A_1$ and $B_1$ correspond to \ce{HIP <=> HIE},
i.e., titration at N$\delta$;
and $A_{10}$ and $B_{10}$ correspond to the tautomer interconversion \ce{HIE <=> HID}.

\textbf{Step 1.}
To obtain $A_0$ and $B_0$, we 
run TI simulations by fixing
$\theta^{x}$ at 0 ($\lambda=0$, HID, N$\delta$ is protonated), varying
$\theta$ at $\left(0.0,0.2,0.4,0.6,\\
0.785,1.0,1.2,1.4,1.5708\right)$, and fitting
the resulting mean forces to the following one-dimensional function,
\begin{equation}
    \frac{\partial U}{\partial\theta}=2A_{0}\sin\left(2\theta\right)\left(\sin^{2}\theta)-B_{0}\right).
\end{equation}

\textbf{Step 2.}
To obtain $A_1$ and $B_1$,
we run TI simulations by fixing
$\theta^{x}$ at 1.5708 ($\lambda=1$, HIE, N$\delta$ is protonated), varying
$\theta$ at $\left(0.0,0.2,0.4,0.6,\\
0.785,1.0,1.2,1.4,1.5708\right)$, and 
fitting
 to the following one-dimensional function,
\begin{equation}
    \frac{\partial U}{\partial\theta}=2A_{1}\sin\left(2\theta\right)\left(\sin^{2}\theta-B_{1}\right).
\end{equation}

\textbf{Step 3.}
To obtain $A_{10}$ and $B_{10}$, we run TI simulations by fixing $\theta$ at 1.5708
($\lambda=0$, HIP, doubly protonated His),
varying $\theta_{x}$ at $\left(0.0,0.2,0.4,0.6,0.785,1.0,1.2,1.4,1.5708\right)$ 
and fitting the resulting mean forces
 to the following one-dimensional function,

\begin{equation}
    \frac{\partial U}{\partial\theta^{x}}=2A_{10}\sin\left(2\theta^{x}\right)\left(\sin^{2}\theta^{x}-B_{10}\right).
\end{equation}

\paragraph{Protocol of parameterization for carboxylic acid model titration}
For double site titration models with two identical microscopic \pka's (e.g., Asp and Glu), the two-dimensional PMF (Eq.~\ref{eq:double})
can be reduced to the following form \cite{Khandogin_Brooks_2005_Biophys.J.},
% Equation...
\begin{equation}
\begin{split}
U^{\rm mod}(\lambda_i,x_i) =
&(R_1 \lambda_i^2 + R_2 \lambda_i + R_3)(x_i + R_4)^2 + R_5 \lambda_i^2 + R_6 \lambda_i
\end{split}\label{eq:carboxyl}
\end{equation}

where $R_1$, ..., $R_6$ are parameters that can be determined via one-dimensional fitting.

%   Determine R1 R2 and R3 by fitting A(lambda) to R1 lambda^2 + R2 lambda + R3
%     R4 = 0.5
%     Determine R5 by fitting A(x) to C1 x^2 + C2 x + R5
%     Determine R6 by fitting B(x) to C1 x^2 + C2 x + R6

\textbf{Step 1.} To generate the data for fitting,
we run TI simulations at the combinations of $\theta$ value of 0.0, 0.4, 0.6, 0.785, 1.0, 1.2, or 1.4
and $\theta^{x}$ value of 0.0, 0.4, 0.6, 0.785, 1.0, 1.2, 1.4, or 1.5708.

\textbf{Step 2.}
We obtain $A$ and $B$ parameters 
at each value of $\theta$, 
$A(\theta)$ and $B(\theta)$,
by fitting
$\left<\partial U/\partial\theta_{x}\right>$
at different $\theta^x$ values to
the derivative of the quadratic function,
% can be obtained at each value of $\theta$ by fitting $<\partial U/\partial\theta_{x}>$ to
\begin{equation}
    \frac{\partial U}{\partial\theta^{x}}=2A(\theta)\sin\left(2\theta_{x}\right)\left(\sin^{2}\theta_{x}-B(\theta)\right),
\end{equation}
where $A(\theta)$ and $B(\theta)$ are the $\theta$-dependent parameters.

\textbf{Step 3.}
We obtain $A$ and $B$ parameters 
at each value of $\theta^x$, $A(\theta^x)$ and $B(\theta^x)$,
by fitting
$\left<\partial U/\partial\theta\right>$
at different $\theta$ values,
% compute $A\left(\theta_{x}\right)$ and $B\left(\theta_{x}\right)$ at each value of $\theta_{x}$ by fitting $\partial U/\partial\theta_{x}$ to

\begin{equation}
    \frac{\partial U}{\partial\theta}=2A(\theta^x)\sin\left(2\theta\right)\left(\sin^{2}\theta-B(\theta^x)\right),
\end{equation}
where $A(\theta^x)$ and $B(\theta^x)$ are the parameters.

\textbf{Step 4.}
We obtain $R_{1}$, $R_{2}$, and $R_{3}$ by fitting $A(\theta_i)$ to

\begin{equation}
    A\left(\theta\right) = R_{1}\sin^{4}\theta + R_{2}\sin^{2}\theta+R_{3},
\end{equation}

\textbf{Step 5.}
We set $R_{4}=0.5$, because the two titrating sites are identical. We then obtain $R_{5}$ by fitting $A(\theta^x)$ to

\begin{equation}
    A\left(\theta^{x}\right)=a_{0}\sin^{4}\theta^{x}+a_{1}\sin^{2}\theta^{x}+R_{5},
\end{equation}
and obtain $R_{6}$ by fitting $B(\theta^x)$ to

\begin{equation}
    B\left(\theta^{x}\right)=a_{0}\sin^{4}\theta^{x}+a_{1}\sin^{2}\theta^{x}+R_{6},
\end{equation}

\subsection{Scripts for TI simulations and parameterization}
To facilitate TI simulations and parameter fitting, we implemented a Shell script 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/blob/main/parameterization/scripts/heat_equil_prod_param.sh}{heat\_equil\_prod\_param.sh} and a Python program 
\href{
https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/blob/main/parameterization/scripts/cphmd_parm_fit.py}{cphmd\_parm\_fit.py}.
% The above parameterization settings are included in the heat\_equil\_prod\_param.sh script and the fitting procedure is included in the cphmd\_parm\_fit.py script, both in the parameterization/scripts folder of the \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/parameterization/scripts}{cphmd-tutorial} gitlab repository. 
Here we use Cys to illustrate the usage.
% A general instruction to use them (here we use Cys as an example) is firstly to run

\begin{lstlisting}
$ ./min.scr cys
$ ./heat_equil_prod_param.sh cys
\end{lstlisting}

After minimization, heating, equilibration, and production thermodynamic integration, 
we obtain $\partial U/\partial\theta$ values saved in files with the extension .lambda as in the folder \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/parameterization/single}{single} (single site titration). 
We then call the Python script to fit the A and B parameters.

\begin{lstlisting}
$ python cphmd_parm_fit.py single
\end{lstlisting}

Four files are generated: du.data, du.fit, du.png, and cphmd.parm. du.data contains the mean force data, du.fit contains the detailed fitting results, du.png is the plot of the fitting (Fig.~\ref{Fig:parm_cys}) , and cphmd.parm contains the A and B parameters that can be directly copied and pasted into the standard CpHMD parameter file. 
The instructions for parameterization of different models are described in the folder \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/parameterization}{parameterization} and the example results are included, as in the folders 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/parameterization/single}{single},
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/parameterization/his}{his},
and 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/parameterization/carboxyl}
{carboxyl}. 
% Here is a single-column checklist that consists of multiple sub-checklists
% \begin{Checklists}

% \begin{checklist}{A list}
% \textbf{Single-column checklists are also straightforward by removing the asterisk}
% \begin{itemize}
% \item First thing let's do an item which breaks across lines to see how that looks
% \item Also remember
% \item And finally
% \end{itemize}
% \end{checklist}

% \begin{checklist}{Another list}
% \textbf{This is some further description.}
% \begin{itemize}
% \item First thing
% \item Also remember
% \item And finally
% \end{itemize}
% \end{checklist}

% \end{Checklists}

### 2022-01-23 Draft/cphmd_tutorial_draft.tex
Initial Overleaf Import

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% LIVECOMS ARTICLE TEMPLATE FOR BEST PRACTICES GUIDE
%%% ADAPTED FROM ELIFE ARTICLE TEMPLATE (8/10/2017)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PREAMBLE
\documentclass[9pt,tutorial]{livecoms}
% Use the 'onehalfspacing' option for 1.5 line spacing
% Use the 'doublespacing' option for 2.0 line spacing
% Use the 'lineno' option for adding line numbers.
% Use the "ASAPversion' option following article acceptance to add the DOI and relevant dates to the document footer.
% Use the 'pubversion' option for adding the citation and publication information to the document footer, when the LiveCoMS issue is finalized.
% The 'bestpractices' option for indicates that this is a best practices guide.
% Omit the bestpractices option to remove the marking as a LiveCoMS paper.
% Please note that these options may affect formatting.

\usepackage{lipsum} % Required to insert dummy text
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\usepackage{hyperref} % Not original - Added by Jack
\usepackage{listings} % Not original - Added by Jack
\DeclareSIUnit\Molar{M}
\usepackage[italic]{mathastext}
\newcommand{\pka}{p\textit{K}$_{\rm a}$}
\graphicspath{{figures/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% IMPORTANT USER CONFIGURATION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\versionnumber}{1.0}  
% you should update the minor version number in preprints and major version number of submissions.
\newcommand{\githubrepository}{\url{https://github.com/myaccount/homegithubrepository}}  %this should be the main github repository for this article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{A Guide to Continuous Constant pH Molecular Dynamics in AMBER and CHARMM v[\versionnumber]}

\author[1]{Jack A. Henderson}
\author[1]{Ruibin Liu}
\author[1,\authfn{2}]{Julie A. Harris}
\author[1,\authfn{3}]{Yandong Huang}
\author[1]{Vinicius M. de Oliveira}
\author[1*]{Jana Shen}
\affil[1]{University of Maryland School of Pharmacy, Baltimore, MD}
%\affil[2]{Institution 2}

\corr{jana.shen@rx.umaryland.edu}{JS}  % Correspondence emails.  FMS and FS are the appropriate authors initials.
%\corr{email2@example.com}{FS}

\orcid{Jack A. Henderson}{0000-0001-6675-7944}
\orcid{Ruibin Liu}{0000-0001-8395-9353}
\orcid{Julie A. Harris}{0000-0002-1130-6457}
\orcid{Yandong Huang}{0000-0002-1452-6383}
\orcid{Vinicius M. de Oliveira}{0000-0003-0927-3825}
\orcid{Jana Shen}{0000-0002-3234-0769}

% \contrib[\authfn{1}]{These authors contributed equally to this work}
% \contrib[\authfn{2}]{These authors also contributed equally to this work}

\presentadd[\authfn{2}]{Mapware, Washington DC}
\presentadd[\authfn{3}]{Jimei University, College of Computer Engineering, Xiamen, China}
% \presentadd[\authfn{4}]{Department, Institute, Country}

%\blurb{This LiveCoMS document is maintained online on GitHub at \githubrepository; to provide feedback, suggestions, or help improve it, please visit the GitHub repository and participate via the issue tracker.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PUBLICATION INFORMATION
%%% Fill out these parameters when available
%%% These are used when the "pubversion" option is invoked
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pubDOI{10.XXXX/YYYYYYY}
\pubvolume{<volume>}
\pubissue{<issue>}
\pubyear{<year>}
\articlenum{<number>}
\datereceived{Day Month Year}
\dateaccepted{Day Month Year}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}
\maketitle

\begin{abstract}
%This particular document provides a skeleton illustrating key sections for a Tutorial document.
%Please see the sample \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates} for additional information on and examples of using the LiveCoMS LaTeX class.
%Here we also assume familiarity with LaTeX and knowledge of how to include figures, tables, etc.; if you want examples, see the sample just referenced.

%In your work, in this particular slot, please provide an abstract of no more than 250 words.
%Your abstract should explain the main contributions of your article, and should not contain any material that is not included in the main text.
%Please note that your abstract, plus the authorship material following it, must not extend beyond the title page or modifications to the LaTeX class will likely be needed.

% Begin Abstract Here

\end{abstract}

\end{frontmatter}

\section{Introduction} % - Vini

Variations in the pH condition of biological systems can lead to drastic changes in the physicochemical processes in these systems. 
Protein folding is one example of a natural pH-dependent process, where the unfolded protein only can reach its native state in a narrow pH range \cite{Pelegrine_Gasparetto_2005_LWT-FoodScienceandTechnology}. 
Thereby, small pH changes variation in the cell can lead to misfolding diseases and even cancer proliferation and metastasis \cite{vanderKamp_Daggett_2010_BiophysicalJournal,Webb_Barber_2011_NatRevCancer,Zheng_Liu_2020_TheInternationalJournalofBiochemistry&CellBiology}. 
The protonation state of residues is also critical in drug design once the ligand-protein binding is susceptible to charge changes \cite{Dominguez_Sussman_2010_Biochemistry,Bax_Edge_2017_ActaCrystallogrDStructBiol}. 
The SARS-CoV main protease (Mpro) is an example of a protein that possesses biological activity in a narrow pH range around 7.0 \cite{Tan_Hilgenfeld_2005_JournalofMolecularBiology}. Out of this range, its active site can collapse by protonation changes of residues that compose it \cite{Verma_Shen_2020_J.Am.Chem.Soc.}. More than that, the rational design of inhibitors must consider how the ligand interactions can alter the protonation state of the titratable residues into the binding site. The solution pH is also critical for the study, development, and fabrication of functional dynamic materials. An example of this is the biopolymer of chitosan, in which its physicochemical behavior is completely different at low and high pH conditions \cite{Morrow_Shen_2015_J.Am.Chem.Soc.}.

Classical molecular dynamics (MD) usually approach these pH-dependent systems by fixing the protonation states of titratable residues accordingly with experimental pKa values or pKa theoretical predictions. Several computational tools are currently used to predict pKa values; among the most used methods, we can cite the empirical ones PropKa and the Poisson-Boltzmann \cite{Olsson_Jensen_2011_J.Chem.TheoryComput.,Warwicker_Warwicker_2011_ProteinsStruct.Funct.Bioinforma.a}, and solver-based methods APBS-PDB 2PQR, DelPhiPKa, and H++ \cite{Unni_Baker_2011_J.Comput.Chem.,Wang_Alexov_2016_Bioinformatics,Anandakrishnan_Onufriev_2012_NucleicAcidsResearch}. Such tools can estimate the protonation states before running MD simulations, but they do not directly consider the dynamic flexibility of the molecules, which may lead to incorrect pKa predictions. Also, even assuming an accurate protonation states assignment at the beginning of MD, the conformational changes along the simulation can modify the electrostatic environment around titratable residues, affecting their protonation state. Thus, this relationship between conformational dynamics and the protonation states is not considered in the classical MD, and addressing this relationship is one of the most advantages of constant pH molecular dynamics (CpHMD) over conventional MD.

Over the past decades, much effort has been applied to developing and improving constant pH methods \cite{Chen_Shen_2014_Mol.Simulat.}. We can divide the most applied methods into discrete and continuous constant pH. The former periodically interrupts MD and uses Metropolis Monte Carlo (MC) criterion to sample the protonation states during simulation \cite{Burgi_Gunsteren_2002_ProteinsStruct.Funct.Bioinforma.}. In this case, the protonation of titratable groups is settled discretely, assuming one of the two states (protonated or deprotonated). On the other hand, the continuous CpHMD treats the protonation states of the ionizable sites with an auxiliary set of continuous titration coordinates, using an extended Hamiltonian $\lambda$-dynamics approach \cite{Chen_Shen_2014_Mol.Simulat.}. This method allows the system to escape local energy minima by accessing partially protonated states temporarily. For years, both MC and continuous CpHMD simulations have been performed and validated by pKa calculations for several systems in different environments \cite{Chen_Shen_2014_Mol.Simulat.,Radak_Roux_2017_J.Chem.TheoryComput.}. Also, CpHMD has been crucial to investigate the molecular mechanisms involved in the pH dependency of protein folding \cite{MartinsdeOliveira_PereiraLeite_2018_BiophysicalJournal,Yue_Shen_2018_Phys.Chem.Chem.Phys.}, enzyme catalysis \cite{Huang_Shen_2018_J.Phys.Chem.Lett.}, conformational changes \cite{Sarkar_Roitberg_2020_J.Phys.Chem.B}, protein/ligand binding \cite{Ellis_Shen_2016_J.Phys.Chem.Lett.,Harris_Shen_2017_J.Phys.Chem.Lett.}, activation of channels and transporters \cite{Huang_Shen_2020_}, and viral uncoating \cite{Chen_Shen_2016_J.Phys.Chem.Lett.}.

Our research group has exhaustively explored continuous CpHMD for different solvent models and force fields (FF). Using the GBNeck2 implicit solvent model implemented in Amber and hybrid-solvent based on the CHARMM platform, we accurately predicted pKa values for a large set of proteins \cite{Huang_Shen_2018_J.Chem.Inf.Model.,Harris_Shen_2019_J.Chem.Inf.Model.}. We also showed that these methods are reliable to predict cysteine and lysine protonation states, which are frequently used in covalent drug design \cite{Liu_Shen_2019_J.Am.Chem.Soc.,Harris_Shen_2020_J.Chem.TheoryComput.}. The active sites of SARS and MERS coronaviruses proteases were also investigated applying CpHMD, and we could clarify how proton-coupled conformational changes can alter their activity \cite{Verma_Shen_2020_J.Am.Chem.Soc.,Henderson_Shen_2020_J.Chem.Phys.}. Performing CpHMD simulations for Human Renin, we detailed its pH-dependent structure-dynamic-function relationship \cite{Ma_Shen_2020_J.Chem.Inf.Model.}, and explored the molecular mechanisms involved in inhibitors' binding and selectivity in BACE1 and $\beta$-Secretase \cite{Ellis_Shen_2015_J.Am.Chem.Soc.,Ellis_Shen_2016_J.Phys.Chem.Lett.,Harris_Shen_2017_J.Phys.Chem.Lett.}. Such examples are only a small sample of the vast application possibilities that CpHMD covers.

\begin{table*}[]
\begin{center}
\begin{tabular}{lp{0.5\linewidth}c}
System &  Topic  & CpHMD Method \\ 
\hline
% Add H. Line (Soluble)
BACE1   & Investigated pH-dependent ligand binding site dynamics, small molecule inhibitor interactions, and identify proton coupled allostery  & Hybrid-Solvent \\
BACE2   & Investigated how protonation state changes of a small molecule inhibitor affect binding with comparisons to BACE1 & Hybrid-Solvent \\
Cat-D   & Investigated pH-dependent ligand binding site dynamics and compared them to BACE1 & Hybrid-Solvent \\
Plasmepsin & Dissect the pH-dependent interaction of the ligand binding site, identify proton coupled allostery, elucidate the catalytic roles of the active site & Hybrid-Solvent \\
Renin   & Explore the acid/base roles of the catalytic dyad and conformational dynamics of the ligand binding site & Hybrid-Solvent \\
Silk Protein & & \\
SNase   &    & \\
Kinases &  Discovery Reactive Lys  & Implicit-Solvent \\
PLpro   &  Identify reactive Cys residues and elucidate proton-coupled conformational dynamics  & Implicit-Solvent \\
Mpro    & Explore protonation state dependent substrate binding site dynamics and identify reactive Cys resdiues & Implicit-Solvent \\
% Add H. Line (Subsection: Folding)
BBL     & & \\
NTL9    & & \\
% Add H. Line (Transmemb)
NhaA    & Addressed a controversy surrounding the proton binding residues involved in the sodium proton exchange mechanism & Membrane-Enabled Hybrid-Solvent \\
NapA    & Dissected the competitive ion binding mechanism and identified the proton binding sites. & Membrane-Enabled Hybrid-Solvent \\
AcrB    & & Membrane-Enabled Hybrid-Solvent \\
M2      & & Membrane-Enabled Hybrid-Solvent \\
\end{tabular}
\end{center}
\caption{Protein systems that CpHMD has been used on.}
\begin{minipage}{\columnwidth}
Possible Minipage
\end{minipage}
\label{Fixed-Charge_simulations}
\end{table*}

%Here you would explain what problem you are tackling and briefly motivate your work.

%In this particular template, we have removed most of the usage examples which occur in \texttt{sample-document.tex} to provide a minimal template you can modify; however, we retain a couple of examples illustrating more unusual features of our templates/article class, such as the checklists, and information on algorithms and pseudocode.

%Keep in mind, as you prepare your manuscript, that you should plan for a representative image  which will be used to highlight your article on the journal website and publications. Usually, this would be one of your figures, but it must also be uploaded separately upon article submission. We give specific guidelines for this image on the journal website in the section on article submission (see \url{https://livecomsjournal.github.io/authors/policies/index.html#article-submission}).

%Additionally, for well-formatted manuscripts, we recommend that you let LaTeX handle figure/table placement for you as much as possible, so please avoid specifying strenuous float instructions like `[h!]` and `[H]` as much as possible.

% This section gives the background and idea/purpose of CpHMD
% Importance of pH in simulations OK
% How Traditional MD Handles Titratable Sites OK
% How are titratable sites treated without pHMD (PropKA, Delphi,  H++, ect...) OK
% Idea of pHMD OK
% Brief description of types of pHMD (DpHMD and CpHMD) OK
% Advantages of CpHMD OK
% Examples of CpHMD in action should range from CHARMM Hybrid-Solvent CpHMD to Amber CpHMD simulations probing for reactive Cys and Lys targets.

% Figure Concept: A figure showing several different protein systems that we have investigated with CpHMD.

% Topics to Cover (Might Need it's own section):
%   1.) Brief Description of how titratable sites are handle, i.e. lambda dynamics and tautomers
%   2.) Give an overview of the extended Hamiltonian.
%   3.) Discuss the importance of pH-based replica exchange and its applications on CPUs and GPUs

% Figure Concept: Here would be a good place to show the classic MD protocol with the CpHMD portion and we could show a small example of the replica exchange process.

\subsection{Constant pH molecular dynamics}

In continuous CpHMD, we define a titration coordinate, $\lambda_i$, where $i$ represents the ionizable residues \cite{Huang_Shen_2018_J.Chem.Inf.Model.}. We attribute to this coordinate an implicit function of an underlying coordinate to assure $\lambda_i$ remains in a range between 0 and 1. Thus, the $\lambda_i$ expression is given by:

\begin{equation}
    \lambda_i = \sin^2(\theta_i).
\end{equation}

When $\lambda_i = 0$, the $i$th titratable group is fully protonated, and it is fully deprotonated when $\lambda_i = 1$. In simulation analysis, we usually apply cutoffs for protonated states ($\lambda^P < 0.1$), and for deprotonated states ($\lambda^U > 0.9$) \cite{}. After defined the titration coordinate, we can use an extended Hamiltonian to allow the simultaneous propagation of spatial (real) and titration (virtual) coordinates:

\begin{equation}
\begin{split}
\label{eq:Hamiltonian}
\mathcal{H}(\{r_a\},\{\theta_i\})=
&\sum_a\frac{1}{2}m_a\dot{r}_a^2+\sum_i\frac{1}{2}m_i\dot{\theta}_i^2+U^{\rm int}(\{r_a\})\\
&+U^{\rm hybr}(\{r_a\},\{\theta_i\})+U^{\ast}(\{\theta_i\}),
\end{split}
\end{equation}
where $a$ and $i$ represent the index for the spatial and titration-site coordinates, respectively. The two first terms of eq. \ref{eq:Hamiltonian} describe the kinetic energy for the atoms and the titration site. $U^{\rm int}$ is the titration-independent bond and non-bond energies from the non-titrating atoms. The fourth term, $U^{\rm hybr}$, depends on the solvent treatment. Herein, we will present the fully implicit solvent and the hybrid-solvent model, which $U^{\rm hybr}$ for both models includes Coulomb, vdW, and generalized Born (GB) energies and is written as

\begin{equation}
\label{eq:Uhybr}
U^{\rm hybr}(\{r_a\},\{\theta_i\})=U^{\rm elec}(\{r_a\},\{\theta_i\})+
U^{\rm vdW}(\{r_a\},\{\theta_i\})+U^{\rm GB}(\{r_a\},\{\theta_i\}).
\end{equation}

In eq. \ref{eq:Uhybr}, for the calculation of the Coulomb and GB electrostatic energies, we incorporate the atomic partial charges on the titrating residue and compute the linear interpolation between the values in the deprotonated and protonated states. Thus, we can write the charge on atom $j$, $q_j$, on the titrating residue $i$ as
%
\begin{equation}
q_j(\lambda) = (1-\lambda_i) q_j^{prot} +  \lambda_i q_j^{unprot}.
\label{eq:qlamb}
\end{equation}

The second term of eq. \ref{eq:Uhybr} represents the linear attenuation of the van der Waals (vdW) interactions between the titrating residue $i$ and its surrounding protein region:

\begin{equation}
U^{vdW} = (1-\lambda_i)U^{vdW}(i, p),
\end{equation}
where $p$ is the index of all other atom types, and $U^{vdW}$ is the protonation independent vdW interaction energy (Lennard-Jones 6-12 potential).

Finally, the energy component that only affects the titratable groups is given by the last term of eq. \ref{eq:Hamiltonian} and consists of three biasing potentials, 
\begin{equation}
\label{eq:biasingpotential}
U^{\ast}(\{\theta_i\})=-U^{model}(\lambda_i)+U^{barr}(\lambda_i)+U^{pH}(\lambda_i),
\end{equation}

The first term of $U^{\ast}$ describes a potential of mean force (PMF) along the $\lambda$-coordinate, considering $\lambda_i$ as an isolated model compound in the solvent,

\begin{equation}
\label{eq:Umod_1d}
U^{model}=A(\lambda_i-B)^2,
\end{equation}
where $A$ and $B$ are parameters that can be determined via thermodynamic integration (TI), which are performed by MD simulations fixing several $\lambda(\theta_i)$ values, and the parameters are obtained fitting the derivative of the mean force. The PMF is second-order in $\lambda$ and $x$ for residues with two tautomeric states, where $x$ describes the tautomeric state. In this case, the two-dimensional PMF can be written as a bivariate polynomial,
%
\begin{equation}
\begin{split}
U^{mod}(\lambda_i,x_i) =
&a_0 \lambda_i^2 x_i^2 + a_1 \lambda_i^2 x_i + a_2 \lambda_i x_i^2 + a_3 \lambda_i^2 \\
&+ a_4 x_i^2 + a_5 \lambda_i x_i + a_6 \lambda_i + a_7 x_i + a_8,
\end{split}\label{eq:double}
\end{equation}
where $x_{i}=\sin^{2}\left(\theta^{x}_{i}\right)$ where the $\theta^{x}_{i}$ are unbound titration variables that ensure that the $x_{i}$ are bound by 0 and 1.
%

$U^{barr}$, the second term of eq. \ref{eq:biasingpotential}, describes a barrier potential in the center of the titration coordinate, 
%
\begin{equation}
U^{barr}(\lambda_i) = -4\beta\left(\lambda_i - \frac{1}{2} \right)^2,
\label{eq:Ubarr}
\end{equation}
where $\beta$ is a parameter that determines the height of the barrier. Such potential increases the sampling time in the $\lambda$  coordinate endpoints and avoids unphysical vdW and electrostatic interactions associated with mixed titration state.

Finally, the last term of eq. \ref{eq:biasingpotential} is given by
%
\begin{equation}
U^{pH}(\lambda_i)=ln(10)k_BT(pH-\textrm{p}{K_a}^{mod})\lambda_i, 
\label{eq:UpH}
\end{equation}
% 
where $U^{pH}$ represents the pH-dependency of the deprotonation free energy, $k_B$ is the Boltzmann constant, $T$ is the system temperature, and $\textrm{p}{K_a}^{mod}$ is the model {\pka} for the titratable residue $i$.

\subsection{Scope} % - Ruibin 
This tutorial mainly covers system preparation, production run and restart, and post-processing data for the two mostly used CpHMD implementations, i.e., GBNeck2 implicit solvent based and GPU-accelerated implementation (GBNeck2-CpHMD) in the Amber pmemd.cuda platform, and the hybridge solvent and OpenMPI based implementation (Hybrid-CpHMD) in the CHARMM platform. For the GBNeck2-CpHMD version, we will first present how to prepare a system in explicit solvent like removing unwanted hetero atoms, ligands, ions, and waters, adding missing residues and atoms, adding dummy hydrogens, and post processing Amber parm7 file for updating exclusion list and changing GB radii. After that, we talk about how to pre-equilibrate the system before running CpHMD production simulations. For the production runs, we provide instructions on how to use conventional pH replica exchange on HPC and GPUs or asynchronous pH replica exchange on GPU clusters, and restarting existing jobs if necessary. For the Hybrid-CpHMD version, the system preparation process is smilar to GBNeck2-CpHMD but we mainly depend on CHARMM commands along with some mmtsb tools and we will learn how to add a pre-equilibrated water box, to miminize hydrogens with restraints and to head and equilibrate the whole system gradually. Instructions about how to use conventional pH replica exchange on HPC for Hybrid-CpHMD are followed as well. For all CpHMD simulations, choices about titratable residue types and pH conditions to use will be discussed for different use cases. At last, we will learn how to post process CpHMD data or the lambda files, and how to interpret the data with examples and explanations. We won't try to cover much traditional molecular dynamics analysis techniques but necessary solvent accessible surface area (SASA) and hydrogen bond analysis tools will be used to explain {\pka} upshifting or downshifting results. For all the topics, we will provide open-sourced bash or python scripts we used daily in our research.
% This section needs to describe what is covered in this review.
% Introduce CpHMD in the AMBER and CHARMM Packages, State the Types of CpHMD in AMBER and CHARMM, i.e. AMBER: Async pHREX with GBNeck2 and PME on GPUs, CHARMM: pHREX Hybrid-Solvent CpHMD on CPUs
% 1.) How to prepare a system for CpHMD
% 2.) Running CpHMD in CHARMM Equilibration to Production
% 3.) Running CpHMD in AMBER Equilibration to Production
% 4.) Analyzing Lambda Files, all the classic lambda values analysis procedures.

%Tutorials should endeavor to cover the specific task at hand, and also highlight how the steps might need to be modified (or additional care might need to be taken at particular points) to handle more general cases.

%The scope of the tutorial, as well as the expected proficiencies / outcomes for researchers who complete the tutorial, should be clearly defined.
%This will often happen in a specific section or subsection in the article itself.

We also have a GBSW implicit solvent based and a PME/GRF all-atom based implemenations in the CHARMM platform, but users are not encouraged to use them for speed issues. We are currently developing a GPU-accelerated all-atom implementation in the Amber pmemd.cuda platform, and the readers are welcomed to contact us for usage when it is finally finished.

\section{Prerequisites} % - Ruibin
CpHMD is based on classical molecular dynamics (MD) simulations, so it is a requirement that the user has some knowledge of MD and experience in running some MD jobs, especially in the Amber or CHARMM platform. Users with experience in other platforms like Gromacs, NAMD, or OpenMM are encouraged to go through some introductory tutorials for Amber or CHARMM to get started. For running a CpHMD simulation, it is important to know your system and research goal. If the user wants to know the {\pka} values for protein residues or their protonation states at certain pH conditions, the GBNeck2-CpHMD version is the best choice and Amber experience is required. If the user wants to investigate detailed proton-coupled conformational transormation, ion transfer, or ligand-protein binding/unbinding, especially when membrane system is involved, the Hybrid-CpHMD is currently the only choice and CHARMM experience is necessary. Like all computational chemistry calculations, it is important to know your system as much as possible before any serious simulations. With them, we can decide what salt concentration, temperature, pH range, and titratable residue types to use in CpHMD. Also, an expectation of time scale of the interested biological process is important to decide how long is required to run the CpHMD job.
% Who is the tutorial designed for, beginners or people with MD experience?
% What knoweledge should the user have prior to running CpHMD?
% Why should a person choose to run a simulation with CpHMD
% What should a person know before running a CpHMD simulation, pH-range, important titratable site... ect..

%Here you would identify prerequisites/background knowledge that are assumed by your work, as well as any software/license requirements.

\subsection{Background knowledge}
Knowledge of classical MD is required to understand what we can and cannot achieve through a MD simulation. Concepts in MD like force field, equation of motion, time step, convergence, thermostat, barostat, and so on are important to understand why we need them and the parameters we choose in our minimization, heating, equilibrating, and production runs. Of course, chemistry knowledge of pH and its significance in chemical, material, or biological systems are essential in entering this field.

Both the GBNeck2-CpHMD in Amber and Hybrid-CpHMD in CHARMM are primarily for Linux especially for HPC environments. The users are expected to have a good knowledge to use Linux command line tools for accessing and editing files, submitting jobs to servers, and retrieving simulation result files from remote Linux machines. Bash and/or Python scripting skills are required for pre- and post- processing files. For HPC environments, it is required to know the specific job submission and management system like SLURM or SGE installed in the HPCs. For GPU platform, knowledge of advantages of GPU acceleration in molecular dynamics and using tools such as 'nvidia-smi' to check GPU status and 'nvcc --version' to check CUDA version is necessary as well. The user can use any data visualisation tool(s), but matplotlib in Python and Jupyter notebooks is the first choice in our tool sets.  
% Tutorials should clearly define what concepts or abilities researchers will need to complete the tutorial (e.g., some proficiency in Python; experience with Jupyter notebooks; knowledge of classical MD; etc).

\subsection{Software/system requirements} % - All
% CHARMM Package Version, number of processors recommended for peak performance 
% AMBER Package version, Recommended GPUs
For GBNeck2-CpHMD in Amber, Amber 16 is the minimum version. Requirements for installing Amber can be found in the \href{https://ambermd.org/Installation.php}{Amber website}. For the GPU version, or specifically pmemd.cuda, the Nvidia GPU cards with at least compute compatibility sm3.0 should be installed with the latest driver. CUDA 10.1 is the primary CUDA version for running it but early versions from CUDA 7.5 also works. After successfully installing CUDA, it is important to set three environmental variables in Linux. CUDA\_HOME and LD\_LIBRARY\_PATH must point to the directories where the desired CUDA version is installed. We also need to specify CUDA\_VISIBLE\_DEVICES to GPU device indices. For example, 
\begin{lstlisting}
export CUDA_HOME=/usr/local/cuda-10.1
export LD_LIBRARY_PATH=\
            /usr/local/cuda-10.1/lib64
export CUDA_VISIBLE_DEVICES=0,1
\end{lstlisting}
will direct the Linux system to use the CUDA 10.1 version installed in the /usr/local directory and GPU with indices 0 and 1 will be used in calculations. After installing Amber, we should set the AMBERHOME environmental variable as the Amber package installation folder. 
\begin{lstlisting}
export AMBERHOME=/path/to/Amber
\end{lstlisting}

To prepare systems for CpHMD simulations, we usually use \href{http://blue11.bch.msu.edu/mmtsb/Main_Page}{mmtsb tool set} for CHARMM and our \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep}{cphmd-prep} tools for Amber. We also use a script-level \href{https://gitlab.com/shenlab-amber-cphmd/async_ph_replica_exchange}{asynchronous replica exchange} implementation in our lab or GPU pH replica exchange simulations. Scripts in the \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-analysis}{cphmd-analysis} are used for analyzing CpHMD lambda files, including convergence checking, {\pka} calculation, and replica profiles. Some simple trajectory based analysis like RMSD can also be done but the users can use any tools from cpptraj, pytraj, vmd, mdanalysis, and more. Note that these free tools usually depends on other free softwares, including OpenMM, Ambertools, scipy, pandas, matplotlib, etc. The instructions on how to install those softwares are covered in their individual pages and not discussed here. A general suggestion to use Python depending packages is to use Anaconda or Miniconda. With them, the user can create an environment specific for CpHMD related calculations. For example, after installing Anaconda or Miniconda, we can type 
\begin{lstlisting}
conda create --name cphmd python=3.7
\end{lstlisting}
Note that we choose Python version 3.7 because the OpenMM package currently only supports up to Python 3.7. Activate the environment named 'cphmd' by
\begin{lstlisting}
conda activate cphmd
\end{lstlisting}
Many packages can be installed through the conda and pip tool.
\begin{lstlisting}
conda install -c omnia pdbfixer parmed
conda install -c conda-forge ambertools \
                 mdanalysis matplotlib 
conda install -c anaconda numpy scipy pandas 
pip install f90nml
\end{lstlisting}
Whenever the user wants to do some calculations or analysis related to CpHMD, they can activate this 'cphmd' environment and use the installed tools without conflicting with other software. 
% Python Packages and Version required for Async-Replica Exchange 
% Python Packages and Version required for the CpHMD-Analysis Lambda Parser

%Tutorials should clearly define what system and/or software requirements the researcher will need to complete the tutorial (e.g., VMD version 1.9 or newer, AMBER, etc.). Tutorials requiring specific software packages must provide instructions and files for the referenced version of the software.

% \section{Content and links}

% A tutorial will normally draw on additional files and materials; clearly indicate where and how these are available, with links, and how they are being archived for the long-term and maintained so they stay current.
% You will likely want to reference your GitHub repository as a central point to access all of this information, and then the GitHub repository may link out to other content as needed.

% \section{Checklists}
% Tutorials do not necessarily require the use of a checklist as in Best Practices documents; however, they can include these if desired.
% Several useful checklist formats are available, with examples presented in \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates}.
% One example is shown here.

\section{Workflow for CpHMD simulations} % - Jack/Ruibin
% State an outline of the steps necessary for running the CpHMD simulations in general.

% Figure concept: Show a diagram of the steps needed to run a CpHMD simulation and images representing those steps.

 %--------------------------------------------- 
\begin{figure*}
    \centering
    \includegraphics[width=6.6in]{figs/workflow_scheme.pdf}
    \caption{\textbf{Generalized CpHMD workflow} 
    *Mock Figure*  
    }
    \label{fig:intro}
\end{figure*}
 %---------------------------------------------

\subsection{Hybrid-Solvent CpHMD in CHARMM} % - Jack

Here, the steps and processes involved in running CpHMD simulations in CHARMM are presented. 
In the first part of each section (System Preparation, Equilibration, and Production) an example of how to run CpHMD on the small soluble protein BBL (PDBid: 1w4h)\cite{Ferguson_JMolBiol_2005_v353_p427} is used as an example case. 
To follow along the please download the CpHMD tutorial from our \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/pH-REX_CpHMD_BBL_tutorial_in_CHARMM}{GitLab page}.
In the second part of the system preparation and equilibration section, supplemental information is provided discussing how to setup and run CpHMD simulation on transmembrane proteins.
Example input files for select portions of this process can be found from our \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/MEMB_pH-REX_CpHMD_CHARMM_Examples}{GitLab page}.

\subsubsection{System Preparation}

% Figure Concept: Show image of the ASP in the double protonated (syn-Configuration) State

To begin one needs to obtain a protein structure from the \href{https://www.rcsb.org/}{Protein Data Bank} (PDB).\cite{Berman_NucleicAcidsRes_2000_v28_p235}
If there are any missing residues in the downloaded protein structure they need to be filled in.
I good tool for this is SWISS-MODEL\cite{Waterhouse_NucleicAcidsRes_2018_v46_p296}, which allows you to replace missing residues based on homology to like proteins.
In the first part of this section we will go through an example for preparing the small protein BBL (PDBid: 1w4h)\cite{Ferguson_JMolBiol_2005_v353_p427} and in the second part discuss the preparation of transmembrane proteins.

With the protein structure for BBL downloaded from PDB we clean the structure using the MMTSB\cite{Feig_JMolGraphModel_2004_v22_p377} tool \href{http://blue11.bch.msu.edu/mmtsb/convpdb.pl}{convpdb.pl}. 
Additionally, all histidine residue names need to be changed to HSP and this is done in one step with the following command. 
%
\begin{lstlisting}[language=bash]
$ perl convpdb.pl -out charmm22 -segnames
-renumber 1 1w4h.pdb | sed "s/HSD/HSP/g" 
> start.pdb
\end{lstlisting}
%
Here, the output is formatted for the CHARMM22 protein force fields (-out charmm22), the segment names are included in the output (-segnames), the resids are renumbered starting at 1, and the input file is "1w4h.pdb."
Changing all HSDs to HSPs allows for titration of both $\delta$ and $\epsilon$ protons later.
Although the resids have been renumbered starting from one this does not have to be the case. 
Next we add all the missing hydrogen atoms to the protein using the HBUILD facility\cite{Brunger_Proteins_1988_v4_p148} in CHARMM.
For ASP and GLU, dummy protons are placed in the $syn$ position, and the force fields contains a dihedral energy barrier (raised to 3 kcal/mol) to maintain them in this position, so they don't move to the less favorable and rarely sampled $anti$ position ("ghost" position).\cite{Khandogin_BiophysJ_2005_v89_p141, Mongan_JComputChem_2004_v25_p2038} 
Addition of the hydrogens requires two rounds of energy minimization: first, with 50 steps of steepest decent (SD), and second 10 steps of adopted basis Newton-Raphson (ABNR) with the a heavy atom harmonically restrained (50 kcal/mol) in both cases.
This will help remove any unfavorable hydrogen positions. 
In most cases the N- and C-terminus ends of the protein are truncated and require capping. 
For the N-terminus use the CH$_3$CO (patch ACE) and for the C-terminus use NH$_2$ (patch CT2) or NHCH$_3$ (patch CT3).
In special cases the free ionized form, NH$_{3}^{+}$ and COO$^+$, is used and can be set to be titratable when CpHMD is turned on. 
The input file "step1\_add\_h.inp" is an example of how to perform these steps on BBL and if CHARMM is installed can be run with the following command.
%
\begin{lstlisting}[language=bash]
$ charmm -i step1_add_h.inp -o step1.out
\end{lstlisting}
%
Additionally, the example script will check for disulfide bonds using a sulfur to sulfur atom distance of 2.5 \mbox{\AA}. 
After running any input file make sure to visualize the output to confirm all protonation states and additional features are correctly constructed. 
Now We move on to making a water box to solvate the system using a small box of pre-equilibrated TIP3P waters.
This is done by generating a plan of water boxes along the X- and Y-axis, then stacking planar water boxes along the Z-axis.
After that, the box is reshaped to a truncated octahedron. 
In the example script the water is placed such that $\sim$10$\mbox{\AA}$ of padding is between the protein and the edge of the periodic boundary condition.
Use the following command and input file to construct the water box.
%
\begin{lstlisting}[language=bash]
$ charmm -i step2_make_waterbox.inp -o step2.out
\end{lstlisting}
%
In the final step, the protein is solvated with the water box by placing the protein at the center of the box and removing all water molecules within 2.8 $\mbox{\AA}$ of the protein heavy atoms. 
This step involves several phases of minimization; first the protein heavy atoms are fixed and the system is minimized using both SB and ABNR for 50 steps each, then five stage of minimization are used with reducing harmonic restraints (100, 50, 25, 5, 0 kcal/mol) on the protein heavy atoms using 50 steps of SD and 100 steps of ABNR each stage. 
Run this step with the following command and input file.
%
\begin{lstlisting}[language=bash]
$ charmm -i step3_solvate.inp -o step3.out
\end{lstlisting}
%
With the system built you can now move on to the next step. 
Again, always remember to visualize the system and make sure everything has been constructed to your expectations.

In some cases the protein crystal structure will include water molecular resolved in the PDB. 
This can be kept with a few modifications to the protocol above. 
First move all the water molecules into a separate PDB file and make sure the PDB file is properly formatted with the correct names, resnames, and has the correct column spacing.
Most error messages involving the handling of the crystal waters comes from the separate crystal water PDB not being properly formatted. 
You will need to make a new input file to add hydrogens to the water oxygen positions, like what was done with the protein and save the new water molecules as a CRD and PSF file. 
Now you will need to make another input file that reads in the protein coordinates and the crystal water coordinates and outputs a file with them merged together. 
The rest of the process now is similar to what was done above. 
You just need to make sure that in step3 solvate step the you don't accidentally delete these waters.

\noindent
\textbf{Transmembrane protein system preparation details}

Transmembrane proteins require a few additional steps to their preparation because they require a more rigorous equilibration. 
First we obtain the protein structure from the database \href{https://opm.phar.umich.edu/}{Orientations of Proteins in Membranes} (OPM). 
The OPM database orients the protein in the lipid bilayer using theoretical calculations with comparisons to experimental data, so an accurate placement of the protein in a lipid bilayer can be achieved.\cite{Lomize_Bioinformatics_2006_v22_p623}
Just like for soluble proteins any missing residues need to be added and then the missing hydrogens can be added.
Before the transmembrane protein system can be simulated with CpHMD, the lipid bilayer needs to be equilibrated using fixed charge MD simulations and this is done using any scalable MD engine like NAMD,\cite{Phillips_JComputChem_2005_v26_p1781} OpenMM,\cite{Eastman_PLoSComputBiol_2017_v13_pe1005659} or GROMACS.\cite{Abraham_SoftwareX_2015_v1-2_p19}
Thus, the protonation states of titratable groups need to be assigned based on any "Anticipated" pH condition described by wither experimental of physiological pH conditions. 
If there is not a pre-described pH condition the physiological pH (7.4) condition can be used giving the titratable residues the resulting charge states: Asp(-), Glu(-), Arg(+), Lys(+), Cys(0). 
Since His has a {\pka} near this pH condition special attention should be given to its protonation (singly or doubly protonated) and tautomeric state (HSD or HSE), based on any hydrogen bonding or salt-bridge interactions it can have. 
Like the soluble proteins the N- and C-terminus ends should be capped and if charged caps are used their protonation states should be considered and assigned. 
This is especially so for N-amino group since its solution {\pka} is 8.5,\cite{Nozaki_MethodsEnzymol_1967_v11_p715} only one pH unit above physiological pH.
Once hydrogens have been added and protonation states assigned, a short minimization is required to remove any poorly placed hydrogen positions.
Again, the system can be minimized in two round with 50 steps of SD followed by 10 steps of ABNR first with the heavy atoms either fixed or greatly harmonically restraint in the first round and with a reduced restraint in the second round.
With the protein prepared the rest of the system can be built and partially equilibrated to relax the systems structures (protein, lipids, ect.) following the CHARMM-GUI protocol for placing the lipids, solvating the system, and adding ions to achieve a solution salt concentration and to charge neutralize the system.\cite{Jo_PLoSONE_2007_v2_pe880}

\subsubsection{Equilibration}

\begin{checklist}{Equilibration Requirements}
\textbf{A list of files needed for the equilibration of a soluble protein}
\begin{itemize}
\item system PDB file 
\item system CRD file 
\item system PBC file
\item equilibration$\_$hphmd.inp
\end{itemize}
\end{checklist}

Equilibration of soluble protein systems for CpHMD is performed in several stages carried out at a single pH.
Since, CpHMD is turned on in the equilibration stages the residues allowed to titrate must be chosen, for this example Asp, Glu, and His are allowed to titrate, because of the pH range used during the production phase.
The pH chosen is normally the crystal pH condition of the PDB structure or a suitable pH for your system like physiological pH 7.4.
Before starting the simulations determine implicit solvent ionic strength (0.15 M in the example) and temperature (298 K in the example).
To use CpHMD in CHARMM add the following \href{https://hpc.nih.gov/apps/charmm/c42b2html/phmd.html}{PHMD module} to your input file after streaming in the GBSW parameters and reading the CpHMD parameters. 
%
\begin{lstlisting}
PHMD PAR 23 wri 25 PH 7.4 NPRI 500 PHFRQ 5 
    MASS 10.0 BARR 0.0 BETA 5.0 TEMP 300 LAM -
    sele resn ASP .or. resn GLU .or. resn HSP end
\end{lstlisting}
%
Here, PH sets the pH environment of the simulation, NPRI sets the frequency to print the lambda values, PHFRQ sets the update frequency of the lambda file, TEMP sets the temperature for the configuration of the titration coordinates and should be the same as the simulation temperature, MASS is the fictitious mass of the titration coordinate, BARR is the quadratic barrier height for each titration coordinate, BETA is the friction coefficient, and LAM tells the PHMD module to output lambda values.
The last line in the module states which residues should be allowed to titrate. 
The length of each stage depends on the system and should be adjusted accordingly, stages lasting 40 to 120 ps have been used by group for previous projects.
For the BBL example all heating/equilibration stages were run for 20 ps, since the system is small and stable. 
Firstly the system is heated, here we use an NVT ensemble and set and initial temperature to what was used to resolve the either the X-ray crystal or the cryo-EM structure. 
If the system does not require heating the initial and final temperatures can be set to be the same value and the system can equilibrate the water positions under the NVT ensemble.
Next we run a series of NPT ensemble equilibration stages with diminishing harmonic constraint on the heavy atom positions. 
The idea here is to slowly relax the system removing any high energy interactions and to allow for the initialization protonation states.
Using four equilibration stages the harmonic constraints on the protein heavy atoms are reduced with the following values 5.0, 1.0, 0.1 (kcal/mol), and then in the last stage all harmonic restraints are released.
After each equilibration make sure the energies of the system have equilibrated.
You are now ready to start the production run. 
The example equilibration input provided moves through each heating/equilibration stage in one input file (equilibration$\_$hphmd.inp).

\noindent
\textbf{Transmembrane protein system preparation details}

The equilibration of transmembrane proteins is a little more involved than soluble proteins, because of the equilibration of the lipid membrane. 
The transmembrane protein system has now undergone an initial relaxation according to the CHARMM-GUI protocol, and as mentioned in the system preparation fixed charge simulations need to be conducted to equilibrate the lipid bilayer.
Here, the protein heavy atoms are harmonically restrained using a 1.0 kcal/mol/$\mbox{\AA}^2$ until the lipid bilayer is equilibrating. 
Initially, the simulation is briefly simulated with a 0.2 ns NVT ensemble simulation.
Then the a much longer NPT ensemble simulation is run to fully equilibrate the simulation. 
These simulations should be run until the surface area per lipid, bilayer thickness, and lipid order parameters are near or within error limits of experimental lipid composition you are simulating.\cite{Huang_Shen_2020_}
These three lipid bilayer metrics are frequently used to assess the phase behaviour and fine structure of the simulated bilayer.\cite{Klauda_JPhysChemB_2010_v114_p7830}
Simulations up to 100 ns can be necessary for proper equilibration of the lipid bilayer.

Following equilibration of the lipid bilayer the system can now be equilibrated at a single pH with CpHMD. 
To start this processes, the dummy protons need to be added back on to the all Asp, Glu, and His side chains. 
This can be done in the exact same way as it was done in the system preparation of soluble proteins and should also be minimized in a similar way.
Now three short equilibrations can be run using CpHMD at a single pH, example inputs for these three equilibration are provided on our GitLab site [Add GitLab Information]. 
Like a soluble protein the pH used should be the crystal pH, physiological pH, or a experimentally relevant pH.
The first two simulations are performed with the protein harmonically restrained (1.0 kcal/mol/$\mbox{\AA}^2$) for 0.1 ns, but the first simulation uses the NVT ensemble and the second uses the NPT ensemble.
The third simulations continues with an NPT ensemble for 0.8 ns with all restraints removed. 
For membrane-enable CpHMD simulations the tmemb features needs to be added to the \href{https://hpc.nih.gov/apps/charmm/c42b2html/gbsw.html}{GBSW module} in the CHARMM input file. 
%
\begin{lstlisting}
gbsw hybrid sgamma 0.0 nang 50 conc 0.15 temp 310 
     tmemb 30 msw 2.5 rcyln 15                      
     sele SOLU end
\end{lstlisting}
%
Here, sgamma is the non-polar surface tension coefficient, nang is the number of angular integration points, and conc is the salt concentration.
The inputs variables tmemb, msw, and rcyln are used to build a low dielectric slab to represent the lipid bilayer in the GBSW model with a high dielectric hole centered around the protein.
The tmemb variable sets the slab thickness ($\mbox{\AA}$) and should be set to the average difference in the $Z$ positions of the lipid C2 atoms from the top and bottom leaflets, this will vary for different lipid types. 
The msw variable is the half membrane switching length\cite{Im_BiophysJ_2003_v85_p2900} for the dielectric transition region between the low-dielectric slab and the bulk solution, and is commonly set to 2.5 $\mbox{\AA}$.
Finally, rcyln is the radius of the high dielectric hole in the low dielectric slab and should set so there is minimal overlap between the protein and lipids.
The protein selection, the last line on in the GBSW module, required the protein selection to place the hole at the center of the protein.
Without restraints on the protein heavy atoms, a cylindrical restraint with a harmonic force constant (0.1 kcal/mol/$mbox{\AA}$) needs to be applied to the protein to prevent lateral drift of the protein. 
This can be done using the MMFP facility in CHARMM.

\subsubsection{Production}

\begin{checklist}{Production Run Requirements}
\textbf{A list of files needed for the production run of a soluble protein}
\begin{itemize}
\item system PSF file
\item system CRD file 
\item system PBC file
\item final equilibration RST file
\item production$\_$hphrex.inp
\item rep.cmd
\end{itemize}
\end{checklist}

The CpHMD production stage is the same for both soluble and transmembrane proteins. 
The only differences are in the set up of the GBSW module to include a low dielectric slab to represent the lipid bilayer for transmembrane simulations, which is discussed in the equilibration stage. 
To begin the production run one must chose a pH range to simulation over and how many 
pH replicas to use. 
For instance, a protein of interest may be active over a particular pH range or you might want to capture conformational dynamics that include specific titratable residues such as Asp or His. 
Next you need to decide how many pH replicas to use for your production run. 
Somethings to consider are how many CPUs you have available and how many CPUs you want to allocate to each replica. 
Simulations with CHARMM do not scale well past 8 CPUs, so for a job using 16 replicas you will need 128 CPUs. 
The spacing of the pH replicas needs to be close enough to obtain sufficient exchange probabilities (e.g., 20$\%$).
Typically, to obtain overlap between potential energy distribution of neighboring replicas a pH spacing in increments of 0.25 or 0.5 is required. 
Exchange attempts are made every 500 to 1000 steps, and a typical CpHMD simulation of this type will have an average exchange rate of $\sim$40$\%$, if exchange rates are too low more replicas can be added to fill in the spacing between replicas.
To achieve enhanced sampling several replicas show move across all pH replicas.
To use pH-based replica exchange in CHARMM the following \href{https://hpc.nih.gov/apps/charmm/c39b2html/repdstr.html}{REPD} module must be used.
%
\begin{lstlisting}
REPD NREP 8 PHMD EXLM FREQ 1000 UNIT 17
\end{lstlisting}
%
Here, the REPD command initiates replica exchange, NREP states the number of replicas to use, in this case 8, and PHMD states this is pH-based replica exchange and allows for the exchange of $\lambda$ along with the spatial coordinates.
Next, EXLM allows for Hamiltonian exchange and FREQ is how often an exchange is attempted. 
Then to run several replicas in parallel CHARMM reads in a rep.cmd file for each pH replica with the syntax, rep.cmd$\_$0, rep.cmd$\_$1, to rep.cmd$\_$(NREPS), using the following command in the input file.
%
\begin{lstlisting}
stream rep.cmd
\end{lstlisting}
% 
Notice this command lacks the '$\_0$' portion of the name.
In each of the rep.cmd file you will have the command that states the pH environment of the at replica.
%
\begin{lstlisting}
PHMD RESPH OLDPH @crysph NEWPH @ph PKATEMP @temp
\end{lstlisting}
%
Where OLDPH is the pH used in the equilibration and NEWPH is the pH of the replica.
finally, PKATEMP should be set to the same temperature of the simulation. 
The rest of the replica exchange set up follows a standard molecular dynamics run in CHARMM.
Production run inputs for BBL (production$\_$hphrex.inp and rep.cmd) can be used to run one 2 ns stage of pH-Replica Exchange CpHMD.

\subsubsection{General MD and CpHMD Settings}
% CHARMM Specific
% Things to Include: 
%   -) FFs
%   -) Thermostats
%   -) Baristas
%   -) Cutoffs
%   -) ect...

For all simulations with CHARMM the protein is represented with CHARMM22/CMAP force fields.\cite{MacKerell_JPhysChemB_1998_v102_p3586,MacKerell_JComputChem_2004_v25_p1400}
For the solvent the CHARMM modified TIP3P water model are used and for any ions or lipid molecules that might be included the CHARMM36 force fields are used.\cite{Klauda_JPhysChemB_2010_v114_p7830}
To be consistent with the CHARMM36 development protocol all van der Waals energies are calculated with a switching function from 10 to 12 $\mbox{\AA}$ with a cutoff at 14 $mbox{\AA}$. 
All simulations are conducted under periodic boundary conditions. 
The PME method\cite{Darden_JChemPhys_1993_v98_p10089} is used to calculate the long-range electrostatic energies and forces with a real space calculation cutoff of 12 $\mbox{\AA}$, 1 $\mbox{\AA}$ grid spacing, and a 6th-order spline interpolation. 
In CHARMM the non-bond neighbor list is updated heuristically. 
Additionally, the SHAKE\cite{Ryckaert_JComputPhys_1977_v23_p327} algorithm in CHARMM is used to constrain all bonds involving hydrogens to allow for a 2 fs time step. 
For NPT simulations the temperature is maintained with the N\'{o}se-Hoover thermostat.\cite{Nose_MolPhys_1984_v52_p255, Hoover_PhysRevA_1985_v31_p1695}
The pressure is maintained with the CHARMM implemented Langevin piston pressure-coupling algorithm.\cite{Feller_JChemPhys_1995_v103_p4613}

For CpHMD simulations, the default GBSW radii\cite{Nina_JPhysChemB_1997_v101_p5239, Chen_JAmChemSoc_2006_v128_p3728} are used for the proteins and any small molecules the generic atomic radii in the GBSW radii file or van der Waals radii in the force field\cite{Knight_JComputChem_2011_v32_p2909} should be used. 
As shown in the example inputs the titration ($\lambda$) and tautomer ($\chi$) should be updated every 5 MD steps, allowing for water and lipid relaxation.\cite{Wallace_JChemTheoryComput_2011_v7_p2617}
A friction coefficient of 5 ps$^{-1}$ should be used for Langevin dynamics and the mass of the titration coordinate ($\lambda$ and $\chi$) should be 10 atomic mass units.

\subsection{All-atom PME CpHMD in CHARMM} % - Yandong

Minor modification of the workflow for 
hybrid-solvent CpHMD above is needed 
to setup and run an all-atom PME CpHMD 
simulation in CHARMM.

To maintain a constant net charge during 
all-atom PME CpHMD simulations, 
titratable water molecules  
or coions should be added to the 
system as built above \cite{Wallace_Shen_2012_J.Chem.Phys.}.
In specific, each acidic (ASP or GLU) or 
basic residue (HSP or LYS) is 
coupled with one coion that can convert 
to a hydroxide (TIPU) or hydronium (TIPP) 
\cite{Chen_Shen_2013_Biophys.J.}. 
Together with the salt ions, the coions 
can be placed in the water box with 
the input script step3\_add\_ions.inp.
Likewise, water molecules within 2.8 \AA\ 
of ions and the protein are removed 
with the script step4\_solvate.inp.

To run an all-atom PME CpHMD simulation, 
the GBSW module should be 
removed first such that PME module is 
utilized in either conformational or 
titration dynamics \cite{Huang_Shen_2016_J.Chem.TheoryComput.}. 
Second, the coupling information should 
be specified. 
As to BBL that 
contains 3 ASP, 3 GLU and 2 HSP, below 
are the parameters required when PHMD module 
is invoked.

%
\begin{lstlisting}
PHMD PAR 23 wri 25 PH 7.0 NPRI 500 PHFRQ 5 
    MASS 10.0 BETA 5.0 TEMP 300 LAM -
    sele resn ASP .or. resn GLU .or. TIPU .or. 
    resn HSP .or. TIPP end - 
    qcouple 8 - 
    resi 4  resc 4789 -
    resi 16 resc 4790 -
    resi 20 resc 4791 -
    resi 36 resc 4792 -
    resi 37 resc 4793 -
    resi 39 resc 4794 -
    resi 17 resc 4771 -
    resi 41 resc 4772 
\end{lstlisting}
%
Here TIPU and TIPP are selected as titratable too. 
The parameter qcouple, 8 in this case, indicates 
the count of coupled pairs listed behind. 
The second and forth 
colunms correspond to the indices of 
titratable residues and the coions 
associated respectively.

Finally, the calculated \pka's are 
corrected with the following formula 
to account for the finite-size error 
\cite{Huang_Shen_2016_J.Chem.TheoryComput.}.

\begin{eqnarray}
\label{eq:pka_corr}
\Delta\Delta{G}^{\rm offset} &=&
\frac{2 \pi}{3} \kappa \gamma^{\rm solv} Q 
\left(\frac{N}{V}
- \rho^{\rm pure} \right) \\
\Delta pK_a^{\rm corr} &=& \pm \frac{\Delta\Delta{G}^{\rm offset}}{ln(10)RT},
\end{eqnarray}
where $\gamma^{\rm solv}$ is 0.764 e$\cdot$ {\AA$^2$} 
for TIP3P water model. 
$\kappa$ is the electrostatic constant, 
which is 332.0 kcal 
$\cdot$ {\AA} $\cdot$ mol$^{-1}$e$^{-2}$. 
The solvent number density $\rho^{\rm pure}$ is 
0.0333679 {\AA}$^3$ for water at ambient temperature 
and pressure. 
Q is -1 for acidic groups and +1 for basic groups. 
The negative sign is for acid groups and positive sign 
is for basic groups.
Thus, $\Delta$\pka$^{\rm corr}$ is negative for both 
acidic and basic groups. 
N is the number of solvent and V is the volume of the box 
measured at equilibrium, which can be approximated 
with the initial volume $V^{\rm init}$ times 0.93. $V^{\rm est} = L^3$ for a cubic box and $V^{\rm est} = 0.77L^3$ for a truncated octahedron box, where L denotes the cubic 
box size. R is the Boltzmann constant and T is the temperature.

\subsection{Implicit-Solvent CpHMD in AMBER} % - Ruibin
\subsubsection{System Preparation}
To prepare a protein system in implicit solvent for CpHMD in Amber, the process is similar to the hybrid-solvent CpHMD in CHARMM but much simpler. We have a \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep}{cphmd-prep} package (and soon a web server) to prepare the Amber GBNeck2-CpHMD inputs and instructions to run related CpHMD simulations by one line bash command via the 'CpHMD\_Prepare.sh' script. The script accepts either a PDB ID (-p|--pdbid flag) whose PDB file is downloadable from RCSB to the working machine or a user-prepared PDB file (-f|--pdbfile flag) stored in the machine. After that, PDBFixer (powered by OpenMM) is used to extract the desired chain (-c|--chainid flag) and model (-m|--modelid) protein structure only and remove all other parts like ions, ligands, waters, or unwanted chains. Missing heavy atoms and residues are then added by PDBFixer. Importantly, capping ligands ACE and NH$_2$ are added to cap the N and C terminus to suppress unwanted electrostatic effects from otherwise charged terminus. After the PDB file is prepared, tleap in Amber (or Ambertools) is used to build hydrogen and dummy hydrogen atoms according to the titratable types (-type|--residuetypes flag) set by the user. Specifically, the user can choose one or a combination in 'Asp Glu His Cys Lys'. After the system is prepared, it is required by GBNeck2-CpHMD that the GB radii of His HD1 and HE2, Cys SG, Asp OD1 and OD2, and Glu OE1 and OE2 atoms should be changed to 1.17, 2.00, 1.40, and 1.40 {\AA} respectively; and the interactions between the two dummy hydrogens in Asp and Glu should be excluded in the non-bonded interaction terms. This is automatically done in `CpHMD\_Prepare.sh` too.

Even though our cphmd-prep tool can prepare a system for most cases, the user with different needs can use other tools to do the same or different tasks as instructed by cphmd-prep. Nevertheless, here we show two simple examples of how to use 'CpHMD\_Prepare.sh' to prepare a GBNeck2-CpHMD simulation system. The first is an EFGR kinase with provided PDB ID 5U8L and we want to know the protonation states and possible {\pka's} of His, Cys, and Lys near the psychological pH condition 7.4. 
%
\begin{lstlisting}
CpHMD_Prepare.sh -p 5U8L -m A -pmin 5.5
                 -pmax 9.5 -type 'His Cys Lys'
\end{lstlisting}
%
Note that we omit chain ID(s) because 5U8L only has one chain whose chain ID is A but we do specify the model ID as A because there are two alternative model IDs for 5U8L, A and B. We choose the pH range from 5.5 to 9.5 with a default interval 0.5 units, which covers about 2 pH units from the two directions of 7.4, making the total number of independent pH calculations to be 9. The other important settings are the simulation length, temperature and implicit ionic strength, which are 20 ns, 300 K, and 150 mM by default. Four important files will be generated if the command runs successfully. The 5U8L\_ModelA\_A\_fixed.rst7 file is an Amber topology rst7 file with initial coordinates; the cphmd\_5U8L\_ModelA\_A\_fixed.parm7 file is an Amber parameter parm7 file; the 5U8L\_ModelA\_A\_fixed\_phmdin file is an GBNeck2-CpHMD job control input file; the gbneck2\_input.parm file is the GBNeck2-CpHMD parameter file which is the same for all GBNeck2-CpHMD calculations. Other files including Amber input files are also generated and they will be explained later.

The second example is to quickly estimate the {\pka's} of all Asp, Glu, His, Cys, and Lys residues in SNase with a PDB file 3BDC.pdb. The command is
%
\begin{lstlisting}
CpHMD_Prepare.sh -f 3BDC.pdb -m A -pmin 1.0 
                 -pmax 12.0 -l 4
\end{lstlisting}
%
Here we feed the script with the PDB filename and use a much larger pH range from 1.0 to 12.0 so that the Asp, with a model {\pka} of 3.7, and Lys, with a model {\pka} of 10.4, residue types can both titrate in the pH range. In addition, we use a much shorter length, 4 ns, because we only want a rough estimation of the {\pka's}. We omit the -type flag because the all five residue types are included by default. Similar files as the previous example will be generated. The user may notice that many more Amber input files are generated. They differ in the solveph keyword that is required to perform CpHMD calculations in different pH conditions.

System minimization with constraints on heavy atoms (except for two terminal and capping residues whose all atoms are allowed to move) is performed after preparing the aforementioned Amber input files. The constraint strength is 50 kcal/mol/\AA$^2$. The input file has a name end with *\_fixed\_mini.mdin which can be found in the generated files. The minimization last 1000 steps with the first 500 steps for steepest descent algorithm and the remaining 500 steps for conjugate gradient steps. After minimization, a *\_mini.rst7 is obtained for the next step. For convenience, all commands for minimization, equilibration, and production runs are listed in the README file generated by the 'CpHMD\_Prepare.sh' file, same for the below steps.

\subsubsection{Equilibration}
For implicit solvent based CpHMD, heating is very fast and thus not required. We do have four short steps of equilibration that the constraintes on heavy atoms are gradually released from 5.0, 2.0, 1.0, to 0.0 kcal/mol/\AA$^2$. Each step last 2000 steps or 4 ps and the total length is 16 ps. In equilibration and production steps, CpHMD is turned on by setting the iphmd keyword to 1 for GBNeck2-CpHMD.

\subsubsection{Production}
For production runs, we can either use CPU pH replica exchange (pHREX), GPU independent pH (pHInd), or GPU asynchronous pH replica exchange styles (\href{https://gitlab.com/shenlab-amber-cphmd/async_ph_replica_exchange}{Async-pHREX}). Replica exchange simulations usually need shorter time for convergence in {\pka} predictions but GPU acceleration can overcome the simulation time problem, and thus Async-pHREX is recommended over the other two styles. Although the asynchronous replica exchange we implemented is in the script level, it still combine the significantly good parts of GPU acceleration and replica exchange enhanced sampling, especially for large systems or the GPUs are the old models. We run pHREX just like we run other replica exchange simulations in Amber, i.e., we need a group file which contains the command line flags and inputs for every pH condition and then run with mpirun. For pHInd, we run each pH condition independently on different GPUs in the same machine by changing the CUDA\_VISIBLE\_DEVICES environment variable or even on GPUs in different machines. For Async-pHREX, we have to use a template input file from which temporary input files for different pH conditions at a replica exchange attempt can be generated. Output files are concatenated according to pH conditions instead of replicas in the pHREX method. For all these three methods, we can use the -rex flag to control which method is going to be used and the related commands can be found in the README file as well.
% Here we can describe a few different styles of production run with CpHMD. 
% this section can be broken down into three sections
%   1.) Single-pH runs with CpHMD Simulations in AMBER
%   2.) Async-pHREX with CpHMD simulations in AMBER

\subsubsection{MD Specific Settings}
For best performance regarding {\pka} prediction speed and accuracy, there are several settings that we think should be fixed in Amber GBNeck2-CpHMD. GBNeck2 implicit solvent model works best with the Amber FF14 force field. Langevin thermostat with a collision frequency 1 ps$-1$ is used to maintain system temperature. Non-bonded cutoff is 999.9 \AA which means not cutoff is used practically. CpHMD specific lambda files are usually saved every 1000 MD steps while trajectory files are saved every 10000 steps to save I/O time. 
% Amber Specific
% Things to Include: 
%   -) FFs
%   -) Thermostats
%   -) Baristas
%   -) Cutoffs
%   -) ect...

\section{Lambda File Analysis} % - Jack / Vini

The lambda file outputs provide us the propagation of $\lambda$ (titration state) and $x$ (tautomeric state) coordinates from titratable residues along the CpHMD simulation. Such information allows us to calculate the protonation state changes during the dynamics, the pKa value of each ionizable residue and also monitoring the pKa convergence. In order to facilitated these analysis, our research group developed a Python tool (cphmdanalysis.py) that possesses several classes to run calculations in lambda files and log files, and can be found at GitLab: \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-analysis}{CpHMD-Analysis}. This tool can also plot the unprotonation fraction and pKa of selected titratable residues.

From each simulated pH, we can calculate the population of the $i$th titratable residue in the protonated state ($N_i^{\rm Prot}$, $\lambda_i < 0.1$) and in the unprotonated state ($N_i^{\rm Unprot}$, $\lambda_i > 0.9$). Thus, we can obtain the unprotonated fraction of this residue,

\begin{equation}
S_i = \frac{N_i^{\rm Unprot}}{N_i^{\rm Prot} + N_i^{\rm Unprot}}.
\end{equation}

The convergence of the $S_i$ values during the CpHMD steps indicates that the simulation reached a protonation state sampling equilibrium.

Finally, we can fit the calculated $S_i$ at several pHs using the generalized form of the Henderson-Hasselbach equation,

\begin{equation}
S_i = \frac{1}{1+10^{n(\rm pK_{a,i} - pH)}},
\end{equation}
where $n$ is the Hill coefficient, and indicates the degree of cooperativity between the titratable groups.

% Here we can describe how to analyze CpHMD results
%   1.) Calculations of Unprot. Fractions
%   2.) Calculations pKa  
%   3.) How to Monitor pKa Convergence
%   3.) Additional Analysis could include proton-coupling, ect...

% Figure Concept: Here we should have a 2 panel figure. 
%   -) Panel 1: Raw Lambda Value versus time @ a single pH with a duel y-Axis and over the plot we showing the running S-value. 
%   -) Panel 2: We show a titration Curve

\section{Parameterization} % - Jack/Ruibin
To perform CpHMD simulations on a new titratable residue, accurate values for the parameters of $U^{model}$ are required. The basic method for obtaining these parameters is to simulate a model compound containing an analog of the residue in question at several fixed values of $\theta_{i}$ and $\theta^{x}_{i}$ with $U^{barr}$ and $U^{pH}$ set to 0, compute the derivatives of the free energy with respect to $\theta_{i}$ and $\theta^{x}_{i}$, and fit those derivatives to eq. \ref{eq:Umod_1d} and \ref{eq:double}.

Note that these fittings should be performed on the $\theta$ and $\theta_{x}$ variables directly rather than on $\lambda$ and $x$ because the nonlinearity in the definitions of $\lambda$ and $x$ cause difficulties in the fitting.

Typically we are interested in three different kinds of titratable residues in most proteins, residues, such as Lys, that have a single titratable site, residues, such as Glu and Asp, that have two titratable sites with identical microspcopic pKa's, and residues, such as His, that have two titration sites with different microscopic pKa's. Each of these cases have different forms for their $U^{model}$ and therefore require somewhat different methods.

First, for residues like Lys with a single titratable site, we only need to consider a single titration variable, $\theta$. For these residues we construct a model compound for which the experimental pKa is known, such as the blocked amino acid, and run simulations with fixed values of $\theta$ and fit the resulting $\partial U/\partial\theta$ to eq. \ref{eq:Umod_1d}.

Second, for residues, like His, that contain two nonidentical intertacting titratable sites, eq. \ref{eq:double} can be written as

\begin{align}
U^{model}= & A_{10}\lambda^{2}x^{2}+2\left(A_{1}B_{1}-A_{0}B_{0}\right)\lambda x+\nonumber \\ & 2\left(A_{0}B_{0}-A_{1}B_{1}-A_{10}B_{10}\right)\lambda^{2}x+ \nonumber \\ & A_{1}\lambda^{2}-2A_{1}B_{1}\lambda\label{eq:his}
\end{align}

The six parameters $A_{0}$, $A_{1}$, $B_{0}$, $B_{1}$, $A_{10}$, and $B_{10}$ can be obtained by running simulations on the model compound at different values of $\theta$ and $\theta_{x}$ with $U^{model}$, $U^{barr}$, and $U^{pH}$ set to 0, computing the derivatives of the free energy with respect to $\theta_{i}$ and $\theta^{x}_{i}$, and fitting those derivatives to this equation. Note once again that although eq. \ref{eq:his} is written in terms of $\lambda$ and $x$ for clarity, the fitting should be performed against $\theta$ and $\theta_{x}$ to avoid issues arising from nonlinearities near the endpoints.

Finally, for residues like Asp and Glu, that contain two identical interacting titratable sites, eq. \ref{eq:double} reduces to

\begin{align}
    U^{model}= & \left(R_{1}\lambda^{2}+R_{2}\lambda+R_{3}\right)\left(x-R_{4}\right)^{2}+\left(R_{5}-R_{1}R_{4}^{2}\right)\lambda^{2} \nonumber\\ & -\left(2R_{5}R_{6}+R_{2}R_{4}^{2}\right)\lambda \label{eq:asp}
\end{align}

As before, the six parameters $R_{1}$ to $R_{6}$ can be obtained by running simulations on the model compound at different values of $\theta$ and $\theta_{x}$ with $U^{model}$, $U^{barr}$, and $U^{pH}$ set to 0, computing the derivatives of the free energy with respect to $\theta$ and $\theta^{x}$, and fit those derivatives to this equation. Note once again that although eq. \ref{eq:asp} is written in terms of $\lambda$ and $x$ for clarity, the fitting should be performed against $\theta$ and $\theta_{x}$ to avoid issues arising from nonlinearities near the endpoints.

\subsection{Parameterization settings and fitting procedure}
As discussed above, the key step to parameterizing a new residue is to run simulations with fixed values of $\theta$ and $x$ with $U^{barr}$, $U^{model}$, and $U^{pH}$ set to 0. This can be accomplished in Amber by setting the options prlam to false, prderiv to true, and phtest to true in the phmdin file and setting ph{\_}theta to the desired values of $\theta$ and $x$ and vph{\_}theta to 0 in the phmdstrt file.

As discussed above, the procedure for parameterizing a residue, such as Lys, with a single titratable site is pretty straightforward, just run at several values of $\theta$ (We generally find that $\theta=\left(0.0,0.2,0.4,0.6,0.785,1.0,1.2,1.4,1.5708\right)$ work well). In contrast, parameterizing two site residues is more complicated because the energy surface is two-dimensional. What follows is therefore a description of the method we usually use to parameterize these residues, although other approaches are possible.

First, for residues, such as His, with two titration sites with different microscopic pKa's, we generally run simulations at $\theta_{x}=0$ and $\theta=\left(0.0,0.2,0.4,0.6,0.785,1.0,1.2,1.4,1.5708\right)$. The parameters $A_{0}$ and $B_{0}$ can then be computed by fitting $\partial U/\partial\theta$ to

\begin{equation}
    \frac{\partial U}{\partial\theta}=2A_{0}\sin\left(2\theta\right)\left(\sin^{2}\left(\theta\right)-B_{0}\right).
\end{equation}
Next, the parameters $A_{1}$ and $B_{1}$ can be computed by running simulations at $\theta_{x}=1.5708$ and $\theta=\left(0.0,0.2,0.4,0.6,0.785,1.0,1.2,1.4,1.5708\right)$ and fitting $\partial U/\partial\theta$ to

\begin{equation}
    \frac{\partial U}{\partial\theta}=2A_{1}\sin\left(2\theta\right)\left(\sin^{2}\left(\theta\right)-B_{1}\right).
\end{equation}
Finally, $A_{10}$ and $B_{10}$ can be computed by running simulations at $\theta=1.5708$ and $\theta_{x}=\left(0.0,0.2,0.4,0.6,0.785,1.0,1.2,1.4,1.5708\right)$ and fitting $\partial U/\partial\theta_{x}$ to

\begin{equation}
    \frac{\partial U}{\partial\theta_{x}}=2A_{10}\sin\left(2\theta_{x}\right)\left(\sin^{2}\left(\theta_{x}\right)-B_{10}\right).
\end{equation}

Finally, for residues, such as Asp, that have two identical interacting titrating sites we have to find parameters $R_{1}$ to $R_{6}$. To start with, we generally run simulations at combinations of $\theta=\left(0.0,0.2,0.4,0.6,0.785,1.0,1.2,1.4\right)$ and  $\theta_{x}=\left(0.0,0.2,0.4,0.6,0.785,1.0,1.2,1.4,1.5708\right)$, compute $A\left(\theta\right)$ and $B\left(\theta\right)$ at each value of $\theta$ by fitting $\partial U/\partial\theta_{x}$ to

\begin{equation}
    \frac{\partial U}{\partial\theta_{x}}=2A\left(\theta\right)\sin\left(2\theta_{x}\right)\left(\sin^{2}\left(\theta_{x}\right)-B\left(\theta\right)\right),
\end{equation}
compute $A\left(\theta_{x}\right)$ and $B\left(\theta_{x}\right)$ at each value of $\theta_{x}$ by fitting $\partial U/\partial\theta_{x}$ to

\begin{equation}
    \frac{\partial U}{\partial\theta}=2A\left(\theta_{x}\right)\sin\left(2\theta\right)\left(\sin^{2}\left(\theta\right)-B\left(\theta_{x}\right)\right),
\end{equation}
obtain estimates of $R_{1}$, $R_{2}$, and $R_{3}$ by fitting $A\left(\theta\right)$ to

\begin{equation}
    A\left(\theta\right) = R_{1}\sin^{4}\left(\theta\right)+R_{2}\sin^{2}\left(\theta\right)+R_{3},
\end{equation}
setting $R_{4}=0.5$ (guaranteed because the two titrating sites are identical), obtain an estimate of $R_{5}$ by fitting $A\left(\theta_{x}\right)$ to

\begin{equation}
    A\left(\theta_{x}\right)=a_{0}\sin^{4}\left(\theta_{x}\right)+a_{1}\sin^{2}(\theta_{x})+R_{5},
\end{equation}
and finally obtain an estimate of $R_{6}$ by fitting $B\left(\theta_{x}\right)$ to

\begin{equation}
    B\left(\theta_{x}\right)=a_{0}\sin^{4}\left(\theta_{x}\right)+a_{1}\sin^{2}(\theta_{x})+R_{6},
\end{equation}
% Here is a single-column checklist that consists of multiple sub-checklists
\begin{Checklists}

\begin{checklist}{A list}
\textbf{Single-column checklists are also straightforward by removing the asterisk}
\begin{itemize}
\item First thing let's do an item which breaks across lines to see how that looks
\item Also remember
\item And finally
\end{itemize}
\end{checklist}

\begin{checklist}{Another list}
\textbf{This is some further description.}
\begin{itemize}
\item First thing
\item Also remember
\item And finally
\end{itemize}
\end{checklist}

\end{Checklists}

\section{Author Contributions}
%%%%%%%%%%%%%%%%
% This section mustt describe the actual contributions of
% author. Since this is an electronic-only journal, there is
% no length limit when you describe the authors' contributions,
% so we recommend describing what they actually did rather than
% simply categorizing them in a small number of
% predefined roles as might be done in other journals.
%
% See the policies ``Policies on Authorship'' section of https://livecoms.github.io
% for more information on deciding on authorship and author order.
%%%%%%%%%%%%%%%%

(Explain the contributions of the different authors here)

% We suggest you preserve this comment:
For a more detailed description of author contributions,
see the GitHub issue tracking and changelog at \githubrepository.

\section{Other Contributions}
%%%%%%%%%%%%%%%
% You should include all people who have filed issues that were
% accepted into the paper, or that upon discussion altered what was in the paper.
% Multiple significant contributions might mean that the contributor
% should be moved to authorship at the discretion of the a
%
% See the policies ``Policies on Authorship'' section of https://livecoms.github.io for
% more information on deciding on authorship and author order.
%%%%%%%%%%%%%%%

(Explain the contributions of any non-author contributors here)
% We suggest you preserve this comment:
For a more detailed description of contributions from the community and others, see the GitHub issue tracking and changelog at \githubrepository.

\section{Potentially Conflicting Interests}
%%%%%%%
%Declare any potentially competing interests, financial or otherwise
%%%%%%%

Declare any potentially conflicting interests here, whether or not they pose an actual conflict in your view.

\section{Funding Information}
%%%%%%%
% Authors should acknowledge funding sources here. Reference specific grants.
%%%%%%%
We acknowledge financial support from the National Institutes of Health (R01GM098818 and R01CA256557).

\section*{Author Information}
\makeorcid

\bibliography{Bib/JanaShenLab.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% APPENDICES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\appendix

\end{document}

### 2022-01-23 cphmd_tutorial.tex
Initial Overleaf Import

--- 2022-01-23 Draft/cphmd_tutorial_draft.tex
@@
 \usepackage[version=4]{mhchem}
-\usepackage{siunitx}
+\usepackage{siunitx,bm}
 \usepackage{hyperref} % Not original - Added by Jack
@@
 % you should update the minor version number in preprints and major version number of submissions.
-\newcommand{\githubrepository}{\url{https://github.com/myaccount/homegithubrepository}}  %this should be the main github repository for this article
+\newcommand{\githubrepository}{\url{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial}}  %this should be the main github repository for this article
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{A Guide to Continuous Constant pH Molecular Dynamics in AMBER and CHARMM v[\versionnumber]}
+\title{A Guide to the Continuous Constant pH Molecular Dynamics 
+Methods in Amber and CHARMM v[\versionnumber]}
 
-\author[1]{Jack A. Henderson}
-\author[1]{Ruibin Liu}
-\author[1,\authfn{2}]{Julie A. Harris}
-\author[1,\authfn{3}]{Yandong Huang}
+\author[1,\authfn{1},\authfn{2}]{Jack A. Henderson}
+\author[1,\authfn{1}]{Ruibin Liu}
+% \alsoaffiliation{
+% Joint first author
+% }
+\author[1,\authfn{3}]{Julie A. Harris}
+\author[1,\authfn{4}]{Yandong Huang}
 \author[1]{Vinicius M. de Oliveira}
+%\author[1]{Paween Mahinthichaichan}
 \author[1*]{Jana Shen}
@@
 
-\orcid{Jack A. Henderson}{0000-0001-6675-7944}
-\orcid{Ruibin Liu}{0000-0001-8395-9353}
-\orcid{Julie A. Harris}{0000-0002-1130-6457}
-\orcid{Yandong Huang}{0000-0002-1452-6383}
-\orcid{Vinicius M. de Oliveira}{0000-0003-0927-3825}
-\orcid{Jana Shen}{0000-0002-3234-0769}
+\orcid{Jack A. Henderson}{https://orcid.org/0000-0001-6675-7944}
+\orcid{Ruibin Liu}{https://orcid.org/0000-0001-8395-9353}
+\orcid{Julie A. Harris}{https://orcid.org/0000-0002-1130-6457}
+\orcid{Yandong Huang}{https://orcid.org/0000-0002-1452-6383}
+\orcid{Vinicius M. de Oliveira}{https://orcid.org/0000-0003-0927-3825}
+\orcid{Jana Shen}{https://orcid.org/0000-0002-3234-0769}
 
-% \contrib[\authfn{1}]{These authors contributed equally to this work}
+\contrib[\authfn{1}]{These authors contributed equally to this work}
 % \contrib[\authfn{2}]{These authors also contributed equally to this work}
 
-\presentadd[\authfn{2}]{Mapware, Washington DC}
-\presentadd[\authfn{3}]{Jimei University, College of Computer Engineering, Xiamen, China}
+\presentadd[\authfn{2}]{Scorpion Therapeutics, Boston, MA}
+\presentadd[\authfn{3}]{ComputChem LLC, Baltimore, MD}
+\presentadd[\authfn{4}]{Jimei University, College of Computer Engineering, Xiamen, China}
 % \presentadd[\authfn{4}]{Department, Institute, Country}
 
-%\blurb{This LiveCoMS document is maintained online on GitHub at \githubrepository; to provide feedback, suggestions, or help improve it, please visit the GitHub repository and participate via the issue tracker.}
+\blurb{This LiveCoMS document is maintained online on GitLab at \githubrepository.
+To provide feedback, suggestions, or help improve it, please visit the GitLab repository and participate via the issue tracker.}
 
@@
 
+% \lstnewenvironment{listings}[1][] 
+%  {\lstset{frame=off,escapechar=`,breaklines, breakindent=10pt, breakatwhitespace=True, linewidth=3.4in, #1}}
+% {}
+
+\lstnewenvironment{plaincontent}[1][] 
+ {\lstset{frame=shadowbox,escapechar=`,breaklines, breakindent=10pt, breakatwhitespace=True, linewidth=3.4in, #1}}
+{}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % Begin Abstract Here
-
+Like temperature and pressure, solution pH is an important environmental variable in biomolecular simulations. 
+Virtually all proteins depend on pH to maintain their structure and function. 
+In conventional molecular dynamics (MD) simulations of proteins, pH is implicitly accounted for by assigning and fixing protonation states of titratable sidechains. This is a significant drawback, 
+as the assigned protonation states may be wrong and they may change during dynamics.
+In this tutorial, we guide the reader in learning and using the various continuous constant pH MD methods in Amber and CHARMM packages, which have been applied to predict {\pka} values and elucidate proton-coupled conformational dynamics of a variety of proteins including enzymes and membrane transporters.
 \end{abstract}
@@
 
-\section{Introduction} % - Vini
-
-Variations in the pH condition of biological systems can lead to drastic changes in the physicochemical processes in these systems. 
-Protein folding is one example of a natural pH-dependent process, where the unfolded protein only can reach its native state in a narrow pH range \cite{Pelegrine_Gasparetto_2005_LWT-FoodScienceandTechnology}.
-Thereby, small pH changes variation in the cell can lead to misfolding diseases and even cancer proliferation and metastasis \cite{vanderKamp_Daggett_2010_BiophysicalJournal,Webb_Barber_2011_NatRevCancer,Zheng_Liu_2020_TheInternationalJournalofBiochemistry&CellBiology}.
-The protonation state of residues is also critical in drug design once the ligand-protein binding is susceptible to charge changes \cite{Dominguez_Sussman_2010_Biochemistry,Bax_Edge_2017_ActaCrystallogrDStructBiol}.
-The SARS-CoV main protease (Mpro) is an example of a protein that possesses biological activity in a narrow pH range around 7.0 \cite{Tan_Hilgenfeld_2005_JournalofMolecularBiology}.
-Out of this range, its active site can collapse by protonation changes of residues that compose it \cite{Verma_Shen_2020_J.Am.Chem.Soc.}.
-More than that, the rational design of inhibitors must consider how the ligand interactions can alter the protonation state of the titratable residues into the binding site.
-The solution pH is also critical for the study, development, and fabrication of functional dynamic materials.
-An example of this is the biopolymer of chitosan, in which its physicochemical behavior is completely different at low and high pH conditions \cite{Morrow_Shen_2015_J.Am.Chem.Soc.}.
-
-Classical molecular dynamics (MD) usually approach these pH-dependent systems by fixing the protonation states of titratable residues accordingly with experimental pKa values or pKa theoretical predictions.
-Several computational tools are currently used to predict pKa values; among the most used methods, we can cite the empirical ones PropKa and the Poisson-Boltzmann \cite{Olsson_Jensen_2011_J.Chem.TheoryComput.,Warwicker_Warwicker_2011_ProteinsStruct.Funct.Bioinforma.a}, and solver-based methods APBS-PDB 2PQR, DelPhiPKa, and H++ \cite{Unni_Baker_2011_J.Comput.Chem.,Wang_Alexov_2016_Bioinformatics,Anandakrishnan_Onufriev_2012_NucleicAcidsResearch}.
-Such tools can estimate the protonation states before running MD simulations, but they do not directly consider the dynamic flexibility of the molecules, which may lead to incorrect pKa predictions.
-Also, even assuming an accurate protonation states assignment at the beginning of MD, the conformational changes along the simulation can modify the electrostatic environment around titratable residues, affecting their protonation state.
-Thus, this relationship between conformational dynamics and the protonation states is not considered in the classical MD, and addressing this relationship is one of the most advantages of constant pH molecular dynamics (CpHMD) over conventional MD.
-
-Over the past decades, much effort has been applied to developing and improving constant pH methods \cite{Chen_Shen_2014_Mol.Simulat.}.
-We can divide the most applied methods into discrete and continuous constant pH.
-The former periodically interrupts MD and uses Metropolis Monte Carlo (MC) criterion to sample the protonation states during simulation \cite{Burgi_Gunsteren_2002_ProteinsStruct.Funct.Bioinforma.}.
-In this case, the protonation of titratable groups is settled discretely, assuming one of the two states (protonated or deprotonated).
-On the other hand, the continuous CpHMD treats the protonation states of the ionizable sites with an auxiliary set of continuous titration coordinates, using an extended Hamiltonian $\lambda$-dynamics approach \cite{Chen_Shen_2014_Mol.Simulat.}.
-This method allows the system to escape local energy minima by accessing partially protonated states temporarily.
-For years, both MC and continuous CpHMD simulations have been performed and validated by pKa calculations for several systems in different environments \cite{Chen_Shen_2014_Mol.Simulat.,Radak_Roux_2017_J.Chem.TheoryComput.}.
-Also, CpHMD has been crucial to investigate the molecular mechanisms involved in the pH dependency of protein folding \cite{MartinsdeOliveira_PereiraLeite_2018_BiophysicalJournal,Yue_Shen_2018_Phys.Chem.Chem.Phys.}, enzyme catalysis \cite{Huang_Shen_2018_J.Phys.Chem.Lett.}, conformational changes \cite{Sarkar_Roitberg_2020_J.Phys.Chem.B}, protein/ligand binding \cite{Ellis_Shen_2016_J.Phys.Chem.Lett.,Harris_Shen_2017_J.Phys.Chem.Lett.}, activation of channels and transporters \cite{Huang_Shen_2020_}, and viral uncoating \cite{Chen_Shen_2016_J.Phys.Chem.Lett.}.
-
-Our research group has exhaustively explored continuous CpHMD for different solvent models and force fields (FF).
-Using the GBNeck2 implicit solvent model implemented in Amber and hybrid-solvent based on the CHARMM platform, we accurately predicted pKa values for a large set of proteins \cite{Huang_Shen_2018_J.Chem.Inf.Model.,Harris_Shen_2019_J.Chem.Inf.Model.}.
-We also showed that these methods are reliable to predict cysteine and lysine protonation states, which are frequently used in covalent drug design \cite{Liu_Shen_2019_J.Am.Chem.Soc.,Harris_Shen_2020_J.Chem.TheoryComput.}.
-The active sites of SARS and MERS coronaviruses proteases were also investigated applying CpHMD, and we could clarify how proton-coupled conformational changes can alter their activity \cite{Verma_Shen_2020_J.Am.Chem.Soc.,Henderson_Shen_2020_J.Chem.Phys.}.
-Performing CpHMD simulations for Human Renin, we detailed its pH-dependent structure-dynamic-function relationship \cite{Ma_Shen_2020_J.Chem.Inf.Model.}, and explored the molecular mechanisms involved in inhibitors' binding and selectivity in BACE1 and $\beta$-Secretase \cite{Ellis_Shen_2015_J.Am.Chem.Soc.,Ellis_Shen_2016_J.Phys.Chem.Lett.,Harris_Shen_2017_J.Phys.Chem.Lett.}.
-Such examples are only a small sample of the vast application possibilities that CpHMD covers.
-
-\begin{table*}[]
-\begin{center}
-\begin{tabular}{lp{0.5\linewidth}c}
-System &  Topic  & CpHMD Method \\ 
-\hline
-% Add H. Line (Soluble)
-BACE1   & Investigated pH-dependent ligand binding site dynamics, small molecule inhibitor interactions, and identify proton coupled allostery  & Hybrid-Solvent \\
-BACE2   & Investigated how protonation state changes of a small molecule inhibitor affect binding with comparisons to BACE1 & Hybrid-Solvent \\
-Cat-D   & Investigated pH-dependent ligand binding site dynamics and compared them to BACE1 & Hybrid-Solvent \\
-Plasmepsin & Dissect the pH-dependent interaction of the ligand binding site, identify proton coupled allostery, elucidate the catalytic roles of the active site & Hybrid-Solvent \\
-Renin   & Explore the acid/base roles of the catalytic dyad and conformational dynamics of the ligand binding site & Hybrid-Solvent \\
-Silk Protein & & \\
-SNase   &    & \\
-Kinases &  Discovery Reactive Lys  & Implicit-Solvent \\
-PLpro   &  Identify reactive Cys residues and elucidate proton-coupled conformational dynamics  & Implicit-Solvent \\
-Mpro    & Explore protonation state dependent substrate binding site dynamics and identify reactive Cys resdiues & Implicit-Solvent \\
-% Add H. Line (Subsection: Folding)
-BBL     & & \\
-NTL9    & & \\
-% Add H. Line (Transmemb)
-NhaA    & Addressed a controversy surrounding the proton binding residues involved in the sodium proton exchange mechanism & Membrane-Enabled Hybrid-Solvent \\
-NapA    & Dissected the competitive ion binding mechanism and identified the proton binding sites. & Membrane-Enabled Hybrid-Solvent \\
-AcrB    & & Membrane-Enabled Hybrid-Solvent \\
-M2      & & Membrane-Enabled Hybrid-Solvent \\
-\end{tabular}
-\end{center}
-\caption{Protein systems that CpHMD has been used on.}
-\begin{minipage}{\columnwidth}
-Possible Minipage
-\end{minipage}
-\label{Fixed-Charge_simulations}
-\end{table*}
-
 %Here you would explain what problem you are tackling and briefly motivate your work.
-
-%In this particular template, we have removed most of the usage examples which occur in \texttt{sample-document.tex} to provide a minimal template you can modify; however, we retain a couple of examples illustrating more unusual features of our templates/article class, such as the checklists, and information on algorithms and pseudocode.
 
@@
 
-% This section gives the background and idea/purpose of CpHMD
-% Importance of pH in simulations OK
-% How Traditional MD Handles Titratable Sites OK
-% How are titratable sites treated without pHMD (PropKA, Delphi,  H++, ect...) OK
-% Idea of pHMD OK
-% Brief description of types of pHMD (DpHMD and CpHMD) OK
-% Advantages of CpHMD OK
-% Examples of CpHMD in action should range from CHARMM Hybrid-Solvent CpHMD to Amber CpHMD simulations probing for reactive Cys and Lys targets.
-
-% Figure Concept: A figure showing several different protein systems that we have investigated with CpHMD.
-
-% Topics to Cover (Might Need it's own section):
-%   1.) Brief Description of how titratable sites are handle, i.e. lambda dynamics and tautomers
-%   2.) Give an overview of the extended Hamiltonian.
-%   3.) Discuss the importance of pH-based replica exchange and its applications on CPUs and GPUs
-
-% Figure Concept: Here would be a good place to show the classic MD protocol with the CpHMD portion and we could show a small example of the replica exchange process.
-
-\subsection{Constant pH molecular dynamics}
-
-In continuous CpHMD, we define a titration coordinate, $\lambda_i$, where $i$ represents the ionizable residues \cite{Huang_Shen_2018_J.Chem.Inf.Model.}.
-We attribute to this coordinate an implicit function of an underlying coordinate to assure $\lambda_i$ remains in a range between 0 and 1.
-Thus, the $\lambda_i$ expression is given by:
-
-\begin{equation}
-    \lambda_i = \sin^2(\theta_i).
-\end{equation}
-
-When $\lambda_i = 0$, the $i$th titratable group is fully protonated, and it is fully deprotonated when $\lambda_i = 1$.
-In simulation analysis, we usually apply cutoffs for protonated states ($\lambda^P < 0.1$), and for deprotonated states ($\lambda^U > 0.9$) \cite{}.
-After defined the titration coordinate, we can use an extended Hamiltonian to allow the simultaneous propagation of spatial (real) and titration (virtual) coordinates:
-
-\begin{equation}
-\begin{split}
-\label{eq:Hamiltonian}
-\mathcal{H}(\{r_a\},\{\theta_i\})=
-&\sum_a\frac{1}{2}m_a\dot{r}_a^2+\sum_i\frac{1}{2}m_i\dot{\theta}_i^2+U^{\rm int}(\{r_a\})\\
-&+U^{\rm hybr}(\{r_a\},\{\theta_i\})+U^{\ast}(\{\theta_i\}),
-\end{split}
-\end{equation}
-where $a$ and $i$ represent the index for the spatial and titration-site coordinates, respectively.
-The two first terms of eq.
-\ref{eq:Hamiltonian} describe the kinetic energy for the atoms and the titration site.
-$U^{\rm int}$ is the titration-independent bond and non-bond energies from the non-titrating atoms.
-The fourth term, $U^{\rm hybr}$, depends on the solvent treatment.
-Herein, we will present the fully implicit solvent and the hybrid-solvent model, which $U^{\rm hybr}$ for both models includes Coulomb, vdW, and generalized Born (GB) energies and is written as
-
-\begin{equation}
-\label{eq:Uhybr}
-U^{\rm hybr}(\{r_a\},\{\theta_i\})=U^{\rm elec}(\{r_a\},\{\theta_i\})+
-U^{\rm vdW}(\{r_a\},\{\theta_i\})+U^{\rm GB}(\{r_a\},\{\theta_i\}).
-\end{equation}
-
-In eq.
-\ref{eq:Uhybr}, for the calculation of the Coulomb and GB electrostatic energies, we incorporate the atomic partial charges on the titrating residue and compute the linear interpolation between the values in the deprotonated and protonated states.
-Thus, we can write the charge on atom $j$, $q_j$, on the titrating residue $i$ as
-%
-\begin{equation}
-q_j(\lambda) = (1-\lambda_i) q_j^{prot} +  \lambda_i q_j^{unprot}.
-\label{eq:qlamb}
-\end{equation}
-
-The second term of eq. \ref{eq:Uhybr} represents the linear attenuation of the van der Waals (vdW) interactions between the titrating residue $i$ and its surrounding protein region:
-
-\begin{equation}
-U^{vdW} = (1-\lambda_i)U^{vdW}(i, p),
-\end{equation}
-where $p$ is the index of all other atom types, and $U^{vdW}$ is the protonation independent vdW interaction energy (Lennard-Jones 6-12 potential).
-
-Finally, the energy component that only affects the titratable groups is given by the last term of eq. \ref{eq:Hamiltonian} and consists of three biasing potentials, 
-\begin{equation}
-\label{eq:biasingpotential}
-U^{\ast}(\{\theta_i\})=-U^{model}(\lambda_i)+U^{barr}(\lambda_i)+U^{pH}(\lambda_i),
-\end{equation}
-
-The first term of $U^{\ast}$ describes a potential of mean force (PMF) along the $\lambda$-coordinate, considering $\lambda_i$ as an isolated model compound in the solvent,
-
-\begin{equation}
-\label{eq:Umod_1d}
-U^{model}=A(\lambda_i-B)^2,
-\end{equation}
-where $A$ and $B$ are parameters that can be determined via thermodynamic integration (TI), which are performed by MD simulations fixing several $\lambda(\theta_i)$ values, and the parameters are obtained fitting the derivative of the mean force.
-The PMF is second-order in $\lambda$ and $x$ for residues with two tautomeric states, where $x$ describes the tautomeric state.
-In this case, the two-dimensional PMF can be written as a bivariate polynomial,
-%
-\begin{equation}
-\begin{split}
-U^{mod}(\lambda_i,x_i) =
-&a_0 \lambda_i^2 x_i^2 + a_1 \lambda_i^2 x_i + a_2 \lambda_i x_i^2 + a_3 \lambda_i^2 \\
-&+ a_4 x_i^2 + a_5 \lambda_i x_i + a_6 \lambda_i + a_7 x_i + a_8,
-\end{split}\label{eq:double}
-\end{equation}
-where $x_{i}=\sin^{2}\left(\theta^{x}_{i}\right)$ where the $\theta^{x}_{i}$ are unbound titration variables that ensure that the $x_{i}$ are bound by 0 and 1.
-%
-
-$U^{barr}$, the second term of eq. \ref{eq:biasingpotential}, describes a barrier potential in the center of the titration coordinate, 
-%
-\begin{equation}
-U^{barr}(\lambda_i) = -4\beta\left(\lambda_i - \frac{1}{2} \right)^2,
-\label{eq:Ubarr}
-\end{equation}
-where $\beta$ is a parameter that determines the height of the barrier.
-Such potential increases the sampling time in the $\lambda$  coordinate endpoints and avoids unphysical vdW and electrostatic interactions associated with mixed titration state.
-
-Finally, the last term of eq. \ref{eq:biasingpotential} is given by
-%
-\begin{equation}
-U^{pH}(\lambda_i)=ln(10)k_BT(pH-\textrm{p}{K_a}^{mod})\lambda_i, 
-\label{eq:UpH}
-\end{equation}
-% 
-where $U^{pH}$ represents the pH-dependency of the deprotonation free energy, $k_B$ is the Boltzmann constant, $T$ is the system temperature, and $\textrm{p}{K_a}^{mod}$ is the model {\pka} for the titratable residue $i$.
-
-\subsection{Scope} % - Ruibin 
-This tutorial mainly covers system preparation, production run and restart, and post-processing data for the two mostly used CpHMD implementations, i.e., GBNeck2 implicit solvent based and GPU-accelerated implementation (GBNeck2-CpHMD) in the Amber pmemd.cuda platform, and the hybridge solvent and OpenMPI based implementation (Hybrid-CpHMD) in the CHARMM platform.
-For the GBNeck2-CpHMD version, we will first present how to prepare a system in explicit solvent like removing unwanted hetero atoms, ligands, ions, and waters, adding missing residues and atoms, adding dummy hydrogens, and post processing Amber parm7 file for updating exclusion list and changing GB radii.
-After that, we talk about how to pre-equilibrate the system before running CpHMD production simulations.
-For the production runs, we provide instructions on how to use conventional pH replica exchange on HPC and GPUs or asynchronous pH replica exchange on GPU clusters, and restarting existing jobs if necessary.
-For the Hybrid-CpHMD version, the system preparation process is smilar to GBNeck2-CpHMD but we mainly depend on CHARMM commands along with some mmtsb tools and we will learn how to add a pre-equilibrated water box, to miminize hydrogens with restraints and to head and equilibrate the whole system gradually.
-Instructions about how to use conventional pH replica exchange on HPC for Hybrid-CpHMD are followed as well.
-For all CpHMD simulations, choices about titratable residue types and pH conditions to use will be discussed for different use cases.
-At last, we will learn how to post process CpHMD data or the lambda files, and how to interpret the data with examples and explanations.
-We won't try to cover much traditional molecular dynamics analysis techniques but necessary solvent accessible surface area (SASA) and hydrogen bond analysis tools will be used to explain {\pka} upshifting or downshifting results.
-For all the topics, we will provide open-sourced bash or python scripts we used daily in our research.
-% This section needs to describe what is covered in this review.
-% Introduce CpHMD in the AMBER and CHARMM Packages, State the Types of CpHMD in AMBER and CHARMM, i.e. AMBER: Async pHREX with GBNeck2 and PME on GPUs, CHARMM: pHREX Hybrid-Solvent CpHMD on CPUs
-% 1.) How to prepare a system for CpHMD
-% 2.) Running CpHMD in CHARMM Equilibration to Production
-% 3.) Running CpHMD in AMBER Equilibration to Production
-% 4.) Analyzing Lambda Files, all the classic lambda values analysis procedures.
-
-%Tutorials should endeavor to cover the specific task at hand, and also highlight how the steps might need to be modified (or additional care might need to be taken at particular points) to handle more general cases.
-
-%The scope of the tutorial, as well as the expected proficiencies / outcomes for researchers who complete the tutorial, should be clearly defined.
-%This will often happen in a specific section or subsection in the article itself.
-
-We also have a GBSW implicit solvent based and a PME/GRF all-atom based implemenations in the CHARMM platform, but users are not encouraged to use them for speed issues.
-We are currently developing a GPU-accelerated all-atom implementation in the Amber pmemd.cuda platform, and the readers are welcomed to contact us for usage when it is finally finished.
-
-\section{Prerequisites} % - Ruibin
-CpHMD is based on classical molecular dynamics (MD) simulations, so it is a requirement that the user has some knowledge of MD and experience in running some MD jobs, especially in the Amber or CHARMM platform.
-Users with experience in other platforms like Gromacs, NAMD, or OpenMM are encouraged to go through some introductory tutorials for Amber or CHARMM to get started.
-For running a CpHMD simulation, it is important to know your system and research goal.
-If the user wants to know the {\pka} values for protein residues or their protonation states at certain pH conditions, the GBNeck2-CpHMD version is the best choice and Amber experience is required.
-If the user wants to investigate detailed proton-coupled conformational transormation, ion transfer, or ligand-protein binding/unbinding, especially when membrane system is involved, the Hybrid-CpHMD is currently the only choice and CHARMM experience is necessary.
-Like all computational chemistry calculations, it is important to know your system as much as possible before any serious simulations.
-With them, we can decide what salt concentration, temperature, pH range, and titratable residue types to use in CpHMD.
-Also, an expectation of time scale of the interested biological process is important to decide how long is required to run the CpHMD job.
-% Who is the tutorial designed for, beginners or people with MD experience?
-% What knoweledge should the user have prior to running CpHMD?
-% Why should a person choose to run a simulation with CpHMD
-% What should a person know before running a CpHMD simulation, pH-range, important titratable site... ect..
-
-%Here you would identify prerequisites/background knowledge that are assumed by your work, as well as any software/license requirements.
-
-\subsection{Background knowledge}
-Knowledge of classical MD is required to understand what we can and cannot achieve through a MD simulation.
-Concepts in MD like force field, equation of motion, time step, convergence, thermostat, barostat, and so on are important to understand why we need them and the parameters we choose in our minimization, heating, equilibrating, and production runs.
-Of course, chemistry knowledge of pH and its significance in chemical, material, or biological systems are essential in entering this field.
-
-Both the GBNeck2-CpHMD in Amber and Hybrid-CpHMD in CHARMM are primarily for Linux especially for HPC environments.
-The users are expected to have a good knowledge to use Linux command line tools for accessing and editing files, submitting jobs to servers, and retrieving simulation result files from remote Linux machines.
-Bash and/or Python scripting skills are required for pre- and post- processing files.
-For HPC environments, it is required to know the specific job submission and management system like SLURM or SGE installed in the HPCs.
-For GPU platform, knowledge of advantages of GPU acceleration in molecular dynamics and using tools such as 'nvidia-smi' to check GPU status and 'nvcc --version' to check CUDA version is necessary as well.
-The user can use any data visualisation tool(s), but matplotlib in Python and Jupyter notebooks is the first choice in our tool sets.
-% Tutorials should clearly define what concepts or abilities researchers will need to complete the tutorial (e.g., some proficiency in Python; experience with Jupyter notebooks; knowledge of classical MD; etc).
-
-\subsection{Software/system requirements} % - All
-% CHARMM Package Version, number of processors recommended for peak performance 
-% AMBER Package version, Recommended GPUs
-For GBNeck2-CpHMD in Amber, Amber 16 is the minimum version.
-Requirements for installing Amber can be found in the \href{https://ambermd.org/Installation.php}{Amber website}.
-For the GPU version, or specifically pmemd.cuda, the Nvidia GPU cards with at least compute compatibility sm3.0 should be installed with the latest driver.
-CUDA 10.1 is the primary CUDA version for running it but early versions from CUDA 7.5 also works.
-After successfully installing CUDA, it is important to set three environmental variables in Linux.
-CUDA\_HOME and LD\_LIBRARY\_PATH must point to the directories where the desired CUDA version is installed.
-We also need to specify CUDA\_VISIBLE\_DEVICES to GPU device indices.
-For example,
-\begin{lstlisting}
-export CUDA_HOME=/usr/local/cuda-10.1
-export LD_LIBRARY_PATH=\
-            /usr/local/cuda-10.1/lib64
-export CUDA_VISIBLE_DEVICES=0,1
-\end{lstlisting}
-will direct the Linux system to use the CUDA 10.1 version installed in the /usr/local directory and GPU with indices 0 and 1 will be used in calculations.
-After installing Amber, we should set the AMBERHOME environmental variable as the Amber package installation folder.
-\begin{lstlisting}
-export AMBERHOME=/path/to/Amber
-\end{lstlisting}
-
-To prepare systems for CpHMD simulations, we usually use \href{http://blue11.bch.msu.edu/mmtsb/Main_Page}{mmtsb tool set} for CHARMM and our \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep}{cphmd-prep} tools for Amber.
-We also use a script-level \href{https://gitlab.com/shenlab-amber-cphmd/async_ph_replica_exchange}{asynchronous replica exchange} implementation in our lab or GPU pH replica exchange simulations.
-Scripts in the \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-analysis}{cphmd-analysis} are used for analyzing CpHMD lambda files, including convergence checking, {\pka} calculation, and replica profiles.
-Some simple trajectory based analysis like RMSD can also be done but the users can use any tools from cpptraj, pytraj, vmd, mdanalysis, and more.
-Note that these free tools usually depends on other free softwares, including OpenMM, Ambertools, scipy, pandas, matplotlib, etc.
-The instructions on how to install those softwares are covered in their individual pages and not discussed here.
-A general suggestion to use Python depending packages is to use Anaconda or Miniconda.
-With them, the user can create an environment specific for CpHMD related calculations.
-For example, after installing Anaconda or Miniconda, we can type
-\begin{lstlisting}
-conda create --name cphmd python=3.7
-\end{lstlisting}
-Note that we choose Python version 3.7 because the OpenMM package currently only supports up to Python 3.7. Activate the environment named 'cphmd' by
-\begin{lstlisting}
-conda activate cphmd
-\end{lstlisting}
-Many packages can be installed through the conda and pip tool.
-\begin{lstlisting}
-conda install -c omnia pdbfixer parmed
-conda install -c conda-forge ambertools \
-                 mdanalysis matplotlib 
-conda install -c anaconda numpy scipy pandas 
-pip install f90nml
-\end{lstlisting}
-Whenever the user wants to do some calculations or analysis related to CpHMD, they can activate this 'cphmd' environment and use the installed tools without conflicting with other software. 
-% Python Packages and Version required for Async-Replica Exchange 
-% Python Packages and Version required for the CpHMD-Analysis Lambda Parser
-
-%Tutorials should clearly define what system and/or software requirements the researcher will need to complete the tutorial (e.g., VMD version 1.9 or newer, AMBER, etc.).
-Tutorials requiring specific software packages must provide instructions and files for the referenced version of the software.
-
-% \section{Content and links}
-
-% A tutorial will normally draw on additional files and materials; clearly indicate where and how these are available, with links, and how they are being archived for the long-term and maintained so they stay current.
-% You will likely want to reference your GitHub repository as a central point to access all of this information, and then the GitHub repository may link out to other content as needed.
-
-% \section{Checklists}
-% Tutorials do not necessarily require the use of a checklist as in Best Practices documents; however, they can include these if desired.
-% Several useful checklist formats are available, with examples presented in \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates}.
-% One example is shown here.
-
-\section{Workflow for CpHMD simulations} % - Jack/Ruibin
-% State an outline of the steps necessary for running the CpHMD simulations in general.
-
-% Figure concept: Show a diagram of the steps needed to run a CpHMD simulation and images representing those steps.
-
- %--------------------------------------------- 
-\begin{figure*}
-    \centering
-    \includegraphics[width=6.6in]{figs/workflow_scheme.pdf}
-    \caption{\textbf{Generalized CpHMD workflow} 
-    *Mock Figure*  
-    }
-    \label{fig:intro}
-\end{figure*}
- %---------------------------------------------
-
-\subsection{Hybrid-Solvent CpHMD in CHARMM} % - Jack
-
-Here, the steps and processes involved in running CpHMD simulations in CHARMM are presented. 
-In the first part of each section (System Preparation, Equilibration, and Production) an example of how to run CpHMD on the small soluble protein BBL (PDBid: 1w4h)\cite{Ferguson_JMolBiol_2005_v353_p427} is used as an example case.
-To follow along the please download the CpHMD tutorial from our \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/pH-REX_CpHMD_BBL_tutorial_in_CHARMM}{GitLab page}.
-In the second part of the system preparation and equilibration section, supplemental information is provided discussing how to setup and run CpHMD simulation on transmembrane proteins.
-Example input files for select portions of this process can be found from our \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/MEMB_pH-REX_CpHMD_CHARMM_Examples}{GitLab page}.
-
-\subsubsection{System Preparation}
-
-% Figure Concept: Show image of the ASP in the double protonated (syn-Configuration) State
-
-To begin one needs to obtain a protein structure from the \href{https://www.rcsb.org/}{Protein Data Bank} (PDB).\cite{Berman_NucleicAcidsRes_2000_v28_p235}
-If there are any missing residues in the downloaded protein structure they need to be filled in.
-I good tool for this is SWISS-MODEL\cite{Waterhouse_NucleicAcidsRes_2018_v46_p296}, which allows you to replace missing residues based on homology to like proteins.
-In the first part of this section we will go through an example for preparing the small protein BBL (PDBid: 1w4h)\cite{Ferguson_JMolBiol_2005_v353_p427} and in the second part discuss the preparation of transmembrane proteins.
-
-With the protein structure for BBL downloaded from PDB we clean the structure using the MMTSB\cite{Feig_JMolGraphModel_2004_v22_p377} tool \href{http://blue11.bch.msu.edu/mmtsb/convpdb.pl}{convpdb.pl}.
-Additionally, all histidine residue names need to be changed to HSP and this is done in one step with the following command. 
-%
-\begin{lstlisting}[language=bash]
-$ perl convpdb.pl -out charmm22 -segnames
--renumber 1 1w4h.pdb | sed "s/HSD/HSP/g" 
-> start.pdb
-\end{lstlisting}
-%
-Here, the output is formatted for the CHARMM22 protein force fields (-out charmm22), the segment names are included in the output (-segnames), the resids are renumbered starting at 1, and the input file is "1w4h.pdb."
-Changing all HSDs to HSPs allows for titration of both $\delta$ and $\epsilon$ protons later.
-Although the resids have been renumbered starting from one this does not have to be the case. 
-Next we add all the missing hydrogen atoms to the protein using the HBUILD facility\cite{Brunger_Proteins_1988_v4_p148} in CHARMM.
-For ASP and GLU, dummy protons are placed in the $syn$ position, and the force fields contains a dihedral energy barrier (raised to 3 kcal/mol) to maintain them in this position, so they don't move to the less favorable and rarely sampled $anti$ position ("ghost" position).\cite{Khandogin_BiophysJ_2005_v89_p141, Mongan_JComputChem_2004_v25_p2038}
-Addition of the hydrogens requires two rounds of energy minimization: first, with 50 steps of steepest decent (SD), and second 10 steps of adopted basis Newton-Raphson (ABNR) with the a heavy atom harmonically restrained (50 kcal/mol) in both cases.
-This will help remove any unfavorable hydrogen positions. 
-In most cases the N- and C-terminus ends of the protein are truncated and require capping. 
-For the N-terminus use the CH$_3$CO (patch ACE) and for the C-terminus use NH$_2$ (patch CT2) or NHCH$_3$ (patch CT3).
-In special cases the free ionized form, NH$_{3}^{+}$ and COO$^+$, is used and can be set to be titratable when CpHMD is turned on. 
-The input file "step1\_add\_h.inp" is an example of how to perform these steps on BBL and if CHARMM is installed can be run with the following command.
-%
-\begin{lstlisting}[language=bash]
-$ charmm -i step1_add_h.inp -o step1.out
-\end{lstlisting}
-%
-Additionally, the example script will check for disulfide bonds using a sulfur to sulfur atom distance of 2.5 \mbox{\AA}. 
-After running any input file make sure to visualize the output to confirm all protonation states and additional features are correctly constructed. 
-Now We move on to making a water box to solvate the system using a small box of pre-equilibrated TIP3P waters.
-This is done by generating a plan of water boxes along the X- and Y-axis, then stacking planar water boxes along the Z-axis.
-After that, the box is reshaped to a truncated octahedron. 
-In the example script the water is placed such that $\sim$10$\mbox{\AA}$ of padding is between the protein and the edge of the periodic boundary condition.
-Use the following command and input file to construct the water box.
-%
-\begin{lstlisting}[language=bash]
-$ charmm -i step2_make_waterbox.inp -o step2.out
-\end{lstlisting}
-%
-In the final step, the protein is solvated with the water box by placing the protein at the center of the box and removing all water molecules within 2.8 $\mbox{\AA}$ of the protein heavy atoms. 
-This step involves several phases of minimization; first the protein heavy atoms are fixed and the system is minimized using both SB and ABNR for 50 steps each, then five stage of minimization are used with reducing harmonic restraints (100, 50, 25, 5, 0 kcal/mol) on the protein heavy atoms using 50 steps of SD and 100 steps of ABNR each stage.
-Run this step with the following command and input file.
-%
-\begin{lstlisting}[language=bash]
-$ charmm -i step3_solvate.inp -o step3.out
-\end{lstlisting}
-%
-With the system built you can now move on to the next step. 
-Again, always remember to visualize the system and make sure everything has been constructed to your expectations.
-
-In some cases the protein crystal structure will include water molecular resolved in the PDB. 
-This can be kept with a few modifications to the protocol above. 
-First move all the water molecules into a separate PDB file and make sure the PDB file is properly formatted with the correct names, resnames, and has the correct column spacing.
-Most error messages involving the handling of the crystal waters comes from the separate crystal water PDB not being properly formatted. 
-You will need to make a new input file to add hydrogens to the water oxygen positions, like what was done with the protein and save the new water molecules as a CRD and PSF file. 
-Now you will need to make another input file that reads in the protein coordinates and the crystal water coordinates and outputs a file with them merged together. 
-The rest of the process now is similar to what was done above. 
-You just need to make sure that in step3 solvate step the you don't accidentally delete these waters.
-
-\noindent
-\textbf{Transmembrane protein system preparation details}
-
-Transmembrane proteins require a few additional steps to their preparation because they require a more rigorous equilibration. 
-First we obtain the protein structure from the database \href{https://opm.phar.umich.edu/}{Orientations of Proteins in Membranes} (OPM). 
-The OPM database orients the protein in the lipid bilayer using theoretical calculations with comparisons to experimental data, so an accurate placement of the protein in a lipid bilayer can be achieved.\cite{Lomize_Bioinformatics_2006_v22_p623}
-Just like for soluble proteins any missing residues need to be added and then the missing hydrogens can be added.
-Before the transmembrane protein system can be simulated with CpHMD, the lipid bilayer needs to be equilibrated using fixed charge MD simulations and this is done using any scalable MD engine like NAMD,\cite{Phillips_JComputChem_2005_v26_p1781} OpenMM,\cite{Eastman_PLoSComputBiol_2017_v13_pe1005659} or GROMACS.\cite{Abraham_SoftwareX_2015_v1-2_p19}
-Thus, the protonation states of titratable groups need to be assigned based on any "Anticipated" pH condition described by wither experimental of physiological pH conditions. 
-If there is not a pre-described pH condition the physiological pH (7.4) condition can be used giving the titratable residues the resulting charge states: Asp(-), Glu(-), Arg(+), Lys(+), Cys(0). 
-Since His has a {\pka} near this pH condition special attention should be given to its protonation (singly or doubly protonated) and tautomeric state (HSD or HSE), based on any hydrogen bonding or salt-bridge interactions it can have.
-Like the soluble proteins the N- and C-terminus ends should be capped and if charged caps are used their protonation states should be considered and assigned. 
-This is especially so for N-amino group since its solution {\pka} is 8.5,\cite{Nozaki_MethodsEnzymol_1967_v11_p715} only one pH unit above physiological pH.
-Once hydrogens have been added and protonation states assigned, a short minimization is required to remove any poorly placed hydrogen positions.
-Again, the system can be minimized in two round with 50 steps of SD followed by 10 steps of ABNR first with the heavy atoms either fixed or greatly harmonically restraint in the first round and with a reduced restraint in the second round.
-With the protein prepared the rest of the system can be built and partially equilibrated to relax the systems structures (protein, lipids, ect.) following the CHARMM-GUI protocol for placing the lipids, solvating the system, and adding ions to achieve a solution salt concentration and to charge neutralize the system.\cite{Jo_PLoSONE_2007_v2_pe880}
-
-\subsubsection{Equilibration}
-
-\begin{checklist}{Equilibration Requirements}
-\textbf{A list of files needed for the equilibration of a soluble protein}
-\begin{itemize}
-\item system PDB file 
-\item system CRD file 
-\item system PBC file
-\item equilibration$\_$hphmd.inp
-\end{itemize}
-\end{checklist}
-
-Equilibration of soluble protein systems for CpHMD is performed in several stages carried out at a single pH.
-Since, CpHMD is turned on in the equilibration stages the residues allowed to titrate must be chosen, for this example Asp, Glu, and His are allowed to titrate, because of the pH range used during the production phase.
-The pH chosen is normally the crystal pH condition of the PDB structure or a suitable pH for your system like physiological pH 7.4.
-Before starting the simulations determine implicit solvent ionic strength (0.15 M in the example) and temperature (298 K in the example).
-To use CpHMD in CHARMM add the following \href{https://hpc.nih.gov/apps/charmm/c42b2html/phmd.html}{PHMD module} to your input file after streaming in the GBSW parameters and reading the CpHMD parameters.
-%
-\begin{lstlisting}
-PHMD PAR 23 wri 25 PH 7.4 NPRI 500 PHFRQ 5 
-    MASS 10.0 BARR 0.0 BETA 5.0 TEMP 300 LAM -
-    sele resn ASP .or. resn GLU .or. resn HSP end
-\end{lstlisting}
-%
-Here, PH sets the pH environment of the simulation, NPRI sets the frequency to print the lambda values, PHFRQ sets the update frequency of the lambda file, TEMP sets the temperature for the configuration of the titration coordinates and should be the same as the simulation temperature, MASS is the fictitious mass of the titration coordinate, BARR is the quadratic barrier height for each titration coordinate, BETA is the friction coefficient, and LAM tells the PHMD module to output lambda values.
-The last line in the module states which residues should be allowed to titrate. 
-The length of each stage depends on the system and should be adjusted accordingly, stages lasting 40 to 120 ps have been used by group for previous projects.
-For the BBL example all heating/equilibration stages were run for 20 ps, since the system is small and stable. 
-Firstly the system is heated, here we use an NVT ensemble and set and initial temperature to what was used to resolve the either the X-ray crystal or the cryo-EM structure. 
-If the system does not require heating the initial and final temperatures can be set to be the same value and the system can equilibrate the water positions under the NVT ensemble.
-Next we run a series of NPT ensemble equilibration stages with diminishing harmonic constraint on the heavy atom positions. 
-The idea here is to slowly relax the system removing any high energy interactions and to allow for the initialization protonation states.
-Using four equilibration stages the harmonic constraints on the protein heavy atoms are reduced with the following values 5.0, 1.0, 0.1 (kcal/mol), and then in the last stage all harmonic restraints are released.
-After each equilibration make sure the energies of the system have equilibrated.
-You are now ready to start the production run. 
-The example equilibration input provided moves through each heating/equilibration stage in one input file (equilibration$\_$hphmd.inp).
-
-\noindent
-\textbf{Transmembrane protein system preparation details}
-
-The equilibration of transmembrane proteins is a little more involved than soluble proteins, because of the equilibration of the lipid membrane. 
-The transmembrane protein system has now undergone an initial relaxation according to the CHARMM-GUI protocol, and as mentioned in the system preparation fixed charge simulations need to be conducted to equilibrate the lipid bilayer.
-Here, the protein heavy atoms are harmonically restrained using a 1.0 kcal/mol/$\mbox{\AA}^2$ until the lipid bilayer is equilibrating. 
-Initially, the simulation is briefly simulated with a 0.2 ns NVT ensemble simulation.
-Then the a much longer NPT ensemble simulation is run to fully equilibrate the simulation. 
-These simulations should be run until the surface area per lipid, bilayer thickness, and lipid order parameters are near or within error limits of experimental lipid composition you are simulating.\cite{Huang_Shen_2020_}
-These three lipid bilayer metrics are frequently used to assess the phase behaviour and fine structure of the simulated bilayer.\cite{Klauda_JPhysChemB_2010_v114_p7830}
-Simulations up to 100 ns can be necessary for proper equilibration of the lipid bilayer.
-
-Following equilibration of the lipid bilayer the system can now be equilibrated at a single pH with CpHMD. 
-To start this processes, the dummy protons need to be added back on to the all Asp, Glu, and His side chains. 
-This can be done in the exact same way as it was done in the system preparation of soluble proteins and should also be minimized in a similar way.
-Now three short equilibrations can be run using CpHMD at a single pH, example inputs for these three equilibration are provided on our GitLab site [Add GitLab Information]. 
-Like a soluble protein the pH used should be the crystal pH, physiological pH, or a experimentally relevant pH.
-The first two simulations are performed with the protein harmonically restrained (1.0 kcal/mol/$\mbox{\AA}^2$) for 0.1 ns, but the first simulation uses the NVT ensemble and the second uses the NPT ensemble.
-The third simulations continues with an NPT ensemble for 0.8 ns with all restraints removed. 
-For membrane-enable CpHMD simulations the tmemb features needs to be added to the \href{https://hpc.nih.gov/apps/charmm/c42b2html/gbsw.html}{GBSW module} in the CHARMM input file. 
-%
-\begin{lstlisting}
-gbsw hybrid sgamma 0.0 nang 50 conc 0.15 temp 310 
-     tmemb 30 msw 2.5 rcyln 15                      
-     sele SOLU end
-\end{lstlisting}
-%
-Here, sgamma is the non-polar surface tension coefficient, nang is the number of angular integration points, and conc is the salt concentration.
-The inputs variables tmemb, msw, and rcyln are used to build a low dielectric slab to represent the lipid bilayer in the GBSW model with a high dielectric hole centered around the protein.
-The tmemb variable sets the slab thickness ($\mbox{\AA}$) and should be set to the average difference in the $Z$ positions of the lipid C2 atoms from the top and bottom leaflets, this will vary for different lipid types.
-The msw variable is the half membrane switching length\cite{Im_BiophysJ_2003_v85_p2900} for the dielectric transition region between the low-dielectric slab and the bulk solution, and is commonly set to 2.5 $\mbox{\AA}$.
-Finally, rcyln is the radius of the high dielectric hole in the low dielectric slab and should set so there is minimal overlap between the protein and lipids.
-The protein selection, the last line on in the GBSW module, required the protein selection to place the hole at the center of the protein.
-Without restraints on the protein heavy atoms, a cylindrical restraint with a harmonic force constant (0.1 kcal/mol/$mbox{\AA}$) needs to be applied to the protein to prevent lateral drift of the protein.
-This can be done using the MMFP facility in CHARMM.
-
-\subsubsection{Production}
-
-\begin{checklist}{Production Run Requirements}
-\textbf{A list of files needed for the production run of a soluble protein}
-\begin{itemize}
-\item system PSF file
-\item system CRD file 
-\item system PBC file
-\item final equilibration RST file
-\item production$\_$hphrex.inp
-\item rep.cmd
-\end{itemize}
-\end{checklist}
-
-The CpHMD production stage is the same for both soluble and transmembrane proteins. 
-The only differences are in the set up of the GBSW module to include a low dielectric slab to represent the lipid bilayer for transmembrane simulations, which is discussed in the equilibration stage. 
-To begin the production run one must chose a pH range to simulation over and how many 
-pH replicas to use. 
-For instance, a protein of interest may be active over a particular pH range or you might want to capture conformational dynamics that include specific titratable residues such as Asp or His. 
-Next you need to decide how many pH replicas to use for your production run. 
-Somethings to consider are how many CPUs you have available and how many CPUs you want to allocate to each replica. 
-Simulations with CHARMM do not scale well past 8 CPUs, so for a job using 16 replicas you will need 128 CPUs. 
-The spacing of the pH replicas needs to be close enough to obtain sufficient exchange probabilities (e.g., 20$\%$).
-Typically, to obtain overlap between potential energy distribution of neighboring replicas a pH spacing in increments of 0.25 or 0.5 is required. 
-Exchange attempts are made every 500 to 1000 steps, and a typical CpHMD simulation of this type will have an average exchange rate of $\sim$40$\%$, if exchange rates are too low more replicas can be added to fill in the spacing between replicas.
-To achieve enhanced sampling several replicas show move across all pH replicas.
-To use pH-based replica exchange in CHARMM the following \href{https://hpc.nih.gov/apps/charmm/c39b2html/repdstr.html}{REPD} module must be used.
-%
-\begin{lstlisting}
-REPD NREP 8 PHMD EXLM FREQ 1000 UNIT 17
-\end{lstlisting}
-%
-Here, the REPD command initiates replica exchange, NREP states the number of replicas to use, in this case 8, and PHMD states this is pH-based replica exchange and allows for the exchange of $\lambda$ along with the spatial coordinates.
-Next, EXLM allows for Hamiltonian exchange and FREQ is how often an exchange is attempted. 
-Then to run several replicas in parallel CHARMM reads in a rep.cmd file for each pH replica with the syntax, rep.cmd$\_$0, rep.cmd$\_$1, to rep.cmd$\_$(NREPS), using the following command in the input file.
-%
-\begin{lstlisting}
-stream rep.cmd
-\end{lstlisting}
-% 
-Notice this command lacks the '$\_0$' portion of the name.
-In each of the rep.cmd file you will have the command that states the pH environment of the at replica.
-%
-\begin{lstlisting}
-PHMD RESPH OLDPH @crysph NEWPH @ph PKATEMP @temp
-\end{lstlisting}
-%
-Where OLDPH is the pH used in the equilibration and NEWPH is the pH of the replica.
-finally, PKATEMP should be set to the same temperature of the simulation. 
-The rest of the replica exchange set up follows a standard molecular dynamics run in CHARMM.
-Production run inputs for BBL (production$\_$hphrex.inp and rep.cmd) can be used to run one 2 ns stage of pH-Replica Exchange CpHMD.
-
-\subsubsection{General MD and CpHMD Settings}
-% CHARMM Specific
-% Things to Include: 
-%   -) FFs
-%   -) Thermostats
-%   -) Baristas
-%   -) Cutoffs
-%   -) ect...
-
-For all simulations with CHARMM the protein is represented with CHARMM22/CMAP force fields.\cite{MacKerell_JPhysChemB_1998_v102_p3586,MacKerell_JComputChem_2004_v25_p1400}
-For the solvent the CHARMM modified TIP3P water model are used and for any ions or lipid molecules that might be included the CHARMM36 force fields are used.\cite{Klauda_JPhysChemB_2010_v114_p7830}
-To be consistent with the CHARMM36 development protocol all van der Waals energies are calculated with a switching function from 10 to 12 $\mbox{\AA}$ with a cutoff at 14 $mbox{\AA}$. 
-All simulations are conducted under periodic boundary conditions. 
-The PME method\cite{Darden_JChemPhys_1993_v98_p10089} is used to calculate the long-range electrostatic energies and forces with a real space calculation cutoff of 12 $\mbox{\AA}$, 1 $\mbox{\AA}$ grid spacing, and a 6th-order spline interpolation.
-In CHARMM the non-bond neighbor list is updated heuristically. 
-Additionally, the SHAKE\cite{Ryckaert_JComputPhys_1977_v23_p327} algorithm in CHARMM is used to constrain all bonds involving hydrogens to allow for a 2 fs time step. 
-For NPT simulations the temperature is maintained with the N\'{o}se-Hoover thermostat.\cite{Nose_MolPhys_1984_v52_p255, Hoover_PhysRevA_1985_v31_p1695}
-The pressure is maintained with the CHARMM implemented Langevin piston pressure-coupling algorithm.\cite{Feller_JChemPhys_1995_v103_p4613}
-
-For CpHMD simulations, the default GBSW radii\cite{Nina_JPhysChemB_1997_v101_p5239, Chen_JAmChemSoc_2006_v128_p3728} are used for the proteins and any small molecules the generic atomic radii in the GBSW radii file or van der Waals radii in the force field\cite{Knight_JComputChem_2011_v32_p2909} should be used.
-As shown in the example inputs the titration ($\lambda$) and tautomer ($\chi$) should be updated every 5 MD steps, allowing for water and lipid relaxation.\cite{Wallace_JChemTheoryComput_2011_v7_p2617}
-A friction coefficient of 5 ps$^{-1}$ should be used for Langevin dynamics and the mass of the titration coordinate ($\lambda$ and $\chi$) should be 10 atomic mass units.
-
-\subsection{All-atom PME CpHMD in CHARMM} % - Yandong
-
-Minor modification of the workflow for 
-hybrid-solvent CpHMD above is needed 
-to setup and run an all-atom PME CpHMD 
-simulation in CHARMM.
-
-To maintain a constant net charge during 
-all-atom PME CpHMD simulations, 
-titratable water molecules  
-or coions should be added to the 
-system as built above \cite{Wallace_Shen_2012_J.Chem.Phys.}.
-In specific, each acidic (ASP or GLU) or 
-basic residue (HSP or LYS) is 
-coupled with one coion that can convert 
-to a hydroxide (TIPU) or hydronium (TIPP) 
-\cite{Chen_Shen_2013_Biophys.J.}. 
-Together with the salt ions, the coions 
-can be placed in the water box with 
-the input script step3\_add\_ions.inp.
-Likewise, water molecules within 2.8 \AA\ 
-of ions and the protein are removed 
-with the script step4\_solvate.inp.
-
-To run an all-atom PME CpHMD simulation, 
-the GBSW module should be 
-removed first such that PME module is 
-utilized in either conformational or 
-titration dynamics \cite{Huang_Shen_2016_J.Chem.TheoryComput.}. 
-Second, the coupling information should 
-be specified. 
-As to BBL that 
-contains 3 ASP, 3 GLU and 2 HSP, below 
-are the parameters required when PHMD module 
-is invoked.
-
-%
-\begin{lstlisting}
-PHMD PAR 23 wri 25 PH 7.0 NPRI 500 PHFRQ 5 
-    MASS 10.0 BETA 5.0 TEMP 300 LAM -
-    sele resn ASP .or. resn GLU .or. TIPU .or. 
-    resn HSP .or. TIPP end - 
-    qcouple 8 - 
-    resi 4  resc 4789 -
-    resi 16 resc 4790 -
-    resi 20 resc 4791 -
-    resi 36 resc 4792 -
-    resi 37 resc 4793 -
-    resi 39 resc 4794 -
-    resi 17 resc 4771 -
-    resi 41 resc 4772 
-\end{lstlisting}
-%
-Here TIPU and TIPP are selected as titratable too. 
-The parameter qcouple, 8 in this case, indicates 
-the count of coupled pairs listed behind. 
-The second and forth 
-colunms correspond to the indices of 
-titratable residues and the coions 
-associated respectively.
-
-Finally, the calculated \pka's are 
-corrected with the following formula 
-to account for the finite-size error 
-\cite{Huang_Shen_2016_J.Chem.TheoryComput.}.
-
-\begin{eqnarray}
-\label{eq:pka_corr}
-\Delta\Delta{G}^{\rm offset} &=&
-\frac{2 \pi}{3} \kappa \gamma^{\rm solv} Q 
-\left(\frac{N}{V}
-- \rho^{\rm pure} \right) \\
-\Delta pK_a^{\rm corr} &=& \pm \frac{\Delta\Delta{G}^{\rm offset}}{ln(10)RT},
-\end{eqnarray}
-where $\gamma^{\rm solv}$ is 0.764 e$\cdot$ {\AA$^2$} 
-for TIP3P water model. 
-$\kappa$ is the electrostatic constant, 
-which is 332.0 kcal 
-$\cdot$ {\AA} $\cdot$ mol$^{-1}$e$^{-2}$. 
-The solvent number density $\rho^{\rm pure}$ is 
-0.0333679 {\AA}$^3$ for water at ambient temperature 
-and pressure. 
-Q is -1 for acidic groups and +1 for basic groups. 
-The negative sign is for acid groups and positive sign 
-is for basic groups.
-Thus, $\Delta$\pka$^{\rm corr}$ is negative for both 
-acidic and basic groups. 
-N is the number of solvent and V is the volume of the box 
-measured at equilibrium, which can be approximated 
-with the initial volume $V^{\rm init}$ times 0.93. $V^{\rm est} = L^3$ for a cubic box and $V^{\rm est} = 0.77L^3$ for a truncated octahedron box, where L denotes the cubic 
-box size. R is the Boltzmann constant and T is the temperature.
-
-\subsection{Implicit-Solvent CpHMD in AMBER} % - Ruibin
-\subsubsection{System Preparation}
-To prepare a protein system in implicit solvent for CpHMD in Amber, the process is similar to the hybrid-solvent CpHMD in CHARMM but much simpler.
-We have a \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep}{cphmd-prep} package (and soon a web server) to prepare the Amber GBNeck2-CpHMD inputs and instructions to run related CpHMD simulations by one line bash command via the 'CpHMD\_Prepare.sh' script.
-The script accepts either a PDB ID (-p|--pdbid flag) whose PDB file is downloadable from RCSB to the working machine or a user-prepared PDB file (-f|--pdbfile flag) stored in the machine.
-After that, PDBFixer (powered by OpenMM) is used to extract the desired chain (-c|--chainid flag) and model (-m|--modelid) protein structure only and remove all other parts like ions, ligands, waters, or unwanted chains.
-Missing heavy atoms and residues are then added by PDBFixer.
-Importantly, capping ligands ACE and NH$_2$ are added to cap the N and C terminus to suppress unwanted electrostatic effects from otherwise charged terminus.
-After the PDB file is prepared, tleap in Amber (or Ambertools) is used to build hydrogen and dummy hydrogen atoms according to the titratable types (-type|--residuetypes flag) set by the user.
-Specifically, the user can choose one or a combination in 'Asp Glu His Cys Lys'.
-After the system is prepared, it is required by GBNeck2-CpHMD that the GB radii of His HD1 and HE2, Cys SG, Asp OD1 and OD2, and Glu OE1 and OE2 atoms should be changed to 1.17, 2.00, 1.40, and 1.40 {\AA} respectively; and the interactions between the two dummy hydrogens in Asp and Glu should be excluded in the non-bonded interaction terms.
-This is automatically done in `CpHMD\_Prepare.sh` too.
-
-Even though our cphmd-prep tool can prepare a system for most cases, the user with different needs can use other tools to do the same or different tasks as instructed by cphmd-prep.
-Nevertheless, here we show two simple examples of how to use 'CpHMD\_Prepare.sh' to prepare a GBNeck2-CpHMD simulation system.
-The first is an EFGR kinase with provided PDB ID 5U8L and we want to know the protonation states and possible {\pka's} of His, Cys, and Lys near the psychological pH condition 7.4.
-%
-\begin{lstlisting}
-CpHMD_Prepare.sh -p 5U8L -m A -pmin 5.5
-                 -pmax 9.5 -type 'His Cys Lys'
-\end{lstlisting}
-%
-Note that we omit chain ID(s) because 5U8L only has one chain whose chain ID is A but we do specify the model ID as A because there are two alternative model IDs for 5U8L, A and B.
-We choose the pH range from 5.5 to 9.5 with a default interval 0.5 units, which covers about 2 pH units from the two directions of 7.4, making the total number of independent pH calculations to be 9.
-The other important settings are the simulation length, temperature and implicit ionic strength, which are 20 ns, 300 K, and 150 mM by default.
-Four important files will be generated if the command runs successfully.
-The 5U8L\_ModelA\_A\_fixed.rst7 file is an Amber topology rst7 file with initial coordinates; the cphmd\_5U8L\_ModelA\_A\_fixed.parm7 file is an Amber parameter parm7 file; the 5U8L\_ModelA\_A\_fixed\_phmdin file is an GBNeck2-CpHMD job control input file; the gbneck2\_input.parm file is the GBNeck2-CpHMD parameter file which is the same for all GBNeck2-CpHMD calculations.
-Other files including Amber input files are also generated and they will be explained later.
-
-The second example is to quickly estimate the {\pka's} of all Asp, Glu, His, Cys, and Lys residues in SNase with a PDB file 3BDC.pdb. The command is
-%
-\begin{lstlisting}
-CpHMD_Prepare.sh -f 3BDC.pdb -m A -pmin 1.0 
-                 -pmax 12.0 -l 4
-\end{lstlisting}
-%
-Here we feed the script with the PDB filename and use a much larger pH range from 1.0 to 12.0 so that the Asp, with a model {\pka} of 3.7, and Lys, with a model {\pka} of 10.4, residue types can both titrate in the pH range.
-In addition, we use a much shorter length, 4 ns, because we only want a rough estimation of the {\pka's}.
-We omit the -type flag because the all five residue types are included by default.
-Similar files as the previous example will be generated.
-The user may notice that many more Amber input files are generated.
-They differ in the solveph keyword that is required to perform CpHMD calculations in different pH conditions.
-
-System minimization with constraints on heavy atoms (except for two terminal and capping residues whose all atoms are allowed to move) is performed after preparing the aforementioned Amber input files.
-The constraint strength is 50 kcal/mol/\AA$^2$.
-The input file has a name end with *\_fixed\_mini.mdin which can be found in the generated files.
-The minimization last 1000 steps with the first 500 steps for steepest descent algorithm and the remaining 500 steps for conjugate gradient steps.
-After minimization, a *\_mini.rst7 is obtained for the next step.
-For convenience, all commands for minimization, equilibration, and production runs are listed in the README file generated by the 'CpHMD\_Prepare.sh' file, same for the below steps.
-
-\subsubsection{Equilibration}
-For implicit solvent based CpHMD, heating is very fast and thus not required.
-We do have four short steps of equilibration that the constraintes on heavy atoms are gradually released from 5.0, 2.0, 1.0, to 0.0 kcal/mol/\AA$^2$.
-Each step last 2000 steps or 4 ps and the total length is 16 ps.
-In equilibration and production steps, CpHMD is turned on by setting the iphmd keyword to 1 for GBNeck2-CpHMD.
-
-\subsubsection{Production}
-For production runs, we can either use CPU pH replica exchange (pHREX), GPU independent pH (pHInd), or GPU asynchronous pH replica exchange styles (\href{https://gitlab.com/shenlab-amber-cphmd/async_ph_replica_exchange}{Async-pHREX}).
-Replica exchange simulations usually need shorter time for convergence in {\pka} predictions but GPU acceleration can overcome the simulation time problem, and thus Async-pHREX is recommended over the other two styles.
-Although the asynchronous replica exchange we implemented is in the script level, it still combine the significantly good parts of GPU acceleration and replica exchange enhanced sampling, especially for large systems or the GPUs are the old models.
-We run pHREX just like we run other replica exchange simulations in Amber, i.e., we need a group file which contains the command line flags and inputs for every pH condition and then run with mpirun.
-For pHInd, we run each pH condition independently on different GPUs in the same machine by changing the CUDA\_VISIBLE\_DEVICES environment variable or even on GPUs in different machines.
-For Async-pHREX, we have to use a template input file from which temporary input files for different pH conditions at a replica exchange attempt can be generated.
-Output files are concatenated according to pH conditions instead of replicas in the pHREX method.
-For all these three methods, we can use the -rex flag to control which method is going to be used and the related commands can be found in the README file as well.
-% Here we can describe a few different styles of production run with CpHMD. 
-% this section can be broken down into three sections
-%   1.) Single-pH runs with CpHMD Simulations in AMBER
-%   2.) Async-pHREX with CpHMD simulations in AMBER
-
-\subsubsection{MD Specific Settings}
-For best performance regarding {\pka} prediction speed and accuracy, there are several settings that we think should be fixed in Amber GBNeck2-CpHMD.
-GBNeck2 implicit solvent model works best with the Amber FF14 force field.
-Langevin thermostat with a collision frequency 1 ps$-1$ is used to maintain system temperature.
-Non-bonded cutoff is 999.9 \AA which means not cutoff is used practically.
-CpHMD specific lambda files are usually saved every 1000 MD steps while trajectory files are saved every 10000 steps to save I/O time.
-% Amber Specific
-% Things to Include: 
-%   -) FFs
-%   -) Thermostats
-%   -) Baristas
-%   -) Cutoffs
-%   -) ect...
-
-\section{Lambda File Analysis} % - Jack / Vini
-
-The lambda file outputs provide us the propagation of $\lambda$ (titration state) and $x$ (tautomeric state) coordinates from titratable residues along the CpHMD simulation.
-Such information allows us to calculate the protonation state changes during the dynamics, the pKa value of each ionizable residue and also monitoring the pKa convergence.
-In order to facilitated these analysis, our research group developed a Python tool (cphmdanalysis.py) that possesses several classes to run calculations in lambda files and log files, and can be found at GitLab: \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-analysis}{CpHMD-Analysis}.
-This tool can also plot the unprotonation fraction and pKa of selected titratable residues.
-
-From each simulated pH, we can calculate the population of the $i$th titratable residue in the protonated state ($N_i^{\rm Prot}$, $\lambda_i < 0.1$) and in the unprotonated state ($N_i^{\rm Unprot}$, $\lambda_i > 0.9$).
-Thus, we can obtain the unprotonated fraction of this residue,
-
-\begin{equation}
-S_i = \frac{N_i^{\rm Unprot}}{N_i^{\rm Prot} + N_i^{\rm Unprot}}.
-\end{equation}
-
-The convergence of the $S_i$ values during the CpHMD steps indicates that the simulation reached a protonation state sampling equilibrium.
-
-Finally, we can fit the calculated $S_i$ at several pHs using the generalized form of the Henderson-Hasselbach equation,
-
-\begin{equation}
-S_i = \frac{1}{1+10^{n(\rm pK_{a,i} - pH)}},
-\end{equation}
-where $n$ is the Hill coefficient, and indicates the degree of cooperativity between the titratable groups.
-
-% Here we can describe how to analyze CpHMD results
-%   1.) Calculations of Unprot. Fractions
-%   2.) Calculations pKa  
-%   3.) How to Monitor pKa Convergence
-%   3.) Additional Analysis could include proton-coupling, ect...
-
-% Figure Concept: Here we should have a 2 panel figure. 
-%   -) Panel 1: Raw Lambda Value versus time @ a single pH with a duel y-Axis and over the plot we showing the running S-value. 
-%   -) Panel 2: We show a titration Curve
-
-\section{Parameterization} % - Jack/Ruibin
-To perform CpHMD simulations on a new titratable residue, accurate values for the parameters of $U^{model}$ are required.
-The basic method for obtaining these parameters is to simulate a model compound containing an analog of the residue in question at several fixed values of $\theta_{i}$ and $\theta^{x}_{i}$ with $U^{barr}$ and $U^{pH}$ set to 0, compute the derivatives of the free energy with respect to $\theta_{i}$ and $\theta^{x}_{i}$, and fit those derivatives to eq.
-\ref{eq:Umod_1d} and \ref{eq:double}.
-
-Note that these fittings should be performed on the $\theta$ and $\theta_{x}$ variables directly rather than on $\lambda$ and $x$ because the nonlinearity in the definitions of $\lambda$ and $x$ cause difficulties in the fitting.
-
-Typically we are interested in three different kinds of titratable residues in most proteins, residues, such as Lys, that have a single titratable site, residues, such as Glu and Asp, that have two titratable sites with identical microspcopic pKa's, and residues, such as His, that have two titration sites with different microscopic pKa's.
-Each of these cases have different forms for their $U^{model}$ and therefore require somewhat different methods.
-
-First, for residues like Lys with a single titratable site, we only need to consider a single titration variable, $\theta$.
-For these residues we construct a model compound for which the experimental pKa is known, such as the blocked amino acid, and run simulations with fixed values of $\theta$ and fit the resulting $\partial U/\partial\theta$ to eq.
-\ref{eq:Umod_1d}.
-
-Second, for residues, like His, that contain two nonidentical intertacting titratable sites, eq. \ref{eq:double} can be written as
-
-\begin{align}
-U^{model}= & A_{10}\lambda^{2}x^{2}+2\left(A_{1}B_{1}-A_{0}B_{0}\right)\lambda x+\nonumber \\ & 2\left(A_{0}B_{0}-A_{1}B_{1}-A_{10}B_{10}\right)\lambda^{2}x+ \nonumber \\ & A_{1}\lambda^{2}-2A_{1}B_{1}\lambda\label{eq:his}
-\end{align}
-
-The six parameters $A_{0}$, $A_{1}$, $B_{0}$, $B_{1}$, $A_{10}$, and $B_{10}$ can be obtained by running simulations on the model compound at different values of $\theta$ and $\theta_{x}$ with $U^{model}$, $U^{barr}$, and $U^{pH}$ set to 0, computing the derivatives of the free energy with respect to $\theta_{i}$ and $\theta^{x}_{i}$, and fitting those derivatives to this equation.
-Note once again that although eq.
-\ref{eq:his} is written in terms of $\lambda$ and $x$ for clarity, the fitting should be performed against $\theta$ and $\theta_{x}$ to avoid issues arising from nonlinearities near the endpoints.
-
-Finally, for residues like Asp and Glu, that contain two identical interacting titratable sites, eq. \ref{eq:double} reduces to
-
-\begin{align}
-    U^{model}= & \left(R_{1}\lambda^{2}+R_{2}\lambda+R_{3}\right)\left(x-R_{4}\right)^{2}+\left(R_{5}-R_{1}R_{4}^{2}\right)\lambda^{2} \nonumber\\ & -\left(2R_{5}R_{6}+R_{2}R_{4}^{2}\right)\lambda \label{eq:asp}
-\end{align}
-
-As before, the six parameters $R_{1}$ to $R_{6}$ can be obtained by running simulations on the model compound at different values of $\theta$ and $\theta_{x}$ with $U^{model}$, $U^{barr}$, and $U^{pH}$ set to 0, computing the derivatives of the free energy with respect to $\theta$ and $\theta^{x}$, and fit those derivatives to this equation.
-Note once again that although eq.
-\ref{eq:asp} is written in terms of $\lambda$ and $x$ for clarity, the fitting should be performed against $\theta$ and $\theta_{x}$ to avoid issues arising from nonlinearities near the endpoints.
-
-\subsection{Parameterization settings and fitting procedure}
-As discussed above, the key step to parameterizing a new residue is to run simulations with fixed values of $\theta$ and $x$ with $U^{barr}$, $U^{model}$, and $U^{pH}$ set to 0.
-This can be accomplished in Amber by setting the options prlam to false, prderiv to true, and phtest to true in the phmdin file and setting ph{\_}theta to the desired values of $\theta$ and $x$ and vph{\_}theta to 0 in the phmdstrt file.
-
-As discussed above, the procedure for parameterizing a residue, such as Lys, with a single titratable site is pretty straightforward, just run at several values of $\theta$ (We generally find that $\theta=\left(0.0,0.2,0.4,0.6,0.785,1.0,1.2,1.4,1.5708\right)$ work well).
-In contrast, parameterizing two site residues is more complicated because the energy surface is two-dimensional.
-What follows is therefore a description of the method we usually use to parameterize these residues, although other approaches are possible.
-
-First, for residues, such as His, with two titration sites with different microscopic pKa's, we generally run simulations at $\theta_{x}=0$ and $\theta=\left(0.0,0.2,0.4,0.6,0.785,1.0,1.2,1.4,1.5708\right)$.
-The parameters $A_{0}$ and $B_{0}$ can then be computed by fitting $\partial U/\partial\theta$ to
-
-\begin{equation}
-    \frac{\partial U}{\partial\theta}=2A_{0}\sin\left(2\theta\right)\left(\sin^{2}\left(\theta\right)-B_{0}\right).
-\end{equation}
-Next, the parameters $A_{1}$ and $B_{1}$ can be computed by running simulations at $\theta_{x}=1.5708$ and $\theta=\left(0.0,0.2,0.4,0.6,0.785,1.0,1.2,1.4,1.5708\right)$ and fitting $\partial U/\partial\theta$ to
-
-\begin{equation}
-    \frac{\partial U}{\partial\theta}=2A_{1}\sin\left(2\theta\right)\left(\sin^{2}\left(\theta\right)-B_{1}\right).
-\end{equation}
-Finally, $A_{10}$ and $B_{10}$ can be computed by running simulations at $\theta=1.5708$ and $\theta_{x}=\left(0.0,0.2,0.4,0.6,0.785,1.0,1.2,1.4,1.5708\right)$ and fitting $\partial U/\partial\theta_{x}$ to
-
-\begin{equation}
-    \frac{\partial U}{\partial\theta_{x}}=2A_{10}\sin\left(2\theta_{x}\right)\left(\sin^{2}\left(\theta_{x}\right)-B_{10}\right).
-\end{equation}
-
-Finally, for residues, such as Asp, that have two identical interacting titrating sites we have to find parameters $R_{1}$ to $R_{6}$.
-To start with, we generally run simulations at combinations of $\theta=\left(0.0,0.2,0.4,0.6,0.785,1.0,1.2,1.4\right)$ and  $\theta_{x}=\left(0.0,0.2,0.4,0.6,0.785,1.0,1.2,1.4,1.5708\right)$, compute $A\left(\theta\right)$ and $B\left(\theta\right)$ at each value of $\theta$ by fitting $\partial U/\partial\theta_{x}$ to
-
-\begin{equation}
-    \frac{\partial U}{\partial\theta_{x}}=2A\left(\theta\right)\sin\left(2\theta_{x}\right)\left(\sin^{2}\left(\theta_{x}\right)-B\left(\theta\right)\right),
-\end{equation}
-compute $A\left(\theta_{x}\right)$ and $B\left(\theta_{x}\right)$ at each value of $\theta_{x}$ by fitting $\partial U/\partial\theta_{x}$ to
-
-\begin{equation}
-    \frac{\partial U}{\partial\theta}=2A\left(\theta_{x}\right)\sin\left(2\theta\right)\left(\sin^{2}\left(\theta\right)-B\left(\theta_{x}\right)\right),
-\end{equation}
-obtain estimates of $R_{1}$, $R_{2}$, and $R_{3}$ by fitting $A\left(\theta\right)$ to
-
-\begin{equation}
-    A\left(\theta\right) = R_{1}\sin^{4}\left(\theta\right)+R_{2}\sin^{2}\left(\theta\right)+R_{3},
-\end{equation}
-setting $R_{4}=0.5$ (guaranteed because the two titrating sites are identical), obtain an estimate of $R_{5}$ by fitting $A\left(\theta_{x}\right)$ to
-
-\begin{equation}
-    A\left(\theta_{x}\right)=a_{0}\sin^{4}\left(\theta_{x}\right)+a_{1}\sin^{2}(\theta_{x})+R_{5},
-\end{equation}
-and finally obtain an estimate of $R_{6}$ by fitting $B\left(\theta_{x}\right)$ to
-
-\begin{equation}
-    B\left(\theta_{x}\right)=a_{0}\sin^{4}\left(\theta_{x}\right)+a_{1}\sin^{2}(\theta_{x})+R_{6},
-\end{equation}
-% Here is a single-column checklist that consists of multiple sub-checklists
-\begin{Checklists}
-
-\begin{checklist}{A list}
-\textbf{Single-column checklists are also straightforward by removing the asterisk}
-\begin{itemize}
-\item First thing let's do an item which breaks across lines to see how that looks
-\item Also remember
-\item And finally
-\end{itemize}
-\end{checklist}
-
-\begin{checklist}{Another list}
-\textbf{This is some further description.}
-\begin{itemize}
-\item First thing
-\item Also remember
-\item And finally
-\end{itemize}
-\end{checklist}
-
-\end{Checklists}
+\input{1_intro}
+\input{1_methods}
+\input{1_scope}
+\input{2_prereq}
+\input{3_hphmd}
+\input{4_aphmd}
+\input{5_gphmd}
+\input{6_analysis}
+\input{7_parameterization}
 
@@
 %%%%%%%%%%%%%%%%
+Henderson, Liu, Harris, Huang, de Oliveira, and Shen
+together wrote the manuscript.
+We thank Zhi Yue and Paween Mahinthichaichan for contributing example scripts for the hybrid-solvent CpHMD simulations of transmembrane proteins.
+We thank Jason A. Wallace for a bash script to calculate {\pka} values and track the convergence of unprotonated fractions using Grace.
+% % We suggest you preserve this comment:
+% For a more detailed description of author contributions,
+% see the GitHub issue tracking and changelog at \githubrepository.
 
-(Explain the contributions of the different authors here)
-
-% We suggest you preserve this comment:
-For a more detailed description of author contributions,
-see the GitHub issue tracking and changelog at \githubrepository.
-
-\section{Other Contributions}
+%\section{Other Contributions}
 %%%%%%%%%%%%%%%
@@
 
-(Explain the contributions of any non-author contributors here)
-% We suggest you preserve this comment:
-For a more detailed description of contributions from the community and others, see the GitHub issue tracking and changelog at \githubrepository.
+% (Explain the contributions of any non-author contributors here)
+% % We suggest you preserve this comment:
+% For a more detailed description of contributions from the community and others, see the GitHub issue tracking and changelog at \githubrepository.
 
@@
 %%%%%%%
-
-Declare any potentially conflicting interests here, whether or not they pose an actual conflict in your view.
+The CpHMD-related software may be of interest to ComputChem LLC, for which
+J.S. is a founder and scientific advisor. 
+J.S. also serves on the scientific advisory board of EoCys LLC.
 
@@
 %%%%%%%
-We acknowledge financial support from the National Institutes of Health (R01GM098818 and R01CA256557).
+We acknowledge financial support from the National Institutes of Health (R01GM098818 and R01CA256557 to Shen and R44GM134756 to Harris) and National Science Foundation (CBET1932963 to Shen).
 
@@
 
-\bibliography{Bib/JanaShenLab.bib}
+\bibliography{cphmd}
 

### 2022-08-08 Old/1_intro_old.tex
Revision updates

\section{Introduction} % - Vini 
\subsection{Background}
Solution pH is tightly regulated in biological environments \cite{Casey_Orlowski_2010_Nat.Rev.Mol.CellBiol.}. 
In normal adult cells, intracellular pH$_{\rm i}$ is about 7.2 and extracellular pH$_{\rm e}$ is about 7.4 \cite{Webb_Barber_2011_Nat.Rev.Cancer};
however, in cancer cells, the pH gradient is reversed,
which promotes cancer progression, metabolic adaptation, and metastasis
\cite{Webb_Barber_2011_Nat.Rev.Cancer,White_Barber_2017_J.CellSci.}. 
Virtually all proteins depend on pH to maintain their structure and function
\cite{Casey_Orlowski_2010_Nat.Rev.Mol.CellBiol.,Roos_Boron_1981_Physiol.Rev.}.
% Protein folding is one example of a pH-dependent process, where the unfolded protein only can reach its native state in a narrow pH range. 
A small change in pH can perturb the functions of many enzymes.
 %The protonation state of residues is also critical in drug design once the ligand-protein binding is susceptible to charge changes \cite{Dominguez_Sussman_2010_Biochemistry,Bax_Edge_2017_ActaCrystallogrDStructBiol}. 
For example, SARS coronavirus main proteases have the maximal
cleavage activity in a narrow pH range around 7.0 \cite{Tan_Hilgenfeld_2005_J.Mol.Biol.}. 
Below the optimum pH, the active site collapses
due to the protonation state change of a histidine \cite{Tan_Hilgenfeld_2005_J.Mol.Biol.,Verma_Shen_2020_J.Am.Chem.Soc.}. 
The function of human $\beta$-secretase 1 peaks at pH 4.5,
\cite{Shimizu_Nukina_2008_Mol.CellBiol.}; at low or high pH
the protonation state of the catalytic dyad changes, 
which results in the closure of the ``flap'' that lies above
the active site and blockage of 
substrate entrance \cite{Ellis_Shen_2015_J.Am.Chem.Soc.}.
%  the rational design of inhibitors must consider how the ligand interactions can alter the protonation state of the titratable residues into the binding site. 
Solution pH is also an important parameter in the creation of dynamic, stimuli-responsive materials \cite{Li_Payne_2019_Biofabrication}. 
For example, pH triggers the formation of hydrogel films
from the biopolymer chitosan
\cite{Morrow_Shen_2015_J.Am.Chem.Soc.}; 
addition of surfactants allows the chitosan film to be reconfigured via a second pH signal to achieve different mechanical properties \cite{Tsai_Shen_2018_Chem.Mater.}.

In a conventional molecular dynamics (MD) simulation,
solution pH is implicitly taken into account by
fixing the protonation states of titratable sites according to the experimental {\pka} values of
model compounds (or peptides) or 
traditional {\pka} calculations based on a static structure. 
Traditional {\pka} methods include empirical
approaches, e.g.,
PropKa \cite{Olsson_Jensen_2011_J.Chem.TheoryComput.},
Rosetta \cite{Kilambi_Gray_2012_Biophys.J.},
Poisson-Boltzmann solvers
MCCE \cite{Song_Gunner_2009_J.Comput.Chem.},
APBS/PDB2PQR \cite{Unni_Baker_2011_J.Comput.Chem.}, 
DelPhiPKa \cite{Wang_Alexov_2016_Bioinformatics},  
H++ \cite{Anandakrishnan_Onufriev_2012_NucleicAcidsRes.},
and other variants \cite{WarwickerJim_WarwickerJim_2011_Proteins}.
These tools are fast; however, 
the predicted protonation states are often incorrect
for deeply buried sites
\cite{Alexov_Word_2011_Proteins},
enzyme active sites \cite{Huang_Shen_2018_J.Phys.Chem.Lett.},
transmembrane proteins \cite{Huang_Shen_2016_Nat.Commun.},
and cysteines or lysines in general \cite{Liu_Shen_2019_J.Am.Chem.Soc.,Harris_Shen_2020_J.Chem.TheoryComput.,Liu_Shen_2021_J.Med.Chem.a}.
% due to a significant {\pka} shift in the protein environment relative to the solution (model) value.
% 2) Even if the assignment is correct based on the initial experimental structure, a change in protonation (or tautomer) state may occur during dynamics.. 
Also, even assuming an accurate assignment of protonation states at the beginning of MD, the conformational changes sampled
by the simulation may modify the electrostatic environment around the titratable groups, changing their protonation states. 
To account for 
the direct coupling between conformational dynamics and protonation state changes and to offer more accurate \pka\ predictions, constant pH MD methods have been developed.

Over the past two decades, much progress has been made in the
development of constant pH MD methods.
% Regarding the system modeling, the C$\alpha$ structure-based model CpHMD is one of the highest CG level approaches; such model treats amino acid residues as beads centered in $\alpha$ carbon, and titratable residues charge is also placed in the center of the bead.  
% \cite{Contessoto_Leite_2016_J.Chem.TheoryComput.}. 
% The titratable Martini is another example of CpHMD-CG, but in turn, it can treat each residue for several beads that represent a group of two to five heavy atoms \cite{Grunewald_Marrink_2020_J.Chem.Phys.}. 
% Finally, the most sophisticated models can explicitly treat all atoms of the system, but even with several ways to model the simulated systems, the CpHMD methods are grouped into discrete and continuous categories. 
% \cite{Chen_Khandogin_2008_Curr.Opin.Struct.Biol.,Chen_Shen_2014_Mol.Simulat.}. 
In a discrete (also known as stochastic) constant pH MD (DpHMD), the MD trajectory is periodically
interrupted by the Metropolis Monte-Carlo (MC) sampling of 
protonation states \cite{Baptista_Soares_2002_J.Chem.Phys.,Mongan_McCammon_2004_J.Comput.Chem.,Swails_Roitberg_2014_J.Chem.TheoryComput.,Stern_Stern_2007_J.Chem.Phys.,Chen_Roux_2015_J.Chem.TheoryComput.}. 
Here, protonation and deprotonation of titratable groups is
sampled discretely, assuming one of the two states. 
In contrast, the continuous constant pH MD (CpHMD) 
\cite{Lee_Brooks_2004_Proteins,Khandogin_Brooks_2005_Biophys.J.,Khandogin_Brooks_2006_Biochemistry,Wallace_Shen_2011_J.Chem.TheoryComput.,Wallace_Shen_2012_J.Chem.Phys.,Huang_Shen_2016_J.Chem.TheoryComput.}
treats the protonation states of ionizable sites using an auxiliary set of continuous titration coordinates based on an extended Hamiltonian $\lambda$-dynamics approach \cite{Kong_Brooks_1996_J.Chem.Phys.}. 
This method allows the system to escape local energy minima by transiently accessing the partially protonated states, which are unphysical.
Although the latter is a major caveat, the CpHMD method offers significantly faster convergence and importantly it can be implemented for implicit-, hybrid-, and fully explicit-solvent simulations.
On the contrary, although progress is being made, 
e.g., via the approach of nonequilibrium MD \cite{Stern_Stern_2007_J.Chem.Phys.,Chen_Roux_2015_J.Chem.TheoryComput.,Radak_Roux_2017_J.Chem.TheoryComput.}, 
it is very challenging to extend the Monte-Carlo based DpHMD method to fully explicit-solvent simulations
due to the low acceptance rate as a result of the 
large energy change.
%In this tutorial, we will focus on the CpHMD methods.

In the implicit-solvent constant pH scheme, sampling of both conformational dynamics and protonation states is performed in implicit solvent, typically a generalized-Born (GB) model.
Implementations of this scheme include the GBMV-
\cite{Lee_Brooks_2003_J.Comput.Chem.} and 
GBSW- \cite{Im_Brooks_2003_J.Comput.Chem.} based CpHMD \cite{Lee_Brooks_2004_Proteins,Khandogin_Brooks_2005_Biophys.J.,Khandogin_Brooks_2006_Biochemistry} in CHARMM  \cite{Brooks_Karplus_2009_J.Comput.Chem.},
and the GBNeck2-based CpHMD
\cite{Huang_Shen_2018_J.Chem.Inf.Model.,Harris_Shen_2019_J.Chem.Inf.Model.},
and the OBC-GB \cite{Onufriev_Bashford_2002_J.Comput.Chem.}
based DpHMD method 
\cite{Mongan_McCammon_2004_J.Comput.Chem.}
in Amber \cite{Case_Kollman_2018_}.
In the hybrid-solvent constant pH scheme, conformational sampling
is conducted with explicit solvent, while a continuum
electrostatic model, e.g., GB or Poisson-Boltzmann (PB)
is used to calculate the solvation forces 
on the $\lambda$ particles for the CpHMD methods
or the energy changes due to protonation-state change
for the DpHMD methods.
Implementations of this scheme include
the PB/TIP3P based stochastic DpHMD method
\cite{Baptista_Soares_2002_J.Chem.Phys.}
for GROMACS \cite{Spoel_Berendsen_2005_J.Comput.Chem.},
the GBSW/TIP3P based CpHMD method \cite{Wallace_Shen_2011_J.Chem.TheoryComput.}
in CHARMM \cite{Brooks_Karplus_2009_J.Comput.Chem.},
and the OBC-GB/TIP3P based DpHMD method 
\cite{Swails_Roitberg_2014_J.Chem.TheoryComput.}
in Amber \cite{Case_Kollman_2018_}.
Note, the GBSW/TIP3P based CpHMD method \cite{Wallace_Shen_2011_J.Chem.TheoryComput.} was extended to the treatment of transmembrane systems \cite{Huang_Shen_2016_Nat.Commun.} by incorporating the membrane GBSW model \cite{Im_Brooks_2003_Biophys.J.}.
Finally, an all-atom or fully explicit-solvent
CpHMD method was implemented in CHARMM
\cite{Brooks_Karplus_2009_J.Comput.Chem.}
as a part of the $\lambda$-dynamics functionality
in the BLOCK module (CPHMD$^{\rm MS\lambda D}$) 
by the Brooks group \cite{Goh_Brooks_2014_Proteins}
and as an extension to the GBSW and hybrid-solvent CpHMD 
methods in the PHMD module by the Shen group
\cite{Wallace_Shen_2012_J.Chem.Phys.,Huang_Shen_2016_J.Chem.TheoryComput.}.
The all-atom CpHMD method by the Shen group is the only one that includes titratable water to compensate for the net charge fluctuation as a result of proton titration in the CpHMD simulation
\cite{Chen_Shen_2013_Biophys.J.}.
An all-atom CpHMD implementation was also developed by the
Grubm\"{u}ller group
\cite{Donnini_Grubmuller_2016_J.Chem.TheoryComput.}
in GROMACS
\cite{Spoel_Berendsen_2005_J.Comput.Chem.}, although it has not been validated for a large number of proteins yet.
Currently, the only all-atom discrete constant pH method is based on nonequilibrium MD,
and it is implemented in NAMD \cite{Phillips_Schulten_2005_J.Comput.Chem.}
by the Roux group
\cite{Chen_Roux_2015_J.Chem.TheoryComput.,Radak_Roux_2017_J.Chem.TheoryComput.}.
% Below we will focus on the CpHMD implementations
% developed by the Shen group.

While accurate prediction of protonation states or {\pka} values is a goal, a major application of constant pH MD is to elucidate the mechanisms of pH-dependent or proton-coupled conformational dynamics. 
One can argue that the latter can also be studied by conventional fixed-protonation-state MD by running simulations with different protonation states.
While this may well be true when the switch of one protonation state is involved, the problem quickly becomes intractable 
when multiple titration sites are involved.
Furthermore,
the identity of the titratable site responsible for proton-coupled conformational changes is
often unclear, which makes the fixed-protonation-state approach unfeasible.
% For example,
% in the spider silk protein \cite{Wallace_Shen_2012_J.Phys.Chem.Lett.},
% the sodium-proton antiporter NhaA \cite{Huang_Shen_2016_Nat.Commun.,Henderson_Shen_2020_Proc.Natl.Acad.Sci.USA} and the proton-coupled drug efflux pump AcrB \cite{Yue_Shen_2017_J.Chem.TheoryComput.}.
Importantly, by running fixed-protonation-state MD one cannot obtain the {\pka} value for the conformational transition \cite{Wallace_Shen_2012_J.Phys.Chem.Lett.,Huang_Shen_2016_Nat.Commun.,Yue_Shen_2017_J.Chem.TheoryComput.,Henderson_Shen_2020_Proc.Natl.Acad.Sci.USA}, 
which can be compared with experiment to validate the microscopic details revealed by simulation.
% Constant pH methods have been reviewed in the past \cite{Chen_Khandogin_2008_Curr.Opin.Struct.Biol.,Wallace_Shen_2009_MethodsEnzymol.,Chen_Shen_2014_Mol.Simulat.}. In this tutorial 
% pH-dependent processes often involve the
% protonation state switches of multiple titratable sites, some of which may sample both protonated and deprotonated states.
% Examples include
% protein folding \cite{Khandogin_Brooks_2006_Proc.Natl.Acad.Sci.USA,Khandogin_Brooks_2007_Proc.Natl.Acad.Sci.USA}
% and pH-triggered self assembly of materials \cite{Morrow_Shen_2015_J.Am.Chem.Soc.}.

Table~\ref{Table:applications} summarizes the applications of the implicit-, hybrid-, and fully explicit-solvent CpHMD methods developed in the Shen group.
% Over the past decade, CpHMD simulations have been applied 
% to proteins in different environments \cite{Chen_Shen_2014_Mol.Simulat.,Radak_Roux_2017_J.Chem.TheoryComput.}. 
The implicit-solvent GBSW- and GBNeck2-CpHMD methods have been mainly applied to {\pka} predictions of proteins
\cite{Khandogin_Brooks_2006_Biochemistry,Wallace_Shen_2009_MethodsEnzymol.,Wallace_Shen_2011_Proteins,Harris_Shen_2019_J.Chem.Inf.Model.,Harris_Shen_2020_J.Chem.TheoryComput.},
while the hybrid-solvent and all-atom CpHMD methods have been additionally applied to elucidate the mechanisms of pH-dependent or proton-coupled conformational processes, e.g.,
protein unfolding
\cite{Yue_Shen_2018_Phys.Chem.Chem.Phys.},
structure-function relationships of proteases and kinases \cite{Ellis_Shen_2015_J.Am.Chem.Soc.,Tsai_Shen_2019_J.Am.Chem.Soc.,Ma_Shen_2021_J.Chem.Inf.Model.,Verma_Shen_2020_J.Am.Chem.Soc.,Henderson_Shen_2020_J.Chem.Phys.}, protein/ligand binding \cite{Ellis_Shen_2016_J.Phys.Chem.Lett.,Harris_Shen_2017_J.Phys.Chem.Lett.,Henderson_Shen_2018_J.Phys.Chem.Lett.}, 
conformational activation of transmembrane channels and transporters \cite{Chen_Shen_2016_J.Phys.Chem.Lett.,Yue_Shen_2017_J.Chem.TheoryComput.,Huang_Shen_2016_Nat.Commun.},
and materials \cite{Morrow_Shen_2015_J.Am.Chem.Soc.,Tsai_Shen_2018_Chem.Mater.}.
A recent application of GBNeck2-CpHMD method is the prediction of nucleophilic cysteine and lysine sites for targeted covalent drug design \cite{Liu_Shen_2019_J.Am.Chem.Soc.,Verma_Shen_2020_J.Am.Chem.Soc.,Henderson_Shen_2020_J.Chem.Phys.,Liu_Shen_2021_J.Med.Chem.a,Liu_Shen_2021_RSCMed.Chem.}. 
In this tutorial, we will discuss GBNeck2-CpHMD in Amber \cite{Huang_Shen_2018_J.Phys.Chem.Lett.,Harris_Shen_2019_J.Chem.Inf.Model.},
hybrid-solvent  \cite{Wallace_Shen_2011_J.Chem.TheoryComput.},
and particle-mesh Ewald (PME) all-atom CpHMD 
\cite{Huang_Shen_2016_J.Chem.TheoryComput.}
in CHARMM \cite{Brooks_Karplus_2009_J.Comput.Chem.}. 
% The GB-based CpHMD methods
% have been validated for a large number of proteins and the latter has additionally revealed the pH-dependent or proton-coupled mechanisms of various soluble and transmembrane proteins.

\begin{table*}
\begin{center}
\begin{tabular}{p{1.2in}p{3in}ll}
System &  Topic  & Method & Reference\\ 
\hline
\multicolumn{2}{l}{\bf Soluble proteins} & \\
Benchmark proteins &  {\pka} calculations & GBSW 
& \cite{Khandogin_Brooks_2006_Biochemistry,Wallace_Shen_2009_MethodsEnzymol.}\\
SNase  & Blind {\pka} predictions for 87 engineered mutant proteins   & GBSW & \cite{Wallace_Shen_2011_Proteins}\\
Peptides and mini-proteins & pH-dependent folding mechanisms & GBSW & \cite{Khandogin_Brooks_2006_Proc.Natl.Acad.Sci.USA,Khandogin_Brooks_2007_Proc.Natl.Acad.Sci.USA,Khandogin_Brooks_2007_J.Am.Chem.Soc.}\\
Benchmark proteins & {\pka} calculations & Hybrid & \cite{Wallace_Shen_2011_J.Chem.TheoryComput.}\\
NTL9, BBL & Unfolded states and unfolding mechanism & Hybrid& \cite{Shen_Shen_2010_J.Am.Chem.Soc.,Yue_Shen_2018_Phys.Chem.Chem.Phys.}\\
SNase   & {\pka} calculation for a deeply buried lysine; proton-coupled opening of the site   & Hybrid & \cite{Shi_Shen_2012_Biophys.J.}\\
Spider silk protein & pH sensing residues for dimerization & Hybrid & \cite{Wallace_Shen_2012_J.Phys.Chem.Lett.}\\
BACE1, BACE2, cathepsin D, renin, plasmepsin D & Acid/base roles of the catalytic dyad; binding-induced protonation state change of the inhibitor; pH-dependent conformational dynamics;
proton-coupled dynamics of binding site water;
pH-dependent 
inhibitor binding free energy calculations & Hybrid & \cite{Ellis_Shen_2015_J.Am.Chem.Soc.,Ellis_Shen_2016_J.Phys.Chem.Lett.,Ellis_Shen_2017_J.Comput.Chem.,Harris_Shen_2017_J.Phys.Chem.Lett.,Henderson_Shen_2018_J.Phys.Chem.Lett.,Ma_Shen_2021_J.Chem.Inf.Model.,Henderson_Shen_2021_J.Chem.Inf.Model.}\\
c-Src Kinase & Proton-coupled conformational change of the DFG motif & Hybrid & \cite{Tsai_Shen_2019_J.Am.Chem.Soc.}\\
% BACE1, cathepsin D &  pH-dependent binding free energy calculations & Hybrid & \cite{Harris_Shen_2017_J.Phys.Chem.Lett.}\\
Benchmark proteins & {\pka} calculations for Asp/Glu/His &  All-atom & \cite{Huang_Shen_2016_J.Chem.TheoryComput.} \\
% Plasmepsin & Dissect the pH-dependent interaction of the ligand binding site, identify proton coupled allostery, elucidate the catalytic roles of the active site & Hybrid-Solvent \\
Various enzymes & Prediction of proton donor and nucleophile and physical determinants & Hybrid & \cite{Huang_Shen_2018_J.Phys.Chem.Lett.}\\
Benchmark proteins & {\pka} calculations for Asp/Glu/His/Cys &  GBNeck2 & \cite{Huang_Shen_2018_J.Chem.Inf.Model.,Harris_Shen_2019_J.Chem.Inf.Model.}\\
Various kinases &  Prediction and rationalization of reactive Lys and Cys  & GBNeck2 & \cite{Liu_Shen_2019_J.Am.Chem.Soc.,Liu_Shen_2021_J.Med.Chem.,Liu_Shen_2021_RSCMed.Chem.}\\
Coronavirus papain-like proteases  &   Protonation states of His/Cys; proton-coupled conformational dynamics  & GBNeck2 & \cite{Henderson_Shen_2020_J.Chem.Phys.}\\
Coronavirus main proteases  & Protonation states of His and Cys; proton-coupled conformational change of the binding site & GBNeck2 & \cite{Verma_Shen_2020_J.Am.Chem.Soc.}\\
\multicolumn{2}{l}{\bf Transmembrane proteins} & \\
Proton channel M2 & Proton-coupled channel opening/closing & Membrane-hybrid & \cite{Chen_Shen_2016_J.Phys.Chem.Lett.} \\
Sodium/proton antiporter NhaA   & Identification of proton binding residues; proton-coupled conformational changes & Membrane-hybrid & \cite{Huang_Shen_2016_Nat.Commun.,Henderson_Shen_2020_Proc.Natl.Acad.Sci.USA} \\ Efflux pump AcrB  &  Identification of proton binding residues; proton-coupled conformational transitions & Membrane-hybrid & \cite{Yue_Shen_2017_J.Chem.TheoryComput.}  \\
$\mu$-opioid receptor & {\pka} calculation & Membrane-hybrid & \cite{Vo_Ellis_2021_Nat.Commun.,Mahinthichaichan_Shen_2021_JACSAu}\\
\multicolumn{2}{l}{\bf Materials} & \\
Various surfactants & {\pka} calculation in micelle & GBSW and hybrid& \cite{Morrow_Shen_2011_J.Phys.Chem.B} \\
Fatty acid & {\pka} calculation in micelle and bilayer & All-atom& \cite{Morrow_Shen_2014_J.Chem.Phys.} \\
Fatty acid & pH-dependent micelle/bilayer formation & Hybrid& \cite{Morrow_Shen_2012_J.Chem.Phys.,Morrow_Shen_2013_Langmuir}\\
Peptide & pH-dependent unfolding of $\beta$-sheets; phase transition {\pka} & All-atom &  \cite{Cote_Shen_2014_J.Phys.Chem.C}\\
Polysaccharide chitosan & Glucosamine titration; pH-dependent dissociation of a model crystallite; phase transition {\pka}; pH-dependent interaction with a surfactant & All-atom & \cite{Morrow_Shen_2015_J.Am.Chem.Soc.,Tsai_Shen_2018_Chem.Mater.}\\
\end{tabular}
\end{center}
\caption{\textbf{Applications of the continuous constant pH MD (CpHMD) methods.}
GBSW-CpHMD \cite{Khandogin_Brooks_2005_Biophys.J.,Khandogin_Brooks_2006_Biochemistry},
hybrid-solvent \cite{Wallace_Shen_2011_J.Chem.TheoryComput.} and membrane-enabled hybrid-solvent \cite{Huang_Shen_2016_Nat.Commun.} CpHMD methods are implemented in CHARMM \cite{Brooks_Karplus_2009_J.Comput.Chem.}.
The GRF \cite{Wallace_Shen_2012_J.Chem.Phys.,Chen_Shen_2013_Biophys.J.}
and PME \cite{Huang_Shen_2016_J.Chem.TheoryComput.}  based all-atom 
CpHMD methods are also implemented in CHARMM \cite{Brooks_Karplus_2009_J.Comput.Chem.} and they can be used with co-titrating ions \cite{Wallace_Shen_2012_J.Chem.Phys.} or water \cite{Chen_Shen_2013_Biophys.J.}.
GBNeck2-CpHMD \cite{Huang_Shen_2018_J.Chem.Inf.Model.,Harris_Shen_2019_J.Chem.Inf.Model.} method is implemented in Amber18 \cite{Case_Kollman_2018_}.
The pH replica-exchange protocol \cite{Wallace_Shen_2011_J.Chem.TheoryComput.}
was used for all applications except for those with GBSW-CpHMD where the temperature replica-exchange protocol \cite{Khandogin_Brooks_2006_Biochemistry} was used.
All implementations except for GBNeck2-CpHMD are for CPU computing.
}
\label{Table:applications}
\end{table*}

### 2022-08-08 3_hphmd.tex
Revision updates

\section{Workflow for the hybrid-solvent CpHMD in CHARMM} 
The hybrid-solvent CpHMD method makes uses of explicit solvent for conformational dynamics and a GB model for propagating protonation states
\cite{Wallace_Shen_2011_J.Chem.TheoryComput.}.
As all MD simulations, CpHMD consists of three stages: system preparation, equilibration, and production.
We will explain each stage using a specific protein as an example. 
To follow along, the tutorial files for soluble proteins can be 
downloaded from our GitLab website \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm}{cphmd-tutorial/hphmd\_charmm}
and for transmembrane proteins, 
the tutorial files can be downloaded from \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm}{cphmd-tutorial/memb\_hphmd\_charmm}.
We will use the pH replica-exchange protocol \cite{Wallace_Shen_2011_J.Chem.TheoryComput.}
in all the examples.

 %--------------------------------------------- 
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=6.6in]{figs/workflow_scheme.pdf}
    \caption{\textbf{Overview of the CpHMD workflow} 
    %Step 5: Add 0 and 1 to the S range.
    %Titration plot, S ==> S_i
    }
    \label{fig:intro}
\end{figure*}
 %---------------------------------------------

%------------------------------------------------------------------------------------
\subsection{System Preparation}

To demonstrate CpHMD simulations of 
soluble proteins without retaining the crystal waters,
we use the mini-protein BBL as an example.
The tutorial files for system preparation can be found 
in the directory
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/bbl_sys_prep}{bbl\_sys\_prep}.

\begin{checklist}{Input files for preparation of a soluble protein}
%\begin{checklist}{Files Produced}
%\textbf{A list of files needed for the equilibration of a soluble protein}
\begin{itemize}
% \item System coordinates (.pdb/.crd)
% \item System topology (.psf)
%%\item Crystal Image Transformations (crystal$\_$image.str)
% \item Periodic boundary conditions (.pbc)
\item step1\_add\_h.inp
\item step2a\_make\_waterbox.inp
\item step2b\_solvate.inp 
\end{itemize}
\end{checklist}

\paragraph{Step 1. Prepare a protein structure with dummy hydrogens.}
First, we retrieve the coordinates of a protein structure
from \href{https://www.rcsb.org/}{Protein Data Bank}\cite{Berman_Bourne_2000_NucleicAcidsRes.},
e.g., BBL (PDB: 1w4h \cite{Ferguson_Fersht_2005_J.Mol.Biol.}).
A homology modeling tool such as SWISS-MODEL\cite{Waterhouse_Schwede_2018_NucleicAcidsRes.} 
can be used to add any missing atomic coordinates.
Next, we convert the PDB file from the standard PDB format to the CHARMM PDB format using the MMTSB tool command \cite{Feig_Brooks_2004_J.Mol.Graph.Model.} \href{http://blue11.bch.msu.edu/mmtsb/convpdb.pl}{convpdb.pl}
or a user supplied script followed by a shell command ``sed'' to
change all histidine residue names to HSP to facilitate titration.

%
\begin{lstlisting}
$ perl convpdb.pl -out charmm22 -segnames
-renumber 1 1w4h.pdb | sed "s/HSD/HSP/g" 
> start.pdb
\end{lstlisting}
%
Here, the output PDB file is formatted for CHARMM (-out charmm22); the segment names are included in the output (-segnames); the residue identification numbers (resid) are renumbered starting at 1; and 
the input file is "1w4h.pdb."
The CHARMM default tautomer for His is HSD.
Changing HSD to HSP indicates to the CpHMD program that histidine is titratable (see later discussion of the parameter file).
Although all resid have been renumbered starting from 1, this does not have to be the case.
Next we add hydrogen atoms to the protein using the HBUILD facility\cite{Brunger_Karplus_1988_Proteins} in CHARMM
\cite{Brooks_Karplus_2009_J.Comput.Chem.}.

To allow titration of Asp and Glu sidechains, we place dummy hydrogens on both carboxylate oxygens.
Note, the dummy atoms do not interact with each other 
and only one is turned on when Asp/Glu is protonated
\cite{Khandogin_Brooks_2005_Biophys.J.}.
The dummy hydrogens are placed in the \textit{syn} position, 
because it is considered
more favorable than the \textit{anti} position
(see discussion in Ref \cite{Khandogin_Brooks_2005_Biophys.J.}).
To prevent the dummy hydrogens from moving to the \textit{anti} position, 
in which case nonnative hydrogen bonds may form
and a neutral dummy may lose the ability to gain charge,
in the CpHMD specific force field parameter file,
the C--O bond rotation energy barrier
is increased from the standard 0.5 to 6.0 kcal/mol
(see discussion in Ref \cite{Khandogin_Brooks_2005_Biophys.J.}).

Following the addition of hydrogens, a brief energy minimization is performed to correct unfavorable positions.
Here 50 steps of steepest descent (SD) and 10 steps of adopted basis Newton-Raphson (ABNR) are used, whereby a harmonic restraint with the force constant of 50 kcal/mol/{\AA}$^2$ is placed on the heavy atoms.
In many experiments, the N- and C-terminal ends of the protein are truncated and require capping. 
For N-terminus we use CH$_3$CO (patch ACE),
and for C-terminus we use NH$_2$ (patch CT2) or NHCH$_3$ (patch CT3).
In special cases, the free ionized forms, NH$_{3}^{+}$ and COO$^-$, 
are used and can be set titratable when CpHMD is turned on. 
The input file 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/bbl_sys_prep}
{step1\_add\_h\_prot.inp} is an example of 
how to perform these steps on BBL and can be run with the following command.
%
\begin{lstlisting}[language=bash]
$ charmm -i step1_add_h_prot.inp -o step1.out
\end{lstlisting}
%
The example input also checks for disulfide bonds using a sulfur-to-sulfur distance cutoff of 2.5 {\AA}. 
After running any input file, make sure to visualize the output 
structure to confirm all the dummy hydrogens and additional features are correctly constructed.

\paragraph{Step 2. Prepare a solvated system}
Now we move to protein solvation by first building a large 
water box of the desired size.
This is done by generating a plane of small cubic water boxes 
along the x- and y-axes, and then stacking the planar water boxes along the z-axis.
CHARMM toppar directory contains the cubic water box made
of pre-equilibrated TIP3P waters.
The size of the large water box is calculated based on the longest
dimension of the protein and a padding space between the protein 
and the edges of the water box.
We recommend having at least $\sim$10 {\AA} padding in x, y, and z directions.
The large cubic water box is then reshaped to a truncated octahedron by removing waters in all eight corners. 
The input file
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/bbl_sys_prep}
{step2a\_make\_waterbox.inp}
is used to construct the octahedron water box.
%
\begin{lstlisting}[language=bash]
$ charmm -i step2a_make_waterbox.inp -o 
step2_make_waterbox.out
\end{lstlisting}
%
Once the water box is built, the protein is placed at the center and water molecules within 2.8 {\AA} of the protein heavy atoms are deleted.

The solvated system is subject to energy minimization to release unfavorable contacts. 
First, the protein heavy atoms are fixed and the system undergoes energy minimization using SD and ABNR for 50 steps each. 
Next, a five-stage minimization is conducted with the harmonic constants
of 100, 50, 25, 5, and 0 kcal/mol on the protein heavy atoms using 50 steps of SD and 100 steps of ABNR in each stage. 
The input file
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/bbl_sys_prep}
{step2b\_solvate.inp} can be used.
%
\begin{lstlisting}[language=bash]
$ charmm -i step2b_solvate.inp -o
step2b_solvate.out
\end{lstlisting}
%
Again, always remember to visualize the system and make sure the system has been constructed to your expectations.

\subsection{System preparation that includes crystal waters}

\begin{checklist}{Required inputs}
%\begin{checklist}{Files Produced}
%\textbf{A list of files needed for the equilibration of a soluble protein}
\begin{itemize}
% \item System coordinates (.pdb/.crd)
% \item System topology (.psf)
%%\item Crystal Image Transformations (crystal$\_$image.str)
% \item Periodic boundary conditions (.pbc)
\item step1a\_add\_h\_prot.inp
\item step1b\_add\_h\_crys\_wat.inp
\item step1c\_merge\_prot\_crys\_wat.inp
\item step2a\_make\_waterbox\_crys\_wat.inp
\item step2b\_solvate\_crys\_wat.inp 
\end{itemize}
\end{checklist}

In some cases the X-ray structure contains water molecules and it is desirable to keep them, e.g., those in the enzyme active site.
A tutorial showing how to set up a system with crystal waters can be
found in the directory \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/crys_wat_system_prep}{crys\_wat\_system\_prep}.
Here we use a small protein, staphylococcus nuclease (SNase), as an example,
as the X-ray structure (PDB ID: 3hzx) contains crystal waters.
To keep the crystal waters, just a few additional steps are needed.
First, move all the water molecules into a separate PDB file and make sure the file is properly formatted with the correct column names and column spacing.
In the tutorial the crystal water coordinates have been exported into the the file
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/crys_wat_system_prep}
{xwat$\_$original.pdb}; however, this file needs to be reformatted for use in CHARMM. 
The Python script
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/crys_wat_system_prep}
{mod$\_$xwat.py}
reformats this file to
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/crys_wat_system_prep}
{xwat$\_$.pdb}.
We should note that most error messages involving the handling of crystal waters are due to the PDB file not being properly formatted.

Since the crystal waters resolved in a X-ray structure do not contain hydrogen positions, they can be added using the CHARMM input file
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/crys_wat_system_prep}
{step1b$\_$add$\_$h$\_$crys$\_$wat.inp}.
With hydrogens added to the protein and crystal waters, the two sets of coordinates need to be combined using the CHARMM input file
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/crys_wat_system_prep}
{step1c$\_$merge$\_$prot$\_$crys$\_$wat.inp}.
The rest of the system preparation is very similar to
that without crystal waters.
The only difference to note is when you solvate the system you can merge the segids of the crystal waters (segid: XWAT) with the bulk water (segid: SOLV) using the following command.
%
\begin{lstlisting}[language=bash]
JOIN XWAT SOLV renumber
RENAME SEGID SOLV sele segid XWAT end
\end{lstlisting}
%
Keep in mind this needs to be done after the deletion of water that overlaps with the protein; otherwise you run the risk of accidentally removing some crystal waters.
An example input file is
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/crys_wat_system_prep}
{step2b$\_$solvate$\_$crys$\_$wat.inp}.

\subsection{System preparation for transmembrane proteins}

\begin{checklist}{Required inputs}
\begin{itemize}
\item initial$\_$system$\_$prep/step1-5/*.inp
\item initial$\_$system$\_$prep/step6/*.inp
\item final$\_$system$\_$prep/*.inp
\end{itemize}
\end{checklist}
Due to the presence of a lipid bilayer, the preparation of transmembrane proteins for CpHMD simulations requires additional steps.
We divide the preparation into the initial and final stages.
The initial preparation is for fixed-charged equilibration,
whereas the final preparation is for CpHMD simulations.
The related files can be accessed from the directories
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/}{initial\_system\_prep}
and \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep/}{final\_system\_prep}.

To begin the initial system preparation, a solvated protein-membrane system is built.
%we first prepare the protein structure using `step1\_genpsf.inp'.
We start by retrieving the protein structure from the database \href{https://opm.phar.umich.edu/}{OPM} (Orientations of Proteins in Membranes), which properly orients the protein in the lipid bilayer using theoretical calculations and experimental data.\cite{Lomize_Mosberg_2006_Bioinformatics}
Just like for soluble proteins, any missing residues need to be added.
%and then the missing hydrogens can be added.
Here we adopt the standard protonation states, i.e., Asp(-), Glu(-), Cys(0), Lys(+), and Arg(+).
As the solution or model {\pka} of His is 6.54 \cite{Thurlkill_Pace_2006_ProteinSci.}, 
which is within one unit of the physiological pH, special attention needs to be given to its protonation (singly or doubly protonated) and tautomeric state (HSD or HSE) by considering 
the nearby hydrogen bonding and/or salt-bridge interactions. 
% The protonation state of His can be set by changing the residue name in the PDB file and running the script. 
Like soluble proteins, the N- and C-terminal ends may be capped or if necessary the charged forms may be used.

With the complete protein structure (all heavy atom positions
are in place), we use the script
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5/step1_genpsf.inp}{step1$\_$genpsf.inp} to build
hydrogen positions followed by a short energy minimization in the GB implicit-solvent model to adjust any energetically
unfavorable hydrogen positions.
Next, we build the rest of the system and equilibrate the
lipid bilayer
following the CHARMM-GUI protocol \cite{Jo_Im_2007_PLoSONE}
for placing the lipids, solvating the system, and adding ions. 
The input files for these steps can be found in the directory \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5}{step1-5}.
Below we will briefly go through the steps.

After the protein structure is prepared, we 
place it at the origin and estimate its 
cross section using the script
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5/step2.1_orient.inp}{step2.1$\_$orient.inp}. 
The next step is to add water to the pore of the protein 
using the script \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5/step2.2_pwat.inp}{step2.2$\_$pwat.inp}.
To determine the system dimensions the script \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5/step3.1_size.inp}{step3.1$\_$size.inp} is used. 
This script also estimates the lipid type area and the number of lipids required to be placed between the protein boundary and the periodic boundary. 
Approximately 4 to 5 lipids should be placed between the protein and the periodic boundary. 
Additionally, the script calculates the thickness of the water layer above and below the lipid bilayer. 
With the dimensions of the system determined, the lipid heads using big dummy spheres are placed to estimate the positions of the lipids in the bilayer. This is accomplished by \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5/step3.2_packing.inp}{step3.2$\_$packing.inp}.
Next, the dummy spheres are replaced with lipid molecules from a library of randomized conformations using the script \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5/step4.1_lipid.inp}{step4.1$\_$lipid.inp}.
In this example, 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine (POPC) is used, but if other lipids are desired,
the libraries of randomized conformations of lipid molecules can be found in the \href{https://www.charmm-gui.org/?doc=archive&lib=lipid}{CHARMM-GUI Archive - Individual Lipid Molecule Library}.

After the lipid bilayer is built, a rectangular water box is generated using the script \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5/step4.2_waterbox.inp}{step4.2$\_$waterbox.inp}, which solvates the protein/membrane complex.
Next, the number of ions required to achieve the desired salt concentration is calculated, 
and the ions are randomly placed in the slabs of water above and below the membrane. 
In the final step, all of the system components are combined using the script \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5/step5_assembly.inp}{step5$\_$assembly.inp}.

We now relax the assembled protein/membrane system in six stages with gradually reduced harmonic 
restraints using the fixed-charge MD CHARMM-GUI scripts in the directory
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step6}{step6}.
Following the system relaxation, a much longer (100 ns) simulation in the NPT ensemble is required to fully equilibrate the lipid bilayer.
The protein heavy atoms are harmonically restrained during the simulation.
The progress of the bilayer equilibration is monitored by
the calculation of three properties: surface area per lipid, bilayer thickness, and lipid order parameters \cite{Klauda_Pastor_2010_J.Phys.Chem.B,Huang_Shen_2021_MethodsinMolecularBiology}. 
The system is considered equilibrated when these properties plateau with the simulation time.
Since this can be performed on any scalable or GPU-enabled MD engine using a standard protocol, we leave this step up to the user.

Following the fixed-charge equilibration of the lipid bilayer,
we move to the final system preparation.
Here, we prepare the protein for CpHMD titration by adding dummy hydrogens.
Example input files can be found in the directory \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep}{final\_system\_prep}. 
%to show how the system is prepared for CpHMD equilibration.
The first step to break the final snap shot from the bilayer equilibration into separate components (protein, bilayer, bulk water, pore water, and ions).
We add hydrogens using the script
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep/step1a_add_h.inp}{step1a$\_$add$\_$h.inp}.
PSF and CRD files are generated for the rest of the system components 
with the following scripts:
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep/step1b_genpsf_memb.inp}{step1b$\_$genpsf$\_$memb.inp}, 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep/step1c_genpsf_water_bulk.inp}{step1c$\_$genpsf$\_$water$\_$bulk.inp}, 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep/step1d_genpsf_water_pwat.inp}{step1d$\_$genpsf$\_$water$\_$pwat.inp},
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep/step1e_genpsf_ion_sod.inp}{step1e$\_$genpsf$\_$ion$\_$sod.inp}, and
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep/step1f_genpsf_ion_cla.inp}{step1f$\_$genpsf$\_$ion$\_$cla.inp}.
Once all the PSF and CRD files have been made,
the rest of the system can be rebuilt using the script 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep/step2_assemble.inp}{step2$\_$assemble.inp}.

%--------------------------------------------------------------------------------
\subsection{Equilibration of soluble proteins}
%--------------------------------------------------------------------------------

\begin{checklist}{Files Required}
\begin{itemize}
\item equil\_hphmd.inp
\end{itemize}
\end{checklist}

Equilibration of soluble proteins for CpHMD is performed at a single pH in several stages. 
An example input for the BBL protein is
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/bbl_equil_prod}{equil\_hphmd.inp}.
Since CpHMD is turned on during equilibration, titratable residues must be chosen. 
In the example here, Asp, Glu, and His are allowed to titrate because of the pH range used in the production simulation.
The crystal structure pH condition or physiological pH 7.4 can be used.
We specify the physiological ionic strength of 0.15 M and temperature of 298 K.
Before invoking the PHMD module for CpHMD
simulation, we need to call GBSW using the following command.
%
\begin{lstlisting}
GBSW HYBRID SGAMMA 0.005 NANG 50 CONC 0.15 - 
SELE protein END
\end{lstlisting}
%
Here, \textbf{SGAMMA} (default 0.005) is the non-polar surface tension coefficient, \textbf{NANG} (default 50) is the number of angular integration points, and \textbf{CONC} (default 0.15) is the ionic strength in M.
Except for \textbf{CONC}, which the user may change according to the specific problem, the default parameters should be used.

Now we invoke PHMD and set related options using the following command.

\begin{lstlisting}
PHMD PAR 23 WRI 25 PH 7.4 NPRI 5000 PHFRQ 5 
    MASS 10.0 BARR 2.0 BETA 5.0 TEMP 298 LAM -
    SELE RESN ASP .or. RESN GLU .or. RESN HSP end
\end{lstlisting}
%
Here \textbf{PH} is the simulation pH (default 7);
\textbf{NPRI} is the frequency of printing lambda values (5000 means 10 ps);
\textbf{PHFRQ} is the update frequency for the lambda coordinates (default 5);
\textbf{TEMP} is the temperature for $\lambda$ dynamics (default 298);
\textbf{MASS} is the fictitious mass of the $\lambda$ particles (default 10); 
\textbf{BARR} is the quadratic barrier height (default 2.0 and is overwritten by the values specified in the parameter file phmd.inp); 
\textbf{BETA} is the friction coefficient for lambda dynamics (default 5.0); 
\textbf{LAM} tells the PHMD module to output lambda values; and
\textbf{SELE} specifies the types of residues allowed to titrate (default all residues in phmd.inp file). 
Except for PH, NPRI, and SELE options, 
the default parameters should be used.

The length of each stage depends on the system and should be adjusted accordingly.
In the BBL example \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/bbl_equil_prod}{equil\_hphmd.inp},
the heating/equilibration stages were run for 20 ps, 
because the system is very small and stable. 
First, the system is heated in the NVT ensemble, and next
four stages of equilibration are performed in the NPT ensemble with decreasing harmonic restraints on the heavy atom positions, e.g., 5.0, 1.0, 0.1, and 0 kcal/mol/{\AA}$^2$. 
The idea here is to slowly relax the system by removing any high energy contacts and to allow for the initialization of protonation states at the specified pH (e.g., 7).
Note, the equilibration time for the mini-protein BBL
is short, and one can adjust it accordingly for larger proteins.

\subsection{Equilibration of transmembrane proteins}

\begin{checklist}{Required input files}
\begin{itemize}
\item  cphmd$\_$equil/step0.1$\_$equil.inp
\item  cphmd$\_$equil/step0.2$\_$equil.inp
\item  cphmd$\_$equil/step0.3$\_$equil.inp
\end{itemize}
\end{checklist}

Now three short equilibration steps can be performed using CpHMD at a single pH.
Example inputs are in the directory \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/cphmd_equil}{cphmd\_equil}. 
As for a soluble protein, the specified pH should be the pH used in the experiment (e.g., crystal growth, NMR, etc) or physiological pH.
In the first step, the CpHMD simulation is performed 
in the NVT ensemble
with a harmonic force constant of 1.0 kcal/mol/$\mbox{\AA}^2$ for 0.1 ns using
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/cphmd_equil}{step0.1$\_$equil.inp}.
In the second step, the CpHMD simulation is performed in the NPT ensemble using
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/cphmd_equil}{step0.2$\_$equil.inp}.
In the third step, the simulation (\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/cphmd_equil}{step0.3$\_$equil.inp}) continues in the NPT ensemble for 0.8 ns with all restraints are removed. 
We note, the membrane-enabled CpHMD simulation makes use of the membrane GBSW model
\cite{Im_Brooks_2003_Biophys.J.}.

\begin{lstlisting}
GBSW HYBRID SGAMMA 0.0 NANG 50 CONC 0.15 - 
     TEMP 310 TMEMB 30 MSW 2.5 -
     RCYLN 15 SELE solu END
\end{lstlisting}
%
The options \textbf{TMEMB}, \textbf{MSM}, and \textbf{RCYLN} are used 
to specify a low dielectric slab in the GBSW model with a high dielectric cylinder centered in the protein.
\textbf{TMEMB} specifies the slab thickness ($\mbox{\AA}$) and should be set to the average difference in the $Z$ positions of the lipid C2 atoms from the top and bottom leaflets. This will vary for different lipid types. 
\textbf{MSM} specifies the half membrane switching length for the dielectric transition between the low-dielectric slab and the bulk solution; 
the default is 2.5 {\AA}.
Finally, \textbf{RCYLN} is the radius of a high dielectric water cylinder placed in the center of the protein; 
it should be set to an appropriate value such that the water cylinder covers the entire protein but does not overlap with lipids as much as possible.
Although no positional restraints are placed on the protein, we recommend using a cylindrical restraint with a harmonic force constant of 0.1 kcal/mol/{\AA}$^{2}$ to prevent the protein from a lateral drift. 
This can be done using the MMFP facility in CHARMM \cite{Brooks_Karplus_2009_J.Comput.Chem.}.

\subsection{Production}

\begin{checklist}{Files Produced}
\begin{itemize}
\item CpHMD Trajectories per pH (.dcd)
\item Lambda Values per pH (.lamb)
\item Replica Exchange Info. per pH (.log)
\item Energy Output per pH (.ene)
\item MD Info. per pH (.out)
\item Restarts per pH (.rst)
\end{itemize}
\end{checklist}

The CpHMD production stage is similar for soluble and transmembrane proteins. The only difference is that the GBSW command includes additional options for the lipid bilayer (see discussion of the equilibration steps) for transmembrane protein simulations.
An example input for the production run of a
soluble protein is
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/bbl_equil_prod}{prod$\_$hphrex.inp}.
A similar production input for a transmembrane protein is
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/prod}{prod$\_$hphrex.inp}.
An example job submission file for a HPC cluster is 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/prod}{hphrex.qsub}.

To begin the production run one must choose a pH range and number of pH replicas. 
The proper choice of the pH range depends on the research objective.
For example, the pH range of 1.5--6.5 was used in the hybrid-solvent CpHMD study
of the pH-dependent structure function relationship of the endosomal protein BACE1
in which the catalytic aspartyl residues play a central role \cite{Ellis_Shen_2015_J.Am.Chem.Soc.}.
On the other hand, the pH range 4--8 was used in the membrane-enabled hybrid-solvent 
CpHMD study of the pH-dependent conformational activation of the transmembrane M2 channel,
whereby the titration of four histidines modulates the conformation of the channel \cite{Chen_Shen_2016_J.Phys.Chem.Lett.}.
Next, one needs to choose the total number of pH replicas. 
The spacing of the pH conditions needs to be close enough such that sufficient exchange 
probabilities (e.g., at least 20$\%$) are obtained.
To obtain good overlap between potential energy distributions of neighboring replicas
a pH spacing of 0.5 and sometimes even 0.25 is required. 
Other aspects to consider are the number of available CPUs on the HPC and how many CPUs can be 
allocated to each replica. 
Simulations with CHARMM do not scale well past 8 CPUs, so for a job using 16 replicas
we recommend requesting a total of 128 CPUs with 8 CPUs for each replica. 
We recommend making exchange attempts every 1000 MD steps.
A typical pH replica-exchange CpHMD simulation has an average exchange ratio of $\sim$40$\%$, 
but if the exchange ratios are too low, more replicas can be added to fill in the spacing 
between replicas.
To achieve optimum performance in enhanced sampling, several replicas should move 
across all pH conditions.

To use pH-based replica exchange in CHARMM, one can use the 
\href{https://hpc.nih.gov/apps/charmm/c39b2html/repdstr.html}{REPD} module.
%
\begin{lstlisting}
REPD NREP 8 PHMD EXLM FREQ 5000 UNIT 17
\end{lstlisting}
%
Here, the \textbf{REPD} command initiates replica exchange, and \textbf{NREP} states the number of replicas to use
(in this case 8). 
\textbf{PHMD EXLM} states this is a special type of Hamiltonian exchange (i.e., pH-based) and allows for the exchange of pH conditions of two adjacent replicas.
\textbf{FREQ} states the frequency of an exchange attempt.

Then, to run several replicas in parallel, CHARMM reads in a rep.cmd file for each pH replica 
with the syntax, rep.cmd$\_$0, rep.cmd$\_$1, to rep.cmd$\_$(NREPS), using the following 
command in the input file.
%
\begin{lstlisting}[language=bash]
$ stream rep.cmd
\end{lstlisting}
% 
Notice this command lacks the '$\_0$' portion of the name.
In each of the rep.cmd file you will have the command that states the pH environment of the replica.
%
\begin{lstlisting}
PHMD RESPH OLDPH @crysph NEWPH @ph PKATEMP @temp
\end{lstlisting}
%
Here \textbf{OLDPH} is the pH used in the equilibration and \textbf{NEWPH} is the pH of the replica.
Finally, \textbf{PKATEMP} should be set to the same temperature of the simulation. 
The rest of the replica exchange set up follows a standard MD run in CHARMM.
Production run inputs for BBL (prod$\_$hphrex.inp and rep.cmd) can be used to run one 2 ns stage of pH-Replica Exchange CpHMD.

\subsection{General settings for MD and CpHMD in CHARMM}
% CHARMM Specific
% Things to Include: 
%   -) FFs
%   -) Thermostats
%   -) Baristas
%   -) Cutoffs
%   -) ect...

For all simulations the protein 
is represented by CHARMM22/CMAP force field.\cite{MacKerell_Karplus_1998_J.Phys.Chem.B,MacKerell_Brooks_2004_J.Am.Chem.Soc.}
Note, the current CpHMD methods are not compatible with the CHARMM36 protein force field.
For solvent the CHARMM modified TIP3P water model is used.
For ions or lipid molecules the CHARMM36 force fields are used \cite{Klauda_Pastor_2010_J.Phys.Chem.B}.
The van der Waals energies are calculated with a switching function from 10 to 12 {\AA} and  a cutoff at 14 {\AA}. 
All simulations are conducted under periodic boundary conditions. 
The PME method \cite{Darden_Pedersen_1993_J.Chem.Phys.} is used to calculate the long-range electrostatic energies and forces with a real space calculation cutoff of 12 {\AA} and 1 {\AA} grid spacing, and a 6th-order spline interpolation. 
The non-bond neighbor list is updated heuristically. 
Additionally, the SHAKE \cite{Ryckaert_Berendsen_1977_J.Comput.Phys.} algorithm is used to constrain all bonds involving hydrogens to allow for a 2-fs time step. 
For NPT simulations, the temperature is maintained with the N\'{o}se-Hoover thermostat \cite{Nose_Nose_1984_Mol.Phys.,Hoover_Hoover_1985_Phys.Rev.Aa}.
The pressure is maintained with the Langevin piston pressure-coupling algorithm \cite{Feller_Brooks_1995_J.Chem.Phys.}.
For the CpHMD simulations, the default GBSW radii \cite{Nina_Roux_1997_J.Phys.Chem.B,Chen_Brooks_2006_J.Am.Chem.Soc.} are used for the protein.
For small molecules, the generic atomic radii in the GBSW file may be used upon user verification. 
A CpHMD parameter file phmd.inp contains
the model reference {\pka} values and the parameters in the titration potential of mean force for Asp, Glu, His, Cys, and Lys. 
Parameters for additional titratable groups
can be obtained by following the steps discussed in section Parameterization.
% As shown in the example inputs the titration ($\lambda$) and tautomer ($\chi$) should be updated every 5 MD steps, allowing for water and lipid relaxation.\cite{Wallace_Shen_2011_J.Chem.TheoryComput.}
% A friction coefficient of 5 ps$^{-1}$ should be used for Langevin dynamics and the mass of the titration coordinate ($\lambda$ and $\chi$) should be 10 atomic mass units. https://www.overleaf.com/project/60d33785a3cd6c24e62c6d32

### 2022-08-08 Response.tex

\documentclass[11pt,letterpaper]{businessletter}
\special{papersize=8.5in,11in} %make sure paper size is correct
\usepackage[]{fancyhdr,helvet}
\usepackage[]{amsmath,fancyhdr,helvet,afterpage,overcite,xurl}
\usepackage[colorlinks]{hyperref}
\usepackage[usenames,dvipsnames]{color}
\usepackage[left=1in,right=1in,top=1in,bottom=0in,nohead]{geometry}

% \hypersetup
% {
%  pdftitle = {},
%  colorlinks=true,
%  linkcolor=NavyBlue,
%  citecolor=NavyBlue,
%  urlcolor=blue
% }

%arial font
\renewcommand{\familydefault}{\sfdefault}
\urlstyle{rm}
\newcommand{\pka}{p$K_a$}

%suppress indentation and add space between paragraphs
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}

\signature{Jana Shen}
\begin{document}
\begin{letter}
{
Editor \\
Live Journal of Computational Molecular Science \\
}

\opening{Dear Editor,}
We are pleased to submit a revised manuscript entitled
``A Guide to the Continuous Constant pH Molecular Dynamics 
Methods in Amber and CHARMM v1.0
in \textit{Live Journal of Computational Molecular Science}.

We thank the reviewers for their constructive critiques
and helpful comments as well as suggestions. 
We have made revision to the manuscript to fully address and accommodate them.
Below we present our response and revision.  
The revised excerpts are highlighted in the 
revised manuscript. 
Unless otherwise noted, page numbers refer to the revised version.

\bigskip
{\large \bf Reviewer E}

We thank the reviewer for the positive view of the manuscript and we address the minor comments below.

\textbf{Comment 1:}

\textcolor{Blue}{
The presence of explicit solvent creates a considerable challenge and essentially renders the discrete algorithm essentially impractical. One compromise is to execute the discrete jump of protonation assuming an implicit solvent model, while propagating the unbiased dynamics with explicit solvent. The models here are mixed. The relative balance of protonation states is governed under the implicit solvent, but the conformational sampling results from the explicit solvent. This is certainly an interesting avenue, but the users should be told clearly and unambigusouly that the ionization equilibrium is really that of an implicit solvent model. Please make sure this is clear.
}

\textbf{Response and Revision:} 
We agree with the reviewer.

We discussed the hybrid-solvent constant pH scheme
on page 2. 
To make it more clear to the reader, we now separate the discussion of the three solvent schemes in three paragraphs.
To further clarify the hybrid-solvent CpHMD method, 
we added the sentence in the beginning of ``3. Workflow for the hybrid-solvent CpHMD in CHARMM'':

On page 7, right, just under the section 3 title:

\textcolor{red}{
The hybrid-solvent CpHMD method makes uses of explicit solvent 
for conformational dynamics and a GB model for propagating protonation states [33].
}

\textbf{Comment 2:}

\textcolor{Blue}{
Certainly, lambda-dynamics and the non-equilibrium MD/MC algorithms provide more rigorous solutions to generate constant pH simulation consistently within the explicit solvent model. It is correct to qualify the neMD/MC algorithm as "discrete" in the sense that system only visits the end-states, although is would also be illuminating to point out that the non-equilibrium switches require a continuous coupling parameter lambda like lambda-dynamics. In neMD/MC, lambda is changed from 0 to 1 according to a schedule,while in lambda-dynamics the coupling parameters are free to evolve dynamically. There is some conceptual similarity between lambda-dynamics and neMD/MC and pointing out this similarity may be beneficial to the users.
}

\textbf{Response and Revision:} 
We agree with the reviewer and expanded our discussion 
of the neMD/MC approach.

On page 2, left, last paragraph continued onto the first paragraph on the right:

\textcolor{red}{
Development of an all-atom DpHMD algorithm based on the
hybrid MD/MC approach is challenging, 
as in explicit solvent a switch in protonation state
leads to a large energy change, which results in
almost complete rejection for the MC moves [29].
This problem, which is due to the lack of overlap between explicit solvent
configurations for the protonated and deprotonated states [29],
has been tackled by introducing a free energy calculation [37]
or a short non-equilibrium MD (neMD) [28,29,38]
(see later discussion).
}

Page 2, right, last paragraph, continued to the first paragraph 
on page 3, left:

\textcolor{red}{
To circumvent the aforementioned configuration overlap problem
in a DpHMD framework, B\"{u}rgi and van Gunsteren used thermodynamic integration (TI)
to calculate the titration free energy change for the MC move [37];
however, TI calculations are too expensive
and do not readily converge.
Recently, an all-atom DpHMD method based on neMD and MC
has been implemented in NAMD [54] 
by the Roux group [29,38].
This approach, which was originally proposed by Stern [28],
has the flavor of both TI and $\lambda$ dynamics
methods, as it makes use of a coupling parameter ($\lambda$) in
a short neMD trajectory to gradually change the protonation 
state thus allowing solvent to adjust [29,38].
}

\textbf{Comment 3:}

\textcolor{Blue}{
The text says "Although the latter is a major caveat, the CpHMD method offers significantly faster convergence and importantly it can be implemented for implicit-,hybrid-, and fully explicit-solvent simulations." I am not sure why lambda-dynamics can achieve faster convergence than discrete jumps in the case of implicit solvent simulations. Since implicit solvent simulations permit the purest form of the discrete algorithm, what is the advantage to execute lambda-dynamics then? Isn’t a simple discrete jump with MC conceptually simpler than lambda-dynamics (no need to masses,intermediate states, etc)? Based on my understanding of the scripting for CHARMM and AMBER, I would think that if you have the infrastructure to run lambda-MD, converting this to remove the MD on the lambda and replace it by random discrete jumps of lambda between 0 and 1 could be easily realized. Please comment.
}

\textbf{Response and Revision:} 
We agree with the reviewer that the hybrid MD/MC or neMD/MC DpHMD methods are conceptually simpler
and practically easier to implement than the lambda-dynamics based CpHMD methods. 
In fact, the first DpHMD method was implemented as a Linux Shell
script by Baptista (Baptista et al, JCP 2002). 
As to {\pka} convergence, we regret that the sentence wasn't phrased well.
What we wanted to say was that CpHMD methods converge significantly faster 
for coupled residues and can be readily extended to all-atom simulations.
The convergence advantage is due to the fact that at every MD step the lambda values of all residues are updated,
whereas in a DpHMD method, the protonation state of a single residue is attempted at each MC step.
Although attempting to update multiple sites is possible, 
the acceptance ratio for such a MC move may be lower.
To clarify this, we revised and expanded the corresponding text:

On page 2, left, end of the second last paragraph:

\textcolor{red}{
Besides the advantage that only physical (protonated and deprotonated) states are sampled, the DpHMD method based on
the hybrid MD/MC scheme is conceptually simple
and practically straightforward to implement.
...
Although sampling the unphysical intermediate states is a major caveat, the CpHMD method offers significantly faster convergence for coupled residues,
and importantly it can be readily extended to all-atom simulations.
The faster convergence arises from the fact that at every MD step the $\lambda$ values of 
all titratable sites are updated, whereas in a typical DpHMD implementation, the protonation state of 
a single residue is attempted at each MC step.
}

\textbf{Response and Revision:}

\textbf{Comment 4:}

\textcolor{Blue}{
The text says "The all-atom CpHMD method by the Shen group is the only one that includes titratable water to compensate for the net charge fluctuation as a result of proton titration in the CpHMD simulation [48]." In fact, previous implementation of the neMD/MC algorithm also included charge neutrality (Chen and Roux, JCTC 2015) where protonation was coupled to the conversion of a water molecule into a chloride anion, and deprotonation was coupled to the conversion of a potassium into a water molecule (eqs 30 and 31 in their paper). The NAMD implementation of Radak et al (JCTC 2017) supports this feature also, as multiple sites can be protonated and deprotonated simultaneously. However, Chen and Roux showed that charge neutrality acts as a constraint and decreases the acceptance ratio. It is like to have similar effects in lambda-dynamics. Lastly, charge neutrality is not a prerequisite in simulations with PBC and PME (in fact, charge neutrality is necessarily violated during a charging free energy calculation)
}

\textbf{Response and Revision:} 
It was our oversight to miss this aspect in the work of Roux and coworkers.  
We thank the reviewer for pointing it out. 
We revised the text accordingly.

Page 2, right, second last paragraph:

\textcolor{red}{
To compensate for the net charge fluctuation due to proton titration, the Shen group introduced the first
co-titratable ion [34] or water [50] scheme.
}

Page 3, left, first paragraph:

\textcolor{red}{
Analogous to the titratable ion or water approach of the Shen group [34,50], the neMD/MC implementation of the Roux group used the chemical transformation 
between a counterion and water to maintain charge neutrality of the system [29,38].
Interestingly, while the Roux group reported that charge neutrality 
acts as a constraint and decreases the acceptance ratio in the MC steps [29,38], 
the Shen group observed a slow down in the convergence of protonation state sampling [35,50,52].
}

\textbf{Comment 5:}

\textcolor{Blue}{
Page 2 \\
The text says "While this may well be true when the switch of one protonation state is involved, the problem quickly becomes intractable when multiple titration sites are involved. " Perhaps emphasize that the problem becomes intractable when the number N of sites is large because the number of possible ionization states N goes like 2$^N$.
}

\textbf{Response and Revision:} We thank the reviewer for the great suggestion and we revised the text.

Page 3, left, second paragraph:

\textcolor{red}{
While this may well be true when the switch of one protonation state is involved, the problem becomes intractable when 
the number $N$ of titratable sites is large because
the number of possible protonation states increases as 2$^N$.
}

\textbf{Comment 6:}

\textcolor{Blue}{
Page 3 \\
The equation for the forces on the lambda variable implicitly assumes that there are only changes in charges and Lennard-Jones paramters. But what about other changes in the force field? 
In the CHARMM force field, dihedral angles often depend on the ionization state of a side chain.
}

\textbf{Response and Revision:} 
In the current CpHMD implementations (by the Brooks, Shen, and Grubm\"{u}ller groups) as well as the implementations of the hybrid MD/MC approach (by Mongan, McCammon and Case and later by the Roitberg group), the bonded part is not scaled and the bonded parameters of one of the protonation states are used.
This is certainly a limitation that needs to be addressed in the future. 
We added this clarification.

Page 3, right, below Eq. 2:

\textcolor{red}{
We note that the force field parameters of some residue types (e.g., carboxylates) may
depend on the protonation state; however, in the CpHMD [31,33,35,44,45,48]
as well as the hybrid MD/MC based DpHMD [26,27]
implementations the bonded parameters 
are fixed in one of the protonation states.
This is a limitation that needs to be addressed in the future.
}

\textbf{Comment 7:}

\textcolor{Blue}{
MINOR\\
In the Abstract: "This is a significant drawback, as" --> perhaps change drawback with limitation?
}

\textbf{Response and Revision:} 
We thank the reviewer for catching the imprecise wording.
We made the revision.

Page 1, abstract:

This is a significant 
\textcolor{red}{limitation}, as ...

\textbf{Comment 8:}
\textcolor{Blue}{
Page 5\\
The writing style is clear but perhaps sounds a bit like spoken words with many "We", "you", "our", etc... For example, "Like all computational chemistry calculations, it is important to know your system as much as possible before any serious". Perhaps change by "it is important to know the system of interest" and generally use a slightly less familiar tone.
}

\textbf{Response and Revision:} 
We agree with the reviewer that the writing style is informal. 
We would not use such a style for articles, 
but this is a tutorial so we opted to write in a familiar tone. 
Have said that, we agree with the reviewer's suggestion
and we revised the sentence and made edits throughout 
the paper to make the style less informal.

Page 6, left, first paragraph:

\textcolor{red}{
..., it is important to know the system of interest
before attempting any simulations.
}

\bigskip
{\large \bf Reviewer H}

We appreciate the reviewer's favorable view of the tutorial,
and we address the minor comments below.

\textbf{Comment 1:}
\textcolor{Blue}{
There are some inconsistent file names. 
I noticed two cases: (1) in the lower right code block on page 7, the INP filename should be step2a\_xxx; (2) in the first line on page 9, it should be step1b\_add\_h\_crys\_xwat.inp.
}

\textbf{Response and Revision:}

We thank the reviewer for catching the errors.
We have gone through the tutorial and fixed all inconsistent names; we also added some missing hyperlinks.

\textbf{Comment 2:}
\textcolor{Blue}{
There is a weird line in the lower right of page 10 where has a single 'step0.2\_equil.inp' hyperlink line.
}

\textbf{Response and Revision:} 
This is now fixed.

\textbf{Comment 3:}
\textcolor{Blue}{
There are some broken url links: (1) the url point to GitHub in the first page; (2) the README link in upper left of page 14; (3) the 'cphmd\_parm\_fit.py.' link in upper right of page 18. I checked the authors LaTeX on GitHub and found out the reason is that the authors have a new line for the web url in \url command. Once I delete that, all url links work well.
}

\textbf{Response and Revision:} 
We appreciate that the reviewer found out the problem.

\textbf{Comment 4:}
\textcolor{Blue}{
The url links are not consistent. Some file links are actually folder links. It would be good if the authors can reformat the links.
}

\textbf{Response and Revision:} 
We thank the reviewer for pointing out this issue. 
We have fixed all url links.

\bigskip

{\large \bf Reviewer J}

\textbf{Comment 1:}
\textcolor{Blue}{
It seems that the instructions provided by the authors on the gitlab repo about patching AMBER20 are not completely correct.
Specifically, in AMBER20 it seems there is no longer a "src" folder under the \$AMBERHOME path that points to the installation folder, and instead all the source files are left in the source folder. Furthermore, attempting to patch the latest distribution of amber20 results in several errors as follow:
"Reversed (or previously applied) patch detected! Assume -R?
X out of X hunks FAILED -- saving rejects to file ./src/cuda/base\_simulationConst.cpp.rej" Pushing throught this errors and attempting to rebuild pmemd resulted in the following cmake error:
"9 errors detected in the compilation of ./src/pmemd/src/cuda/gti\_cuda.cu.
CMake Error at pmemd\_cuda\_DPFP\_generated\_gtei\_cuda.cu.o.RELEASE.cmake:278" Nevertheless it seems I was able to run the simulations correctly, despite failing in completing the patching of amber20, so I am not completely sure what happened there, but this issue may affect other readers"
}

\textbf{Response and Revision:} 
We appreciate that the reviewer has taken the time to test the code.
We have updated the gitlab repository to clarify that the src directory that must be patched falls in the Amber source directory rather than \$AMBERHOME.

As to the errors in ``attempting to patch the latest distribution of amber20'', we suggest that they may have to do with applying the CpHMD patch to the Amber20 release version and not the fully patched version. 
We are raising the issue with the Amber developers and will get it resolved.

\textbf{Comment 2:}
\textcolor{Blue}{
The mdin input files generated by cphmd\_prep.sh were actually named with two decimal points, not one as they appear in the tutorial (e.g 3BDC\_chainA\_A\_pH7.50\_prod.mdin instead of 3BDC\_chainA\_A\_pH7.5\_prod.mdin)
}

\textbf{Response and Revision:} 
We thank the reviewer for catching this error.
We updated the script so that the number of decimal 
points in file names depends on the
'-dph' flag (default value is 0.5 with 1 decimal point). 
The files in the GitLab repository were updated accordingly.

\textbf{Comment 3:}
\textcolor{Blue}{
Attempting to run the mdin files generated by cphmd\_prep.sh results in a Fortran runtime error because of an Interger format overflow due to the size of the value of the ntwr flag wich is set to 1*10$^{10}$. Reducing the lenght to 1*10$^6$ gets rid of the error,
ntwr flag which is set to 1*10$^{10}$. Reducing the length to 1*10$^6$ gets rid of the error, although I did not test the effect on the performance.
}

\textbf{Response and Revision:} 
We thank the reviewer for catching this error, which is now fixed by making ntwx the same value as ntwr 
as we do in other running modes.

\textbf{Comment 4:}
\textcolor{Blue}{
In addition to the technical issues highlighted before, I think that the tutorial will benefit from a more extended explanaition about which keywords to use and modify in AMBER in order to run CpHMD simulations, rather than exclusively focusing on how to use cphmd$\_$prep.sh to set up the simulations. In my opinion, the reliance on that script makes the whole process a bit too much of a blackbox.
}

\textbf{Response and Revision:} 
We thank the reviewer for the suggestion and we added two paragraphs before explaining the shell script cphmd$\_$prep.sh.

Page 13, right, first two paragraphs:

\textcolor{red}{
\indent Before explaining {cphmd\_prep.sh}, we briefly review the CpHMD specific files and options in Amber here.
Similar to 3.1 (System Preparation), we first convert the PDB file to the Amber format, cap the terminal groups, add dummy hydrogens to the titratable Asp and Glu residues and change their names to AS2 and GL2.
Besides the standard Amber force field files, tleap requires two CpHMD specific files for building topology and parameters:
{frcmod.phmd},
which specifies the modifications of bonded parameters for AS2 and GL2, 
and {phmd.lib}, which
contains the definitions of AS2 and GL2.
For GBNeck2-CpHMD, set the PBradii to mbondi3 in tleap (see later discussion of modifications). 
For running CpHMD, an additional file
{gbneck2$\_$input.parm} is needed, which contains the
partial charges of the protonated/deprotonated forms of the titratable residues as well as the parameters of the model titration PMFs (see section 7). 
\\
\indent
To run CpHMD, the \textit{pmemd.cuda} command should contain several CpHMD options.
The option
\textbf{-phmdparm} reads in 
the aforementioned CpHMD parameter file {gbneck2\_input.parm}.
The option \textbf{-phmdin} reads in
a CpHMD job control file, which specifies various options for CpHMD simulations. 
We note, except for  
\textbf{MaskTitrRes(:)} (titratable residue types, e.g., 'AS2','GL2','HIP','CYS','LYS')
and \textbf{MaskTitrResTypes} (number of titratable residues, e.g., 5), the default settings in the phmdin file should be kept.
Finally, if specified the option \textbf{-phmdstrt} reads in a CpHMD restart file.
We should also point out that 
for CpHMD runs, the Amber mdin files are
similar to those for the fixed-charge MD using GBNeck2 implicit solvent and Langevin dynamics. 
The two CpHMD specific flags are \textbf{iphmd} and \textbf{solvph}. 
For GBNeck2-CpHMD, \textbf{iphmd} should be set to 1 and \textbf{solvph} should be set to the desired pH condition.
}

\textbf{Comment 5:}
\textcolor{Blue}{
It will good if the authors could provide additional context to the sentence "If the objective is to predict the pKa values of soluble proteins or their protonation states at a certain pH condition, GBNeck2-CpHMD [41, 42] is the best choice..." in the first paragraph of the second column in page 5.
}

\textbf{Response and Revision:}

We revised and expanded the discussion.

Page 6, left, first paragraph:

\textcolor{red}{
If the objective is to predict the {\pka} values of soluble proteins 
or their protonation states at a certain pH condition, we recommend
the GPU-accelerated GBNeck2-CpHMD in Amber [45], 
as the {\pka's} converge rapidly and the accuracies 
for titrating Asp, Glu, His, Cys, and Lys (in solvent-exposed and buried sites) have been validated using a large number of proteins (see references in Table 1).
If the objective is to investigate the detailed proton-coupled conformational dynamics, proton transfer, protein-ligand binding/unbinding, or transmembrane proteins, we currently recommend 
the hybrid-solvent CpHMD [33],
as the accuracy has been extensively validated in terms of
{\pka} values and description of proton-dependent conformational dynamics
(see example applications in Table 1),
although speed is limited due to the use of CPUs.
We note, a GPU-accelerated all-atom PME CpHMD 
implementation in Amber22 [51]
has been recently released by us [52] 
and holds a promise to offer more accurate description of
proton-coupled conformational dynamics for heterogeneous systems
such as protein-ligand complexes and transmembrane proteins.
}

\textbf{Comment 6:}
\textcolor{Blue}{
In page 6, when discussing set up of CUDA environment, it may be better to use a generic CUDA version as they do with the AMBER version.
}

\textbf{Response and Revision:} 
The CpHMD method uses the same CUDA version as Amber. To clarify this, we revised the related sentence.

%'CUDA 10.1 is the primary CUDA version for running it but early versions from CUDA 7.5 also works' to

\textcolor{red}{
A version of CUDA required by Amber needs to be installed.
}

In summary, we have addressed all the comments from the reviewers and made a thorough revision. As a result, we believe the quality of the tutorial has been significantly improved and the revised version is now ready for publication.

\closing{Sincerely,}

\end{letter}
\end{document}

### 2022-08-08 Old/1_methods.tex

--- 2022-01-23 1_methods.tex
@@
 of atom $a$, and $\theta_i$ refers to the titration coordinate of titratable residue $i$. 
-The two first terms of Eq.~\ref{eq:Hamiltonian} give the kinetic energies of real atoms and virtual $\lambda$ particles. 
+The two first terms of Eq.~\ref{eq:Hamiltonian} give the kinetic energies of real atoms 
+and virtual $\lambda$ particles. 
 $U^{\rm int}$ represent the titration-independent bonded and non-bond energies. 
+We note that the force field parameters of some residue types (e.g., carboxylates) may
+depend on the protonation state; however, in the CpHMD \cite{Khandogin_Brooks_2005_Biophys.J.}
+as well as the hybrid MD/MC based DpHMD \cite{Mongan_McCammon_2004_J.Comput.Chem.}
+implementations the bonded parameters 
+are fixed in one of the protonation states.
+This is a limitation that needs to be addressed in the future.
+
 The fourth term, $U^{\rm hybr}$, describes the 

### 2022-08-08 letterhead.tex

\setlength\topmargin{-0.50 in}
{\normalfont
\begin{flushright}
{\Large {\bf \color{Maroon} University of Maryland School of Pharmacy}} 
   %\hspace{\fill} 
%\includegraphics[height=20truept]{OUlogo}
\end{flushright}
\vspace{-0.5cm}
\begin{tabular*}{\linewidth}{c} \hline \end{tabular*}
\begin{flushright}
{\footnotesize {\bf {\em Jana Shen, Ph.D.}} \hspace{\fill} 
        {\em  Department of Pharmaceutical Sciences} }     \\
{\footnotesize {\em Professor} \hspace{\fill} 
        {\em  Computer-Aided Drug Design Center} }                    \\
{\footnotesize {\em Codirector, Computer-Aided Drug Design Center} \hspace{\fill}
        {\em University of Maryland School of Pharmacy} }  \\
{\footnotesize {\em Phone: (410) 706-4187} \hspace{\fill} 
        {\em Baltimore, MD 21201} }          \linebreak
{\footnotesize {\em E-mail: \url{jana.shen@rx.umaryland.edu}} \hspace{\fill} 
        {\em \url{http://www.computchem.org}} }                               \\
        \end{flushright}
\vspace{-0.5cm}
}

### 2022-08-08 Old/3_hphmd_old.tex

\section{Workflow for the hybrid-solvent CpHMD in CHARMM} 
As all MD simulations, CpHMD consists of three stages: system preparation, equilibration, and production.
We will explain each stage using a specific protein as an example. 
To follow along, the tutorial files for soluble proteins can be 
downloaded from our GitLab website \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm}{cphmd-tutorial/hphmd\_charmm}
and for transmembrane proteins, 
the tutorial files can be downloaded from \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm}{cphmd-tutorial/memb\_hphmd\_charmm}.
We will use the pH replica-exchange protocol \cite{Wallace_Shen_2011_J.Chem.TheoryComput.}
in all the examples.

 %--------------------------------------------- 
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=6.6in]{figs/workflow_scheme.pdf}
    \caption{\textbf{Overview of the CpHMD workflow} 
    %Step 5: Add 0 and 1 to the S range.
    %Titration plot, S ==> S_i
    }
    \label{fig:intro}
\end{figure*}
 %---------------------------------------------

%------------------------------------------------------------------------------------
\subsection{System Preparation}

To demonstrate CpHMD simulations of 
soluble proteins without retaining the crystal waters,
we use the mini-protein BBL as an example.
The tutorial files for system preparation can be found 
in the directory
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/bbl_sys_prep}{bbl\_sys\_prep}.

\begin{checklist}{Input files for preparation of a soluble protein}
%\begin{checklist}{Files Produced}
%\textbf{A list of files needed for the equilibration of a soluble protein}
\begin{itemize}
% \item System coordinates (.pdb/.crd)
% \item System topology (.psf)
%%\item Crystal Image Transformations (crystal$\_$image.str)
% \item Periodic boundary conditions (.pbc)
\item step1\_add\_h.inp
\item step2a\_make\_waterbox.inp
\item step2b\_solvate.inp 
\end{itemize}
\end{checklist}

\paragraph{Step 1. Prepare a protein structure with dummy hydrogens.}
First, we retrieve the coordinates of a protein structure
from \href{https://www.rcsb.org/}{Protein Data Bank}\cite{Berman_Bourne_2000_NucleicAcidsRes.},
e.g., BBL (PDB: 1w4h \cite{Ferguson_Fersht_2005_J.Mol.Biol.}).
A homology modeling tool such as SWISS-MODEL\cite{Waterhouse_Schwede_2018_NucleicAcidsRes.} 
can be used to add any missing atomic coordinates.
Next, we convert the PDB file from the standard PDB format to the CHARMM PDB format using the MMTSB tool command \cite{Feig_Brooks_2004_J.Mol.Graph.Model.} \href{http://blue11.bch.msu.edu/mmtsb/convpdb.pl}{convpdb.pl}
or a user supplied script followed by a shell command ``sed'' to
change all histidine residue names to HSP to facilitate titration.

%
\begin{lstlisting}
$ perl convpdb.pl -out charmm22 -segnames
-renumber 1 1w4h.pdb | sed "s/HSD/HSP/g" 
> start.pdb
\end{lstlisting}
%
Here, the output PDB file is formatted for CHARMM (-out charmm22); the segment names are included in the output (-segnames); the residue identification numbers (resid) are renumbered starting at 1; and 
the input file is "1w4h.pdb."
The CHARMM default tautomer for His is HSD.
Changing HSD to HSP indicates to the CpHMD program that histidine is titratable (see later discussion of the parameter file).
Although all resid have been renumbered starting from 1, this does not have to be the case.
Next we add hydrogen atoms to the protein using the HBUILD facility\cite{Brunger_Karplus_1988_Proteins} in CHARMM
\cite{Brooks_Karplus_2009_J.Comput.Chem.}.

To allow titration of Asp and Glu sidechains, we place dummy hydrogens on both carboxylate oxygens.
Note, the dummy atoms do not interact with each other 
and only one is turned on when Asp/Glu is protonated
\cite{Khandogin_Brooks_2005_Biophys.J.}.
The dummy hydrogens are placed in the \textit{syn} position, 
because it is considered
more favorable than the \textit{anti} position
(see discussion in Ref \cite{Khandogin_Brooks_2005_Biophys.J.}).
To prevent the dummy hydrogens from moving to the \textit{anti} position, in which case nonnative hydrogen bonds may form
and a neutral dummy may lose the ability to gain charge,
in the CpHMD specific force field parameter file,
the C--O bond rotation energy barrier
is increased from the standard 0.5 to 6.0 kcal/mol
(see discussion in Ref \cite{Khandogin_Brooks_2005_Biophys.J.}).

Following the addition of hydrogens, a brief energy minimization is performed to correct unfavorable positions.
Here 50 steps of steepest descent (SD) and 10 steps of adopted basis Newton-Raphson (ABNR) are used, whereby a harmonic restraint with the force constant of 50 kcal/mol/{\AA}$^2$ is placed on the heavy atoms.
In most cases the N- and C-terminal ends of the protein are truncated and require capping. 
For N-terminus, we use CH$_3$CO (patch ACE) and 
for C-terminus we use NH$_2$ (patch CT2) or NHCH$_3$ (patch CT3).
In special cases, the free ionized form, NH$_{3}^{+}$ and COO$^+$, is used and can be set titratable when CpHMD is turned on. 
The input file `step1a\_add\_h\_prot.inp' is an example of how to perform these steps on BBL and can be run with the following command.
%
\begin{lstlisting}[language=bash]
$ charmm -i step1a_add_h_prot.inp -o step1a.out
\end{lstlisting}
%
The example input also checks for disulfide bonds using a sulfur-to-sulfur distance cutoff of 2.5 {\AA}. 
After running any input file, make sure to visualize the output 
structure to confirm all the dummy hydrogens and additional features are correctly constructed.

\paragraph{Step 2. Prepare a solvated system}
Now we move to protein solvation by first building a large 
water box of the desired size.
This is done by generating a plane of small cubic water boxes 
along the x- and y-axes, and then stacking the planar water boxes along the z-axis.
CHARMM toppar directory contains the cubic water box made
of pre-equilibrated TIP3P waters.
The size of the large water box is calculated based on the longest
dimension of the protein and a padding space between the protein 
and the edges of the water box.
We recommend having at least $\sim$10{\AA} padding in x, y, and z directions.
The large cubic water box is then reshaped to a truncated octahedron by removing waters in all eight corners. 
The input file `step2a\_make\_waterbox.inp' is used to construct the octahedron water box.
%
\begin{lstlisting}[language=bash]
$ charmm -i step2_make_waterbox.inp -o 
step2_make_waterbox.out
\end{lstlisting}
%
Once the water box is built, the protein is placed at the center and water molecules within 2.8 {\AA} of the protein heavy atoms are deleted.

The solvated system is subject to energy minimization to release unfavorable contacts. 
First, the protein heavy atoms are fixed and the system undergoes energy minimization using SD and ABNR for 50 steps each. 
Next, a five-stage minimization is conducted with the harmonic constants
of 100, 50, 25, 5, and 0 kcal/mol on the protein heavy atoms using 50 steps of SD and 100 steps of ABNR in each stage. 
The input file
``step3\_solvate.inp'' 
can be used.
%
\begin{lstlisting}[language=bash]
$ charmm -i step2b_solvate.inp -o
step2b_solvate.out
\end{lstlisting}
%
Again, always remember to visualize the system and make sure the system has been constructed to your expectations.

\subsection{System preparation that includes crystal waters}

\begin{checklist}{Required inputs}
%\begin{checklist}{Files Produced}
%\textbf{A list of files needed for the equilibration of a soluble protein}
\begin{itemize}
% \item System coordinates (.pdb/.crd)
% \item System topology (.psf)
%%\item Crystal Image Transformations (crystal$\_$image.str)
% \item Periodic boundary conditions (.pbc)
\item step1a\_add\_h\_prot.inp
\item step1b\_add\_h\_crys\_wat.inp
\item step1c\_merge\_prot\_crys\_wat.inp
\item step2a\_make\_waterbox\_crys\_wat.inp
\item step2b\_solvate\_crys\_wat.inp 
\end{itemize}
\end{checklist}

In some cases the X-ray structure contains water molecules and it is desirable to keep them, e.g., those in the enzyme active site.
A tutorial showing how to set up a system with crystal waters can be
found in the directory \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/crys_wat_system_prep}{crys\_wat\_system\_prep}.
Here we use a small protein, staphylococcus nuclease (SNase), as an example,
as the X-ray structure (PDBid: 3hzx) contains crystal waters.
% In the previous example, the solution NMR structure of BBL did not contain crystal waters.\cite{Ferguson_JMolBiol_2005_v353_p427}
To keep the crystal waters, just a few additional steps are needed.
First, move all the water molecules into a separate PDB file and make sure the file is properly formatted with the correct column names and column spacing.
In the tutorial the crystal water coordinates have been exported into the the file 'xwat$\_$original.pdb'; however, this file needs to be formatted for use in CHARMM. 
The Python script 'mod$\_$xwat.py' reformats this file, and in the example the reformatted file is named 'xwat$\_$final.pdb', and you might be able to modify the script for your simulation needs.
Most error messages involving the handling of crystal waters are due to the PDB file not being properly formatted.

Since the crystal waters resolved in a X-ray structure do not contain hydrogen positions, they can be added using the CHARMM input file 'step1b$\_$add$\_$h$\_$xwat.inp'.
With hydrogens added to the protein and crystal waters, the two sets of coordinates need to be combined using the CHARMM input file 'step1c$\_$merge$\_$prot$\_$xwat.inp'.
The rest of the system preparation is very similar to
that without crystal waters.
The only difference to note is when you solvate the system you can merge the segids of the crystal waters (segid: XWAT) with the bulk water (segid: SOLV) using the following command.
%
\begin{lstlisting}[language=bash]
JOIN XWAT SOLV renumber
RENAME SEGID SOLV sele segid XWAT end
\end{lstlisting}
%
Keep in mind this needs to be done after the deletion of water that overlaps with the protein; otherwise you run the risk of accidentally removing some crystal waters.
An example of this is the input file 'step2b$\_$solvate$\_$crys$\_$wat.inp'.

\subsection{System preparation for transmembrane proteins}

\begin{checklist}{Required inputs}
 \begin{itemize}
 \item step1-5/*.inp
  \item step6/*.inp
% \item System Atom Coordinates (.pdb/.crd)
% \item System Atom Topology (.psf)
% \item System Assemtion nbly Info. (.str)
% \item Crystal Image Transformations (crystal$\_$image.str)
% \item Periodic Boundary Conditions (.pbc)
% \item Pre-Equil. System Coordinates (.crd/.pdb)
% \item Pre-Equil. System Trajectory (.dcd)
% \item Pre-Equil. System Information (.out)
% \item Pre-Equil. System Energies (.ene)
% \item Pre-Equil. System Restart (.rst)
\end{itemize}
\end{checklist}
Due to the presence of a lipid bilayer, the preparation of transmembrane proteins for CpHMD simulations requires additional steps.
We divide the preparation into the initial and final stages.
% First, we assemble the system including the protein
% and the solvated lipid bilayer followed by the relaxation of the bilayer. 
The related files can be accessed from the directories
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/}{initial\_system\_prep}
and \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep/}{final\_system\_prep}.

In the initial system preparation, a solvated protein-membrane system is built.
%we first prepare the protein structure using `step1\_genpsf.inp'.
We start by retrieving the protein structure from the database \href{https://opm.phar.umich.edu/}{OPM} (Orientations of Proteins in Membranes), which properly orients the protein in the lipid bilayer using theoretical calculations and experimental data.\cite{Lomize_Mosberg_2006_Bioinformatics}
Just like for soluble proteins, any missing residues need to be added.
%and then the missing hydrogens can be added.
Here we adopt the standard protonation states, i.e., Asp(-), Glu(-), Cys(0), Lys(+), and Arg(+).
As the solution or model {\pka} of His is 6.54 \cite{Thurlkill_Pace_2006_ProteinSci.}, 
which is within one unit of the physiological pH, special attention needs to be given to its protonation (singly or doubly protonated) and tautomeric state (HSD or HSE), based on the nearby hydrogen bonding or salt-bridge interaction. 
% The protonation state of His can be set by changing the residue name in the PDB file and running the script. 
Like soluble proteins, the N- and C-terminal ends may be capped or if necessary the charged forms may be used.

With the complete protein structure (all heavy atom positions
are in place), we use the script
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5}{step1\_genpsf.inp} to build
hydrogen positions followed by a short energy minimization in the GB implicit-solvent model to adjust any energetically
unfavorable hydrogen positions.
Next, we build the rest of the system and equilibrate the
lipid bilayer
following the CHARMM-GUI protocol for placing the lipids, solvating the system, and adding ions.\cite{Jo_Im_2007_PLoSONE}
The input files for these steps can be found in the directory \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5}{step1-5}.
Below we will briefly go through the steps.

After the protein structure is prepared, we 
place it at the origin and estimate its 
cross section using the script
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5}{step2.1$\_$orient.inp}. 
The next step is to add water to the pore of the protein 
using the script \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5}{step2.2$\_$pwat.inp}.
To determine the system dimensions the script \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5}{step3.1$\_$size.inp} is used. This script also assesses the lipid type area and the number of lipids required to be placed between the protein boundary and the periodic boundary condition. 
Approximately 4 to 5 lipids should be placed between the protein and the PBC. 
Additionally, the script will state the thickness of the water slab above and below the lipid bilayer. 
With the dimensions of the system calculated, the lipid heads using big dummy spheres are placed to estimate the positions of the lipids in the bilayer. This is accomplished by \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5}{step3.2$\_$packing.inp}.
Then the dummy spheres are replaced with lipid molecules from a library of randomized conformations using the script \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5}{step4.1$\_$lipid.inp}.
In this example 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine (POPC) lipids are used, but if different lipids are needed other libraries of randomized conformations of lipid molecules can be found on the \href{https://www.charmm-gui.org/?doc=archive&lib=lipid}{CHARMM-GUI Archive - Individual Lipid Molecule Library}.

Next, a rectangular water box is generated using the input file \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5}{step4.2$\_$waterbox.inp}, which solvates the protein and membrane complex.
Next the number of ions required to achieve the desired salt concentration is calculated and randomly placed in the slabs of water above and below the protein and membrane complex. 
In the final step, \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step1-5}{step5$\_$assembly.inp}, all of the system components are combined.
After the system is built, 
all components of the system are relaxed in six equilibration stages with gradually decreased harmonic restraint potentials using the scripts  \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/initial_system_prep/step6}{step6.1$\_$equi.inp to 'step6.6$\_$equi.inp}.

Now we move to the final system preparation stage

%--------------------------------------------------------------------------------
\subsection{Equilibration of soluble proteins}
%--------------------------------------------------------------------------------

\begin{checklist}{Files Required}
%\textbf{A list of files needed for the equilibration of a soluble protein}
\begin{itemize}
% \item System Atom Coordinates (.crd)
% \item System Atom Topology (.psf)
% \item Periodic Boundary Conditions (.pbc)
\item equil\_hphmd.inp
%\item C22 Protein Topologies Parameters (.rtf/.prm)
%\item Water/Ion Topologies and Parameters (toppar$\_$water$\_$ions.str)
%\item C22 Hybrid PHMD Parameters (toppar$\_$phmd$\_$c22.str)
%\item Hybrid PHMD Parameters (phmd$\_$hybrid$\_$g2$\_$c22.in)
%\item GBSW Radii (radius$\_$gbsw.str)
\end{itemize}
\end{checklist}

% \begin{checklist}{Files Produced}
% %\textbf{A list of files needed for the equilibration of a soluble protein}
% \begin{itemize}
% \item Lambda Values (.lamb)
% \item Trajectory (.dcd)
% \item Energy Outputs (.ene)
% \item Restart (.rst)
% \end{itemize}
% \end{checklist}

Equilibration of soluble proteins for CpHMD is performed at a single pH in several stages. 
An example input for the BBL protein is
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/bbl_equil_prod}{equil\_hphmd.inp}.
% , `\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/bbl_equil_prod}{equilibration$\_$hphmd.inp}.'
Since CpHMD is turned on during equilibration, titratable residues must be chosen. 
In the example here, Asp, Glu, and His are allowed to titrate because of the pH range used in the production simulation.
The crystal structure pH condition or physiological pH 7.4 can be used.
We specify the physiological ionic strength of 0.15 M and temperature of 298 K.
Before invoking the PHMD module for CpHMD
simulation, we need to call GBSW using the following command.
%
\begin{lstlisting}
GBSW HYBRID SGAMMA 0.005 NANG 50 CONC 0.15 - 
SELE protein END
\end{lstlisting}
%
Here, SGAMMA (default 0.005) is the non-polar surface tension coefficient, NANG (default 50) is the number of angular integration points, and CONC (default 0.15) is the ionic strength.
Except for CONC, which the user may change according to the specific problem, the default parameters should be used.

Now we invoke PHMD and set related options using the following command.
%\href{https://hpc.nih.gov/apps/charmm/c42b2html/phmd.html}{PHMD module} to your input file. 
%
\begin{lstlisting}
PHMD PAR 23 WRI 25 PH 7.4 NPRI 5000 PHFRQ 5 
    MASS 10.0 BARR 2.0 BETA 5.0 TEMP 298 LAM -
    SELE RESN ASP .or. RESN GLU .or. RESN HSP end
\end{lstlisting}
%
Here PH is the simulation pH (default 7);
NPRI is the frequency of printing lambda values (5000 means 10 ps);
PHFRQ is the update frequency for the lambda coordinates (default 5);
TEMP is the temperature for $\lambda$ dynamics (default 298);
MASS is the fictitious mass of the $\lambda$ particles (default 10); 
BARR is the quadratic barrier height (default 2.0 and is overwritten by the values specified in the parameter file phmd.inp); 
BETA is the friction coefficient for lambda dynamics (default 5.0); 
LAM tells the PHMD module to output lambda values; and
SELE specifies the types of residues allowed to titrate (default all residues in phmd.inp file). 
Except for PH, NPRI, and SELE options, 
the default settings should be used.

The length of each stage depends on the system and should be adjusted accordingly.
%stages lasting 40 to 120 ps have been used by group for previous projects.
In the BBL example \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/bbl_equil_prod}{equil\_hphmd.inp},
the heating/equilibration stages were run for 20 ps, 
because the system is very small and stable. 
First, the system is heated in the NVT ensemble, and next
% If the system does not require heating the initial and final temperatures can be set to be the same value and the system can equilibrate the water positions under the NVT ensemble.
four stages of equilibration are performed in the NPT ensemble with decreasing harmonic restraints on the heavy atom positions, e.g., 5.0, 1.0, 0.1, and 0 kcal/mol/{\AA}$^2$. 
The idea here is to slowly relax the system by removing any high energy contacts and to allow for the initialization of protonation states at the specified pH (e.g., 7).
% Using four equilibration stages the harmonic constraints on the protein heavy atoms are reduced with the following values 5.0, 1.0, 0.1 (kcal/mol), and then in the last stage all harmonic restraints are released.
% After each equilibration make sure the energies of the system have equilibrated.
Note, the equilibration time for the mini-protein BBL
is short, and you can adjust it accordingly for larger proteins.
%You are now ready to start the production run. 
% The example equilibration input provided moves through each heating/equilibration stage in one input file (equilibration$\_$hphmd.inp).

\subsection{Equilibration of transmembrane proteins}

\begin{checklist}{Required input files}
\begin{itemize}
% \item  System Atom Coordinates (.crd)
% \item  System Atom Topology (.psf)
% \item  Pre-Equil. System Coordinates (.pdb)
% %\item  Final Lipid Equilibration Snapshot (.pdb)
% \item  Crystal Image Transformations (crystal$\_$image.str)
%\item *.crd, *.psf, 
\item  step0.1$\_$equil.inp
\item  step0.2$\_$equil.inp
\item  step0.3$\_$equil.inp
\end{itemize}
\end{checklist}

% \begin{checklist}{Files Produced}
% \begin{itemize}
% \item Lipid Equilibration Trajectories
% \item CpHMD Trajectory at a Single pH (.dcd)
% \item Lambda Values (.lamb)
% \item Energy Outputs (.ene)
% \item Restart (.rst)
% \end{itemize}
% \end{checklist}

Here, the protein heavy atoms are harmonically restrained using a 1.0 kcal/mol/{\AA}$^2$ until the lipid bilayer is equilibrated in all equilibration simulations.
Starting from the final snap shot of the pre-equilibration simulations ('Pre-Equil. System Coordinates'), the simulation is briefly run for 0.2 ns using an NVT ensemble.
Then a much longer NPT ensemble simulation is run to fully equilibrate the simulation. 
These simulations should be run until the surface area per lipid, bilayer thickness, and lipid order parameters are near or within error limits of experimental lipid composition you are simulating.\cite{Huang_Shen_2021_MethodsinMolecularBiology}
These three lipid bilayer metrics are frequently used to assess the phase behaviour and fine structure of the simulated bilayer.\cite{Klauda_Pastor_2010_J.Phys.Chem.B}
Simulations up to 100 ns can be necessary for proper equilibration of the lipid bilayer.
Since this can be performed on any scalable MD engine we leave this step up to the user to decide which MD engine they would like to use.
For convenience we have supplied \href{}{example input files} for performing this step using the MD software NAMD.\cite{Phillips_Schulten_2005_J.Comput.Chem.}

Following equilibration of the lipid bilayer,
the system can now be equilibrated at a single pH with CpHMD. 
To start this process, dummy protons need to be added in the same manner as for soluble proteins. 
Example input files are in the directory \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep}{final\_system\_prep}. 
%to show how the system is prepared for CpHMD equilibration.
% The first step to break the final snap shot from the lipid equilibration simulation into separate components (Protein, Bilayer, Bulk Water, Pore Water, Sodium Ions, Chloride Ions, ect.).
First, we add hydrogens 
%or the protein all of the hydrogens can be removed and re-added 
with the CHARMM input file 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep}{step1a$\_$add$\_$h.inp}.
Next, PSF and CRD files are generated for the rest of the system components with the following scripts
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep}{step1b$\_$genpsf$\_$memb.inp}, \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep}{step1c$\_$genpsf$\_$water$\_$bulk.inp}, 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep}{step1d$\_$genpsf$\_$water$\_$pwat.inp},
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep}{step1e$\_$genpsf$\_$ion$\_$sod.inp}, and
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep}{step1f$\_$genpsf$\_$ion$\_$cla.inp}.
Once all the PSF and CRD files have been made,
the rest of the system can be rebuilt using the CHARMM input file 
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/final_system_prep}{step2$\_$assemble.inp}.

Now three short equilibration steps can be performed using CpHMD at a single pH.
Example inputs are in the directory \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/cphmd_equil}{cphmd\_equil}. 
As for a soluble protein, the specified pH should be the pH used in the crystal growth, physiological pH, or an experimentally relevant pH.
In the first step, the CpHMD simulation is performed 
in the NVT ensemble
with the harmonic force constant of 1.0 kcal/mol/$\mbox{\AA}^2$ for 0.1 ns using the input
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/cphmd_equil}{step0.1$\_$equil.inp}
In the second step, the CpHMD simulation is performed 
in the NPT ensemble, but the first simulation ('step0.1$\_$equil.inp') uses the NVT ensemble and the second ('step0.2$\_$equil.inp') uses the NPT ensemble.

\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/cphmd_equil}{step0.2$\_$equil.inp}

The third simulation ('step0.3$\_$equil.inp') continues with an NPT ensemble for 0.8 ns with all restraints removed. 
For membrane-enabled CpHMD simulations the tmemb features needs to be added to the \href{https://hpc.nih.gov/apps/charmm/c42b2html/gbsw.html}{GBSW module} in the CHARMM input file. 
%
\begin{lstlisting}
GBSW HYBRID SGAMMA 0.0 NANG 50 CONC 0.15 - 
     TEMP 310 TMEMB 30 MSW 2.5 -
     RCYLN 15 SELE solu END
\end{lstlisting}
%
The inputs variables tmemb, msw, and rcyln are used to build a low dielectric slab to represent the lipid bilayer in the GBSW model with a high dielectric hole centered around the protein.
The TMEMB variable sets the slab thickness ($\mbox{\AA}$) and should be set to the average difference in the $Z$ positions of the lipid C2 atoms from the top and bottom leaflets. This will vary for different lipid types. 
The MSM variable is the half membrane switching length\cite{Im_Brooks_2003_Biophys.J.} for the dielectric transition region between the low-dielectric slab and the bulk solution, and is commonly set to 2.5 $\mbox{\AA}$.
Finally, rcyln is the radius of the high dielectric hole in the low dielectric slab and should set so there is minimal overlap between the protein and lipids.
The protein selection, the last line in the GBSW module, required the protein selection to place the hole at the center of the protein.
Without restraints on the protein heavy atoms, a cylindrical restraint with a harmonic force constant (0.1 kcal mol$^{-1}$ {\AA}$^{-2}$) needs to be applied to the protein to prevent from a lateral drift. 
This can be done using the MMFP facility in CHARMM.

\subsection{Production}

% \begin{checklist}{Files Required}
% \begin{itemize}
% \item System Atom Coordinates (.crd)
% \item System Atom Topologies (.psf)
% \Item Final Equilibration Restart (.rst)
% \item Prod. Input (prod\_hphrex.inp)
% \item Replica Exchange Input (rep.cmd)
% \end{itemize}
% \end{checklist}

\begin{checklist}{Files Produced}
\begin{itemize}
\item CpHMD Trajectories per pH (.dcd)
\item Lambda Values per pH (.lamb)
\item Replica Exchange Info. per pH (.log)
\item Energy Output per pH (.ene)
\item MD Info. per pH (.out)
\item Restarts per pH (.rst)
\end{itemize}
\end{checklist}

The CpHMD production stage is the same for both soluble and transmembrane proteins. 
The only differences are in the set up of the GBSW module to include a low dielectric slab to represent the lipid bilayer for transmembrane simulations, which is discussed in the equilibration stage.
On our GitLab page we have provided examples of the production run for both the \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/hphmd_charmm/bbl_equil_prod}{soluble proteins} and \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/memb_hphmd_charmm/prod}{transmembrane proteins} simulations titled 'production$\_$hphrex.inp' with an example queue submission file for a HPC cluster ('hphrex.qsub').
To begin the production run one must choose a pH range and the number of pH replicas. 
For instance, a protein of interest may be active over a particular pH range or you might want to capture conformational dynamics that include specific titratable residues, such as Asp or His. 
Next you need to decide how many pH replicas to use for your production run. 
Somethings to consider are how many CPUs you have available and how many CPUs you want to allocate to each replica. 
Simulations with CHARMM do not scale well past 8 CPUs, so for a job using 16 replicas you will need 128 CPUs. 
The spacing of the pH replicas needs to be close enough to obtain sufficient exchange probabilities (e.g., 20$\%$).
Typically, to obtain overlap between potential energy distributions of neighboring replicas a pH spacing in increments of 0.25 or 0.5 is required. 
Exchange attempts are made every 1000 steps, and a typical CpHMD simulation of this type will have an average exchange rate of $\sim$40$\%$, if exchange rates are too low more replicas can be added to fill in the spacing between replicas.
To achieve enhanced sampling several replicas should move across all pH replicas.
To use pH-based replica exchange in CHARMM the following \href{https://hpc.nih.gov/apps/charmm/c39b2html/repdstr.html}{REPD} module must be used.
%
\begin{lstlisting}
REPD NREP 8 PHMD EXLM FREQ 5000 UNIT 17
\end{lstlisting}
%
Here, the REPD command initiates replica exchange, NREP states the number of replicas to use, in this case 8, and PHMD states this is pH-based replica exchange and allows for the exchange of $\lambda$ along with the spatial coordinates.
Next, EXLM allows for Hamiltonian exchange and FREQ is how often an exchange is attempted. 
Then to run several replicas in parallel CHARMM reads in a rep.cmd file for each pH replica with the syntax, rep.cmd$\_$0, rep.cmd$\_$1, to rep.cmd$\_$(NREPS), using the following command in the input file.
%
\begin{lstlisting}[language=bash]
$ stream rep.cmd
\end{lstlisting}
% 
Notice this command lacks the '$\_0$' portion of the name.
In each of the rep.cmd file you will have the command that states the pH environment of the replica.
%
\begin{lstlisting}
PHMD RESPH OLDPH @crysph NEWPH @ph PKATEMP @temp
\end{lstlisting}
%
Here OLDPH is the pH used in the equilibration and NEWPH is the pH of the replica.
Finally, PKATEMP should be set to the same temperature of the simulation. 
The rest of the replica exchange set up follows a standard molecular dynamics run in CHARMM.
Production run inputs for BBL (production$\_$hphrex.inp and rep.cmd) can be used to run one 2 ns stage of pH-Replica Exchange CpHMD.

\subsection{General settings for MD and CpHMD in CHARMM}
% CHARMM Specific
% Things to Include: 
%   -) FFs
%   -) Thermostats
%   -) Baristas
%   -) Cutoffs
%   -) ect...

For all simulations the protein 
is represented by CHARMM22/CMAP force field.\cite{MacKerell_Karplus_1998_J.Phys.Chem.B,MacKerell_Brooks_2004_J.Am.Chem.Soc.}
Note, the CpHMD simulations are currently not compatible with the CHARMM36 protein force field.
For solvent the CHARMM modified TIP3P water model is used.
For ions or lipid molecules the CHARMM36 force fields are used.\cite{Klauda_Pastor_2010_J.Phys.Chem.B}
The van der Waals energies are calculated with a switching function from 10 to 12 {\AA} and  a cutoff at 14 {\AA}. 
All simulations are conducted under periodic boundary conditions. 
The PME method\cite{Darden_Pedersen_1993_J.Chem.Phys.} is used to calculate the long-range electrostatic energies and forces with a real space calculation cutoff of 12 {\AA} and 1 {\AA} grid spacing, and a 6th-order spline interpolation. 
The non-bond neighbor list is updated heuristically. 
Additionally, the SHAKE\cite{Ryckaert_Berendsen_1977_J.Comput.Phys.} algorithm is used to constrain all bonds involving hydrogens to allow for a 2-fs time step. 
For NPT simulations, the temperature is maintained with the N\'{o}se-Hoover thermostat.\cite{Nose_Nose_1984_Mol.Phys.,Hoover_Hoover_1985_Phys.Rev.Aa}
The pressure is maintained with the Langevin piston pressure-coupling algorithm.\cite{Feller_Brooks_1995_J.Chem.Phys.}
For CpHMD simulations, the default GBSW radii \cite{Nina_Roux_1997_J.Phys.Chem.B,Chen_Brooks_2006_J.Am.Chem.Soc.} are used for the protein.
For the small molecule, generic atomic radii in the GBSW file may be used upon user verification. 
A CpHMD parameter file phmd.inp contains
the model {\pka} values and deprotonation potential of mean force for the titratable groups Asp, Glu, His, Cys, and Lys. 
Parameters for additional titratable groups
can be obtained by following the steps discussed in section Parameterization.
% As shown in the example inputs the titration ($\lambda$) and tautomer ($\chi$) should be updated every 5 MD steps, allowing for water and lipid relaxation.\cite{Wallace_Shen_2011_J.Chem.TheoryComput.}
% A friction coefficient of 5 ps$^{-1}$ should be used for Langevin dynamics and the mass of the titration coordinate ($\lambda$ and $\chi$) should be 10 atomic mass units.

### 2022-08-08 Old/4_aphmd.tex

\section{Workflow for all-atom PME CpHMD in CHARMM} 
% - Yandong
Minor modifications to the workflow for 
hybrid-solvent CpHMD are needed 
to set up and run all-atom PME CpHMD 
simulations in CHARMM \cite{Huang_Shen_2016_J.Chem.TheoryComput.}. 
The example inputs for simulations of BBL are downloadable from  \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/ephmd_charmm/}{ephmd\_charmm}.
To maintain charge neutrality during the
all-atom PME CpHMD simulation, 
titratable water molecules  
\cite{Chen_Shen_2013_Biophys.J.}
(or co-ions \cite{Wallace_Shen_2012_J.Chem.Phys.} in an earlier version) should be added.
Specifically, each acidic (Asp or Glu) or 
basic residue (His or Lys) is 
coupled with one titratable water that can convert to a hydroxide (TIPU) or hydronium (TIPP). Details are explained in Ref \cite{Chen_Shen_2013_Biophys.J.}.
Together with salt ions, titratable water molecules are placed in the water box using the input script
\href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/ephmd_charmm/bbl_sys_prep/}
{step3\_add\_ions\_titr.inp}.

To run an all-atom PME CpHMD simulation, make sure
the command GBSW is NOT invoked before calling PHMD.
The additional options need to be added in the PHMD command to specify the residue names of titratable water (TIPU and TIPP) and the total number as well as residue IDs of the coupled pairs.
For example, in the BBL simulation, 3 Asp, 3 Glu, and 2 His
are titratable, and they are coupled to 8 titratable water molecules. 
%
\begin{lstlisting}
PHMD PAR 23 WRI 25 PH 7.0 NPRI 5000 PHFRQ 5 
    MASS 10.0 BETA 5.0 TEMP 298 LAM -
    sele resn ASP .or. resn GLU .or. TIPU .or. 
    resn HSP .or. TIPP end - 
    qcouple 8 - 
    resi 4  resc 4789 -
    resi 16 resc 4790 -
    resi 20 resc 4791 -
    resi 36 resc 4792 -
    resi 37 resc 4793 -
    resi 39 resc 4794 -
    resi 17 resc 4771 -
    resi 41 resc 4772 
\end{lstlisting}
%
    %-If the user does not 
 %    -want the default, can he/she set it 
Here resn TIPU and TIPP declare the two types of titratable water. 
The option qcouple 8 specifies
a total number of 8 coupled pairs. 
Following qcouple, residue IDs for the titratable group 
(e.g., resi 4) and co-titrating water (e.g., resc 4789) are listed. 
Note, all co-titrating pairs need to be specified manually.

% Finally, the calculated \pka's are 
% corrected with the following formula 
% to account for the finite-size error 
% \cite{Huang_Shen_2016_J.Chem.TheoryComput.}.

% \begin{eqnarray}
% \label{eq:pka_corr}
% \Delta\Delta{G}^{\rm offset} &=&
% \frac{2 \pi}{3} \kappa \gamma^{\rm solv} Q 
% \left(\frac{N}{V}
% - \rho^{\rm pure} \right) \\
% \Delta pK_a^{\rm corr} &=& \pm \frac{\Delta\Delta{G}^{\rm offset}}{ln(10)RT},
% \end{eqnarray}
% where $\gamma^{\rm solv}$ is 0.764 e$\cdot$ {\AA$^2$} 
% for TIP3P water model. 
% $\kappa$ is the electrostatic constant, 
% which is 332.0 kcal 
% $\cdot$ {\AA} $\cdot$ mol$^{-1}$e$^{-2}$. 
% The solvent number density $\rho^{\rm pure}$ is 
% 0.0333679 {\AA}$^3$ for water at ambient temperature 
% and pressure. 
% Q is -1 for acidic groups and +1 for basic groups. 
% The negative sign is for acid groups and positive sign 
% is for basic groups.
% Thus, $\Delta$\pka$^{\rm corr}$ is negative for both 
% acidic and basic groups. 
% N is the number of solvent and V is the volume of the box 
% measured at equilibrium, which can be approximated 
% with the initial volume $V^{\rm init}$ times 0.93. $V^{\rm est} = L^3$ for a cubic box and $V^{\rm est} = 0.77L^3$ for a truncated octahedron box, where L denotes the cubic 
% box size. R is the Boltzmann constant and T is the temperature.

### 2022-08-08 Old/cphmd_tutorial.tex

--- 2022-01-23 cphmd_tutorial.tex
@@
 % you should update the minor version number in preprints and major version number of submissions.
-\newcommand{\githubrepository}{\url{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial}}  %this should be the main github repository for this article
+\newcommand{\githubrepository}{\url{
+https://github.com/JanaShenLab/CpHMD-Tutorial}}  %this should be the main github repository for this article
 
@@
 \author[1,\authfn{4}]{Yandong Huang}
-\author[1]{Vinicius M. de Oliveira}
+\author[1]{Vinicius Martins de Oliveira}
 %\author[1]{Paween Mahinthichaichan}
@@
 \presentadd[\authfn{2}]{Scorpion Therapeutics, Boston, MA}
-\presentadd[\authfn{3}]{ComputChem LLC, Baltimore, MD}
+\presentadd[\authfn{3}]{ComputChem, Baltimore, MD}
 \presentadd[\authfn{4}]{Jimei University, College of Computer Engineering, Xiamen, China}
@@
 We acknowledge financial support from the National Institutes of Health (R01GM098818 and R01CA256557 to Shen and R44GM134756 to Harris) and National Science Foundation (CBET1932963 to Shen). 
+We thank Zhi (Shane) Yue for contributing the input files for the membrane-enabled hybrid-solvent CpHMD simulations.
+We thank Paween Mahinthichaichan for providing comments on the input files for the membrane-enabled hybrid-solvent CpHMD simulations.
 

### 2022-08-09 cphmd_tutorial.tex
Updates from Overleaf

@@
 \title{A Guide to the Continuous Constant pH Molecular Dynamics 
-Methods in Amber and CHARMM v[\versionnumber]}
+Methods in Amber and CHARMM Article [v\versionnumber]}
 
@@
 \author[1,\authfn{1}]{Ruibin Liu}
-% \alsoaffiliation{
-% Joint first author
-% }
 \author[1,\authfn{3}]{Julie A. Harris}
 \author[1,\authfn{4}]{Yandong Huang}
-\author[1]{Vinicius M. de Oliveira}
-%\author[1]{Paween Mahinthichaichan}
+\author[1]{Vinicius Martins de Oliveira}
 \author[1*]{Jana Shen}
@@
 
-\corr{jana.shen@rx.umaryland.edu}{JS}  % Correspondence emails.  FMS and FS are the appropriate authors initials.
-%\corr{email2@example.com}{FS}
+\corr{jana.shen@rx.umaryland.edu}{JS}
 
@@
 \contrib[\authfn{1}]{These authors contributed equally to this work}
-% \contrib[\authfn{2}]{These authors also contributed equally to this work}
 
@@
 \presentadd[\authfn{4}]{Jimei University, College of Computer Engineering, Xiamen, China}
-% \presentadd[\authfn{4}]{Department, Institute, Country}
 
-\blurb{This LiveCoMS document is maintained online on GitLab at \githubrepository.
-To provide feedback, suggestions, or help improve it, please visit the GitLab repository and participate via the issue tracker.}
+\blurb{This LiveCoMS document is maintained online on GitHub at \githubrepository.
+To provide feedback, suggestions, or help improve it, please visit the GitHub repository and participate via the issue tracker.}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\pubDOI{10.XXXX/YYYYYYY}
-\pubvolume{<volume>}
-\pubissue{<issue>}
-\pubyear{<year>}
-\articlenum{<number>}
-\datereceived{Day Month Year}
-\dateaccepted{Day Month Year}
+\pubDOI{10.33011/livecoms.4.1.1563}
+\pubvolume{4}
+\pubissue{1}
+\pubyear{2022}
+\articlenum{1563}
+\datereceived{22 January, 2022}
+\dateaccepted{6 August, 2022}
 
@@
 \begin{abstract}
-%This particular document provides a skeleton illustrating key sections for a Tutorial document.
-%Please see the sample \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates} for additional information on and examples of using the LiveCoMS LaTeX class.
-%Here we also assume familiarity with LaTeX and knowledge of how to include figures, tables, etc.; if you want examples, see the sample just referenced.
-
-%In your work, in this particular slot, please provide an abstract of no more than 250 words.
-%Your abstract should explain the main contributions of your article, and should not contain any material that is not included in the main text.
-%Please note that your abstract, plus the authorship material following it, must not extend beyond the title page or modifications to the LaTeX class will likely be needed.
-
-% Begin Abstract Here
 Like temperature and pressure, solution pH is an important environmental variable in biomolecular simulations. 
 Virtually all proteins depend on pH to maintain their structure and function. 
-In conventional molecular dynamics (MD) simulations of proteins, pH is implicitly accounted for by assigning and fixing protonation states of titratable sidechains. This is a significant drawback, 
+In conventional molecular dynamics (MD) simulations of proteins, pH is implicitly accounted for by assigning and fixing protonation states of titratable sidechains. This is a significant limitation, 
 as the assigned protonation states may be wrong and they may change during dynamics.
@@
 \end{frontmatter}
-
-%Here you would explain what problem you are tackling and briefly motivate your work.
-
-%Keep in mind, as you prepare your manuscript, that you should plan for a representative image  which will be used to highlight your article on the journal website and publications.
-Usually, this would be one of your figures, but it must also be uploaded separately upon article submission.
-We give specific guidelines for this image on the journal website in the section on article submission (see \url{https://livecomsjournal.github.io/authors/policies/index.html#article-submission}).
-
-%Additionally, for well-formatted manuscripts, we recommend that you let LaTeX handle figure/table placement for you as much as possible, so please avoid specifying strenuous float instructions like `[h!]` and `[H]` as much as possible.
 
@@
 \section{Author Contributions}
-%%%%%%%%%%%%%%%%
-% This section mustt describe the actual contributions of
-% author. Since this is an electronic-only journal, there is
-% no length limit when you describe the authors' contributions,
-% so we recommend describing what they actually did rather than
-% simply categorizing them in a small number of
-% predefined roles as might be done in other journals.
-%
-% See the policies ``Policies on Authorship'' section of https://livecoms.github.io
-% for more information on deciding on authorship and author order.
-%%%%%%%%%%%%%%%%
 Henderson, Liu, Harris, Huang, de Oliveira, and Shen
@@
 We thank Jason A. Wallace for a bash script to calculate {\pka} values and track the convergence of unprotonated fractions using Grace.
-% % We suggest you preserve this comment:
-% For a more detailed description of author contributions,
-% see the GitHub issue tracking and changelog at \githubrepository.
-
-%\section{Other Contributions}
-%%%%%%%%%%%%%%%
-% You should include all people who have filed issues that were
-% accepted into the paper, or that upon discussion altered what was in the paper.
-% Multiple significant contributions might mean that the contributor
-% should be moved to authorship at the discretion of the a
-%
-% See the policies ``Policies on Authorship'' section of https://livecoms.github.io for
-% more information on deciding on authorship and author order.
-%%%%%%%%%%%%%%%
-
-% (Explain the contributions of any non-author contributors here)
-% % We suggest you preserve this comment:
-% For a more detailed description of contributions from the community and others, see the GitHub issue tracking and changelog at \githubrepository.
 
@@
 J.S. is a founder and scientific advisor.
-J.S. also serves on the scientific advisory board of EoCys LLC.
+J.S. also serves on the scientific advisory board of MatchPoint Therapeutics.
 

### 2022-08-09 Old/2_prereq.tex

--- 2022-01-23 2_prereq.tex
@@
 \cite{Huang_Shen_2018_J.Chem.Inf.Model.,Harris_Shen_2019_J.Chem.Inf.Model.}
-is the best choice and Amber \cite{Case_Kollman_2018_}
+is the best choice and Amber \cite{Case_Kollman_2018}
 experience is required. 
@@
 We note, a GPU-accelerated all-atom PME CpHMD 
-implementation in Amber is forthcoming (Harris, Liu, and Shen, to be published). 
-Like all computational chemistry calculations, it is important to know your system as much as possible before any serious simulations. 
+implementation in Amber22 \cite{Case_Kollman_2022}
+has been recently released \cite{Harris_Shen_Submitted}.
+Like all computational chemistry calculations, it is important to know the system of interest
+before attempting any simulations. 
 With such knowledge, we can decide what salt concentration, temperature, pH range, and titratable residue types to use in CpHMD.
@@
  The next step is to add capping ligands ACE and NH2 to the N- and C-terminus and to modify ASP, GLU and HIS to AS2, GL2, and HIP residues if these residue types are set as titratable.
-For AS2 and GL2, a library file \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/blob/master/Files/phmd.lib}{phmd.lib} and a force field modification file \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/blob/master/Files/frcmod.phmd}{frcmod.phmd} are included in Ambertools after version 19 \cite{Case_Kollman_2018_}, or we can load them to tLeap/Leap from the directory containing the two files.
+For AS2 and GL2, a library file \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/blob/master/Files/phmd.lib}{phmd.lib} and a force field modification file \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep/-/blob/master/Files/frcmod.phmd}{frcmod.phmd} are included in Ambertools after version 19 \cite{Case_Kollman_2018}, or we can load them to tLeap/Leap from the directory containing the two files.
 The last step is to use tLeap/Leap in Amber/ Ambertools to generate topology and parameter files.

### 2022-08-09 Old/5_gphmd.tex

--- 2022-01-23 5_gphmd.tex
@@
 conjugate gradient algorithms for the first and second 500 steps, respectively. 
-To run minimization, we call \textbf{pmemd.MPI} or \textbf{pmemd.cuda} if Amber 20 \cite{Case_Kollman_2021_Amber20} or a later version is used.
+To run minimization, we call \textbf{pmemd.MPI} or \textbf{pmemd.cuda} if Amber 20 \cite{Case_Kollman_2020} or a later version is used.
 \begin{lstlisting}
@@
 $ $AMBERHOME/bin/pmemd.cuda -O
-  -i ${name}_pH7.5_prod.mdin
+  -i ${name}_pH7.50_prod.mdin
   -c ${name}_equil4.rst7 -p ${name}.parm7
-  -r ${name}_pH7.5_prod1.rst7
-  -o ${name}_pH7.5_prod1.out
-  -x ${name}_pH7.5_prod1.nc
+  -r ${name}_pH7.50_prod1.rst7
+  -o ${name}_pH7.50_prod1.out
+  -x ${name}_pH7.50_prod1.nc
   -phmdin ${name}.phmdin
@@
   -phmdstrt ${name}_equil4.phmdrst
-  -phmdout ${name}_pH7.5_prod1.lambda
+  -phmdout ${name}_pH7.50_prod1.lambda
   -phmdrestrt ${name}_pH7.5_prod1.phmdrst
-  -inf pH7.5.mdinfo
+  -inf pH7.50.mdinfo
 \end{lstlisting}
